---
title: "Potential and Limitations of ComBat Harmonization for Heterogenous Neuroimaging Data"
author: "Laurenz Lammer"
format:
  pdf:
    papersize: a4
    keep-tex: true
    header-includes:
      - \usepackage{flafter}
      - \usepackage[section]{placeins}
crossref:
  tbl-prefix: "Table"
  fig-prefix: "Figure"
bibliography: references.bib
editor:
  markdown:
    wrap: sentence
---

# Introduction

In the pursuit of a deeper understanding of the causes, correlates and trajectories of brain structure and function ever greater samples are required[@poldrack_brain_2016].
Hence, numerous multicentre neuroimaging studies have been conducted to facilitate the recruitment and scanning of more participants[@van_horn_multisite_2009].
Even larger sample sizes can be attained by conducting integrative data analyses (IDA) in which data from multiple studies are pooled[@curran_integrative_2009].
For example, Bethlehem et al. recently aggregated over 100,000 brain MRIs to develop brain charts across the human lifespan[@bethlehem_brain_2022].

A central challenge in multicentre studies is that images acquired on different scanners differ systematically[@medawar_estimating_2021].
To confront this concern, several data-harmonization approaches have been established.
Many rely on *travelling brains*, i.e. participants that undergo imaging on multiple scanners [@maikusa_comparison_2021; @yamashita_harmonization_2019; @tian_deep_2022].
However, this is a logistical challenge for any study and impossible when datasets are integrated post-hoc.

Therefore, ComBat, a method originally developed in molecular biology, that does not require travelling participants, has been applied widely [@fortin_harmonization_2018].
For example, all researchers using data of the ENIGMA consortium have been advised to employ it[@radua_increased_2020].
Multiple investigations concluded that ComBat successfully harmonizes data from multiple scanners while preserving the results of meaningful differences like distinct age and gender compositions of samples[@tassi_assessment_2024; @sun_comparison_2022; @eshaghzadeh_torbati_multi-scanner_2021].
Ongoing developments have further advanced the method to handle longitudinal data and to model nonlinear relationships more flexibly[@pomponio_harmonization_2020; @beer_longitudinal_2020].

Although these assessments are very promising, they lack a ground truth against which ComBat's performance could be evaluated.
Instead, quality was appraised based on ComBat's ability to detect significant case-control differences[@sun_comparison_2022] or the significance of age effects on cortical thickness[@beer_longitudinal_2020; @fortin_harmonization_2018].
Measurements of the same participant from multiple sites in close temporal proximity would provide a ground truth.
They should agree perfectly within the limits of measurement error if no systematic differences are present.
Yet, to our knowledge, only one study has thus far examined ComBat's performance in a sample of 16 individuals scanned on two scanners a few months apart[@eshaghzadeh_torbati_multi-scanner_2021].

Moreover, if cerebral measurements pooled in an IDA have been derived using different segmentation algorithms, an additional complication arises.
Volumetrics estimated using different software exhibit even larger discrepancies than those derived using the same software on different scanners[@burkle_scan-rescan_2025].

Therefore, we investigate whether ComBat successfully harmonizes derivatives with compounded scanner and software heterogeneity in a large travelling brain dataset.
Moreover, we provide the first independent evaluation of the performance of ComBat in longitudinal neuroimaging data and its application to white matter hyperintensity volume.

# Methods {#sec-methods}

## Preregistration and code

We preregistered this study at [osf](https://osf.io/7dcuh/overview).
Deviations were limited to the use of absolute value bars in the calculation of percent differences that were missing in the preregistration.
We also conducted multiple addtional explicitly labelled exploratory analyses.
All code used to process, analyse and visualize the data is available on [github]().
All analyses were performed in R version 4.5.0 except for neuroHarmonize which was applied in Python version 3.11.2.

```{r life_age_gender}
#| echo: false
#| warning: false
#| message: false
#| error: false
life_age_gender <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/life_age_gender.csv")
```

## Samples

To test ComBat’s performance in case of compounded scanner and segmentation software effects, we used the scanner-update dataset by Medawar et al. [@medawar_estimating_2021].
This sample consists of 121 healthy participants (median age = 28 years, range = 19–54 years; 61 women) that were scanned on two different 3 Tesla MRI scanners.
The time between sessions ranged from 0.5 to 18 weeks (median = 7.7 weeks).\
Additionally, we used data from the longitudinal LIFE-Adult neuroimaging sample[@engel_cohort_2023] to evaluate the performance of ComBat in longitudinal data.
At baseline, \~2600 participants underwent MR imaging (median age = `{r} round(median(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)` years, range = `{r} round(min(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)`-`{r} round(max(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)` years; `{r} round(mean(life_age_gender[life_age_gender$timepoint == "BL", "GENDER"], na.rm = T), digits = 2)*100`% women) of which \~1600 returned for follow-up imaging after \~6 years.
Both studies were approved by the ethics committee of the Faculty of Medicine at the University of Leipzig and all participants gave written informed consent according to the Declaration of Helsinki.

## Neuroimaging

All scans were obtained on 3 Tesla MRI scanners.
We acquired T1-weighted MPRAGE and fluid-attenuated inversion-recovery (FLAIR) sequences.
In the scanner-update study, participants were scanned on two different 3 Tesla Siemens MRI scanners: MAGNETOM VERIO syngo MR B17A and MAGNETOM SKYRA fit syngo MR E11C. Five participants did not undergo scanning on the SKYRA scanner.
In the LIFE-Adult study, all participants were scanned on the same 3 Tesla Siemens VERIO scanner.
Detailed acquisition parameters have been reported previously[@medawar_estimating_2021; @lammer_generalized_2025].

## Segmentation

We aimed to reuse existing data to minimise research waste and analyse segmentations that provide an overview of widely applied methods with a varying degree of similarity.
In addition to the original FreeSurfer version 6.0.0p1 longitudinal recon-all pipeline employed by Medawar et al., we segmented the MRIs using FreeSurfer 7.4.0 recon-all (cross-sectional), FreeSurfer 7.4.0 samseg (cross-sectional), and SPM 12.7771 CAT-12.9 (cross-sectional).
These four pipelines across two scanners resulted in eight batches of scanner-segmentation combinations.
All processing was performed with default settings on T1-weighted images.
For an exploratory analysis of the harmonizability of white matter hyperintensity volume (WMHV) from different scanners, we additionally used FLAIR images as inputs to FreeSurfer's version 7.4.0 cross-sectional samseg.\
In the LIFE-Adult dataset, we reused existing segmentations from FreeSurfer’s version 5.3.0 longitudinal recon-all pipeline (using only T1-weighted images) and FreeSurfer’s version 7.4.1 longitudinal samseg (using both T1-weighted and FLAIR images)[@lammer_impact_2023; @lammer_generalized_2025].
To investigate the harmonizability of WMHV from different software, we compared the measurements of samseg with those of the longitudinal pipeline of the Lesion Segmentation Toolbox version 3.0.0[@beyer_progression_2024].
For further exploratory analyses, we additionally analysed data from FreeSurfer's cross-sectional recon-all and samseg pipeline with the same versions as above.

## Derivation of composite volumes

We derived volumetric measures of the following regions of interest (ROIs):

-   Total intracranial volume (TIV) - unit: $dm^3$

-   Total grey matter volume (TGMV) - unit: $100 cm^3$

-   Total white matter volume (TWMV) - unit: $100 cm^3$

-   Total cortical volume (TCV) - unit: $100 cm^3$

-   Lateral ventricle volume (LVV) - unit: transformed

-   Hippocampal volume (HCV) - unit: $cm^3$

-   Amygdala volume (AV) - unit: $cm^3$

-   White matter hyperintensity volume (WMHV) - unit: transformed

We summed all bilateral measures over both hemispheres.
Further details on the derivation of composite volumes are available in the [preregistration](https://osf.io/7dcuh/overview).
TIV was not assessed in LIFE-Adult as this measure was not expected to change over time.
Furthermore, in longitudinal recon-all, TIV is always identical in all observations.
Thus, comparing TIV in the two measures based on this pipeline would not make sense.
Additionally, TWMV is not available in FreeSurfer's version 5.3.0 segmentation.

## Harmonization

We harmonized the scanner-update data using ComBat as implemented in neuroHarmonize[@pomponio_harmonization_2020] with gender as a categorical and age as a smooth covariate.
In the LIFE-Adult study, we used both longitudinal ComBat[@beer_longitudinal_2020] with age and gender as linear covariates and neuroHarmonize with the same covariates as above.
As suggested by the authors of neuroHarmonize, we first trained the model on our baseline data and then also applied it to follow-up observations.
In an exploratory analysis of the LIFE-Adult data, we used both methods to harmonize change scores rather than the observed values.
The shared participant identity of observations across batches was not modelled.

## Exclusion criteria, sampling weights and imputation

We included all participants of the scanner-update dataset.
For LIFE-Adult, we included all participants with available segmentations if their scans were labelled as usable by a trained clinician.
For WMHV values, we excluded participants whose scans had been flagged in visual quality control.
This included motion-artifacts, ventricular expansion which led to regression of lesion voxels, or brain pathologies such as stroke or congenital lesions.
We did not exclude outliers, use sampling weights or imputed missing data.
In structural equation models (SEMs), we used full information maximum likelihood.

## Transformations

LVV and WMHV were not normally distributed.
The former could be transformed with an inverse hyperbolic sine transformation.
This was not successful for WMHV.
Therefore, we used the bestnormalize function to perform the optimal non-standardizing transformation excluding the ordered quantile technique.
This was a Box-Cox transformation with lambda = -0.44 in LIFE-Adult and the decadic logarithm in the scanner-update data.

## Analyses

We employed a combination of statistical methods to evaluate the agreement and the magnitude and significance of differences between batches before and after harmonization.
In the scanner-update study, we drew inferences on absolute volume measures.
We used the LIFE-Adult study primarily to investigate ComBat's performance in longitudinal datasets.
Hence, we made inferences on the agreement on change scores calculated before and after harmonization here.

### Intraclass correlation coefficients

We calculated intraclass correlation coefficients (ICCs) for each ROI using the absolute agreement two-way random-effects model without averaging across raters.
This allows us to generalize the results to other scanner-segmentation combinations than the ones present in this study[@mcgraw_forming_1996].
Classical ICC calculations assume independent batches.
This is violated in the scanner-update dataset when batches share scanners or segmentations.
We accounted for this by estimating separate random effects for scanner and segmentation.

$$
volume = \beta_0 + id + scanner + segmentation + \varepsilon
$$

$$
\begin{aligned}
id \sim \mathcal{N}(0, \sigma^2_{id}), \quad
scanner \sim \mathcal{N}(0, \sigma^2_{scanner}), \\
segmentation \sim \mathcal{N}(0, \sigma^2_{segmentation}), \quad
\varepsilon \sim \mathcal{N}(0, \sigma^2_{\varepsilon})
\end{aligned}
$$

For each ROI, we estimated an overall intercept $\beta_0$ and three random intercepts (individual, scanner and segmentation) and an error term with lmer from lme4.
Then we calculated the ICC as:

$$
ICC = \frac{\sigma^2_{id}}{\sigma^2_{id} + \sigma^2_{scanner} + \sigma^2_{segmentation} + \sigma^2_{\varepsilon}}
$$

We derived confidence intervals by bootstrapping 1000 times.
For the LIFE-Adult dataset we followed the same procedure but did not have to model scanner as an additional effect.

### Percent differences

We calculated the percent difference (PD) to quantify the size and direction of the difference between batches for each ROI j across n participants according to:

$$
PD_j = 100 * \frac{2}{n} \sum_{i=1}^{n} \frac{V_{ij} - S_{ij}}{|V_{ij}| + |S_{ij}|}
$$

where $V_{ij}$ and $S_{ij}$ are the volumes of participant i in two different batches.
Absolute value bars are necessary to avoid that changes of a different sign cancel each other out.

### Significance tests

To test for significant differences, we calculated repeated-measures two-sided ANOVAs for the scanner-update dataset and two-sided paired t-tests for the LIFE-Adult data.
The ANOVAs were followed up with pairwise two-sided paired t-tests with Hommel correction.

### Intraclass effect decomposition

Brandmaier et al. have introduced intraclass effect decomposition (ICED) which uses structural equation modelling (SEM) to discern individual sources of variation[@brandmaier_assessing_2018].
This approach allowed us to identify the contributions of individual differences, scanners, segmentations and measurement error to the variance in the scanner-update dataset.
We modelled a SEM in which all eight observed variables indicated one *true* latent variable.
Additionally, all observed variables from one scanner indicated this scanner's latent variable (two latent scanner variables) and all from one segmentation indicated the corresponding segmentation's latent variable (four latent segmentation variables).
We modelled all observations to have one error term with a shared variance.
All variances were orthogonal to one another.
We constrained all factor loadings to 1 and estimated a mean for the *truth*, the second scanner and the second to fourth segmentation latent variables.
The ICED-ICC for each scanner-segmentation combination was calculated as:

$$
ICC = \frac{\sigma^2_{truth}}{\sigma^2_{truth} + \sigma^2_{scanner} + \sigma^2_{segmentation} + \sigma^2_{\varepsilon}}
$$

If a variance estimate was negative with a standard error larger than its distance to 0, we set it to 0 for the ICC calculation.

## Effect Size

We preregistered that an ICC \> 0.75 would be our primary criterion for sufficient harmonization.
Moreover, we considered a PD \< 2% and a p-value \>= 5% to be negligible.

## Statistical Power

A priori power analyses indicated that we had \>99% power to detect small effects (d = 0.2 / $\eta^2$ = 0.01) for all our samples.

## Sensitivity analyses

As a sensitivity analyses, we tested whether harmonization success is sensitive to manual edits.
We excluded manually edited segmentations (scanner-update 6.0.0p; LIFE-Adult 5.3.0), harmonized the reduced datasets, and re-ran our analyses.

## Exploratory analyses

We conducted five exploratory analyses.

1.  To further probe the effects of harmonization in longitudinal data, we tested for additive and multiplicative batch effects before and after harmonization with longitudinal ComBat.
    To that end, we used longCombat's addTest and multTest functions with random intercepts per participant and controlling for age and gender.

2.  We tested if ComBat is sensitive to the template generation employed in longitudinal pipelines.
    Hence, we evaluated its performance on the cross-sectional pipelines in the LIFE-Adult data.

3.  We tested the behaviour of neuroHarmonize when harmonizing change scores directly rather than absolute values from which change scores are calculated subsequently.

4.  We evaluated the success of ComBat in harmonizing absolute values (rather than changes).

5.  We investigated whether we could harmonize WMHV if only scanner differences are present.

# Results {#sec-results}

## Preregistered analyses

### Scanner-update

```{r update_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/outcomes.RData")
update_lme_icc_res <- outcomes$update_lme_icc_res
update_lme_icc_res_no_wmhv <- update_lme_icc_res[update_lme_icc_res$ROI != "WMHV",]
pd_mat <- outcomes$update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- outcomes$update_anova_res
pwcs <- outcomes$update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc")]]
  post <- pwcs[[paste0(outcome, "_pwc_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p <- ifelse(anovas$p < 0.001, formatC(anovas$p, format = "e", digits = 2), anovas$p)
anovas$p_adj <- ifelse(anovas$p_adj < 0.001, formatC(anovas$p_adj, format = "e", digits = 2), anovas$p_adj)
iced <- outcomes$update_iced_res
iced <- iced[iced$ROI != "TIV",]
iced[, 2:ncol(iced)] <- apply(iced[, 2:ncol(iced)], 2, function(x) ifelse(abs(as.numeric(x)) < 0.001 & as.numeric(x) != 0, formatC(as.numeric(x), format = "e", digits = 0), round(as.numeric(x), digits = 3)))
```

Prior to harmonization, ICCs varied widely across ROIs (range: `r round(min(update_lme_icc_res_no_wmhv$ICC), digits = 2)` - `r round(max(update_lme_icc_res_no_wmhv$ICC), digits = 2)`).
ANOVAs returned statistically significant batch differences for all volumes.
Likewise, average percent differences in batch comparisons exceeded 2% in all ROIs.
Differences between segmentations were greater than those between scanners (see @tbl-between-within).\
After harmonization, ICCs improved substantially (range: `r round(min(update_lme_icc_res_no_wmhv$ICC_adj), digits = 2)` - `r round(max(update_lme_icc_res_no_wmhv$ICC_adj), digits = 2)`, see @tbl-update-icc) and ANOVAs revealed no statistically significant differences.
Likewise, no pairwise comparison of two batches was significant after harmonization (see @tbl-update-signif).
The highest average percent difference was `r pd_sub2[which.max(pd_sub2$mean_pd_post), "ROI"]` with `r round(pd_sub2[which.max(pd_sub2$mean_pd_post), "mean_pd_post"], digits = 2)`% (see @tbl-update-pd).\
The ICED analyses produced ICCs \> 0.75 for all scanner-segmentation combinations across all ROIs.
However, this SEM-based metric was largely unaltered after harmonization
Neither scanner showed markedly higher reliability (see @tbl-iced-icc).
\
ComBat did not shift variances.
However, scanner and segmentation means were all close to 0 afterwards indicating the reduction of systematic batch differences (see @tbl-iced-variances and @tbl-iced-means).\
@fig-update to @fig-update4 illustrate the distribution of measurements in the batches before and after harmonization.

```{r}
#| label: tbl-update-icc
#| tbl-cap: "ICCs before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC, ICC_lower_95_CI, ICC_upper_95_CI),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_adj, ICC_lower_95_CI_adj, ICC_upper_95_CI_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricles volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p", "p_adj", "n_sig_pre", "n_sig_post")] |>
  kbl(
    col.names = c(
      "ROI",
      "p-value",
      "p-value harm.",
      "sig. PWCs",
      "sig. PWCs harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; p-value = p-value from ANOVA; harm. = harmonized with neuroHarmonize; ",
      "sig. PWCs = share of significant pairwise comparisons; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-pd
#| tbl-cap: "Percent differences before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percent difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: fig-update
#| fig-cap: "Distribution of measurements by batch before and after harmonization. \\newline TIV = Total Intracranial Volume; TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
#| dpi: 900
library(tidyverse)
library(MetBrewer)
library(patchwork)
library(showtext)
showtext_auto()
colors = met.brewer(name="Java", n=5, type="discrete")
colors <- colors[c(1,2,4,5)]
# load the unadjusted data from the scanner update study 
df_update <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/df_update.csv")
# break each scanner-segmentation combination down to a scanner and a segmentation variable
df_update$scanner <- ifelse(grepl("VER", df_update$SITE), "VERIO", "SKYRA")
df_update <- df_update %>%
  mutate(segmentation = case_when(
    grepl("samseg", SITE) ~ "samseg",
    grepl("cat", SITE) ~ "cat",
    grepl("recon6", SITE) ~ "recon6",
    grepl("recon7", SITE) ~ "recon7"
  ))
# for unscaled age data
QA <- read.csv("/data/pt_nro186_lifeupdate/Data/FREESURFER/QA/QA_list_short.csv")
# remove empty rows
# remove empty rows
QA <- QA[1:121,]
# create id col, add 0 that might got lost at the beginning/end, remove ., add "sub-"
# first add 0 lost at the end
QA$ID <- ifelse(nchar(sub(".*\\.", "", QA$ID)) == 1, paste0(QA$ID, "0"), QA$ID)
QA$id <- ifelse(nchar(QA$ID) != 7, gsub("\\.", "", paste0("sub-" ,QA$ID)), gsub("\\.", "", paste0("sub-", "0", QA$ID)))

# trim trailing spaces
QA$id <- trimws(QA$id)
QA$correction.necessary <- ifelse(is.na(QA$correction.necessary), 0, QA$correction.necessary)
# rename German colnames
QA <- QA %>%
  rename(
    "age" = "Alter_aktuell",
    "gender" = "Geschlecht"
  )
df_update <- df_update %>%
  select(id, TIV, TGMV, TWMV, TCV, LVV, HCV, AV, SITE) %>%
  left_join(QA[,c("id", "correction.necessary", "age", "gender")], by = "id")

# load the harmonized data from the scanner update study 
df_update_adj <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/df_update_adj.csv", col.names = colnames(df_update)[2:8])
# fill in the missing data from the unadjusted df (these cols must not be included during harmonization) 
df_update_adj[,c("id", "SITE", "age", "gender")] <- df_update[,c("id", "SITE", "age", "gender")]
df_update$SITE <- gsub("_", " ", df_update$SITE)
df_update_adj$SITE <- gsub("_", " ", df_update_adj$SITE)
df_update$SITE <- factor(df_update$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_update_adj$SITE <- factor(df_update_adj$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_update_wmhv <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/wmhv_update.csv")
df_update_wmhv_adj <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/wmhv_update_adj.csv")
df_update_wmhv$SITE <- ifelse(df_update_wmhv$scanner == "skyra", "SKYRA samseg", "VERIO samseg")
df_update_wmhv$SITE <- factor(df_update_wmhv$SITE, levels = c("SKYRA samseg", "VERIO samseg"))
df_update_wmhv_adj[,c("id", "SITE")] <- df_update_wmhv[,c("id", "SITE")]
df_update_wmhv$WMHV <- df_update_wmhv$WMHV_norm
df_update_wmhv_adj$WMHV <- df_update_wmhv_adj$X..WMHV_norm  
plots <- list()
comparisons <- list()
dfs <- list(df_update = df_update, df_update_adj = df_update_adj)
wmhv_dfs <- list(df_update_wmhv = df_update_wmhv, df_update_wmhv_adj = df_update_wmhv_adj)
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV", "HCV", "AV")){
  for (name in names(dfs)){
    df <- dfs[[name]]
    suffix <- ""
    y_axis_title <- outcome
    if (grepl("adj", name)){
      suffix <- "_adj"
      y_axis_title <- paste0(outcome, " harmonized")
    }
    ordered_ids <- df %>%
      group_by(id) %>%
      summarise(mean_val = mean(.data[[outcome]], na.rm = TRUE)) %>%
      arrange(mean_val) %>%
      pull(id)
    df$id <- factor(df$id, levels = ordered_ids)
    plt <- ggplot(data = df, aes(x = factor(id), y = .data[[outcome]], color = SITE, group = SITE, shape = SITE)) +
      geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
      scale_shape_manual(values = c(16, 6, 16, 6, 16, 6, 16, 6)) +
      theme_classic(base_family = "Charter") +
      guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) + 
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.1, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.3, "cm"), legend.key.spacing.y = unit(0, "cm"), legend.key.spacing.x = unit(0, "cm"), 
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
      ylab(y_axis_title) + scale_color_manual(values = rep(colors, each = 2))
    if (grepl("adj", name)){
      plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
    }
    plots[[paste0(outcome, suffix)]] <- plt
  }
  y_range <- range(c(plots[[outcome]]$data[[outcome]], plots[[paste0(outcome, "_adj")]]$data[[outcome]]), na.rm = TRUE)
  plots[[outcome]] <- plots[[outcome]] + coord_cartesian(ylim = y_range)
  plots[[paste0(outcome, "_adj")]] <- plots[[paste0(outcome, "_adj")]] + coord_cartesian(ylim = y_range)
  comparison <- wrap_plots(list(plots[[outcome]], plots[[paste0(outcome, "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
  comparisons[[outcome]] <- comparison
}
for (name in names(wmhv_dfs)){
  df <- wmhv_dfs[[name]]
  suffix <- ""
  y_axis_title <- "WMHV"
  if (grepl("adj", name)){
    suffix <- "_adj"
    y_axis_title <- paste0("WMHV harmonized")
  }
  ordered_ids <- df %>%
    group_by(id) %>%
    summarise(mean_val = mean(.data[["WMHV"]], na.rm = TRUE)) %>%
    arrange(mean_val) %>%
    pull(id)
  df$id <- factor(df$id, levels = ordered_ids)
  plt <- ggplot(data = df, aes(x = factor(id), y = .data[["WMHV"]], color = SITE, group = SITE, shape = SITE)) +
    geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
    scale_shape_manual(values = c(16, 6)) +
    theme_classic(base_family = "Charter") +
    guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.1, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.3, "cm"), legend.key.spacing.y = unit(0, "cm"), legend.key.spacing.x = unit(0, "cm"), 
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
    ylab(y_axis_title) + scale_color_manual(values = rep(colors[3], 2))
  if (grepl("adj", name)){
    plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
  }
  plots[[paste0("WMHV", suffix)]] <- plt
}
y_range <- range(c(plots[["WMHV"]]$data[["WMHV"]], plots[[paste0("WMHV", "_adj")]]$data[["WMHV"]]), na.rm = TRUE)
plots[["WMHV"]] <- plots[["WMHV"]] + coord_cartesian(ylim = y_range)
plots[[paste0("WMHV", "_adj")]] <- plots[[paste0("WMHV", "_adj")]] + coord_cartesian(ylim = y_range)
comparison <- wrap_plots(list(plots[["WMHV"]], plots[[paste0("WMHV", "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
comparisons[["WMHV"]] <- comparison

legend_only_plot <- ggplot() + 
  theme_void() +
  annotation_custom(cowplot::get_legend(comparisons[[1]]))
p_combined <- wrap_plots(comparisons[[1]], comparisons[[2]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update2
#| fig-cap: "Distribution of measurements by batch before and after harmonization. \\newline TWMV = Total White Matter Volume; TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
#| dpi: 900
p_combined <- wrap_plots(comparisons[[3]], comparisons[[4]],
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update3
#| fig-cap: "Distribution of measurements by batch before and after harmonization. \\newline LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
#| dpi: 900
p_combined <- wrap_plots(comparisons[[5]], comparisons[[6]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update4
#| fig-cap: "Distribution of measurements by batch before and after harmonization. \\newline AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
#| dpi: 900
p_combined <- wrap_plots(comparisons[[7]], comparisons[[8]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

\FloatBarrier

### LIFE-Adult

```{r life_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
life_lme_icc_res <- outcomes$life_lme_icc_res
life_lme_icc_res$ROI <- gsub("d", "\u0394", life_lme_icc_res$ROI)
pd_mat <- outcomes$life_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 6, ncol = 7))
colnames(pd_sub2) <- c("ROI", "pre", "nh", "combat", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_sub2$ROI <- c("\u0394TGMV", "\u0394TCV", "\u0394LVV",  "\u0394HCV", "\u0394AV", "\u0394WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_nh <- length(which(nh >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_combat <- length(which(combat >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "nh"] <- paste0(n_exceeding_nh, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "combat"] <- paste0(n_exceeding_combat, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pwcs <- outcomes$life_pwcs
pwcs$pwc_wmhv$measure <- "WMHV"
pwcs$pwc_wmhv_combat$measure <- "WMHV"
pwcs$pwc_wmhv_nh$measure <- "WMHV"
pwc <- rbind(pwcs$pwc[,!grepl("adj", colnames(pwcs$pwc))], pwcs$pwc_wmhv)
pwc_combat <- rbind(pwcs$pwc_combat[,!grepl("adj", colnames(pwcs$pwc_combat))], pwcs$pwc_wmhv_combat)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh[,!grepl("adj", colnames(pwcs$pwc_nh))], pwcs$pwc_wmhv_nh)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all$p <- ifelse(pwc_all$p < 0.001, formatC(pwc_all$p, format = "e", digits = 2), pwc_all$p)
pwc_all$p_nh <- ifelse(pwc_all$p_nh < 0.001, formatC(pwc_all$p_nh, format = "e", digits = 2), pwc_all$p_nh)
pwc_all$p_combat <- ifelse(pwc_all$p_combat < 0.001, formatC(pwc_all$p_combat, format = "e", digits = 2), pwc_all$p_combat)
pwc_all$measure <- gsub("d", "\u0394", pwc_all$measure)
```

ICCs for all ROIs except for change in HCV and LVV were below 0.75 both before and after harmonization (see @tbl-life-icc).
Likewise, PDs remained markedly above the 2% threshold (see @tbl-life-pd) and paired t-tests comparing the batches stayed highly significant for all ROIs except amygdala volume (see @tbl-life-signif).
This was the case both for data harmonized with longComBat and neuroHarmonize.
Bland-Altman plots depicting the differences between the two batches before and after harmonization are provided in @fig-ba-dtgmv to @fig-ba-dav.

```{r }
#| label: tbl-life-icc
#| tbl-cap: "ICCs before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
df2 <- life_lme_icc_res %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC, ICC_lower_95_CI, ICC_upper_95_CI
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh, ICC_lower_95_CI_nh, ICC_upper_95_CI_nh
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat, ICC_lower_95_CI_combat, ICC_upper_95_CI_combat
    )
  ) 
df2[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "\u0394 = change in",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-pd
#| tbl-cap: "Percent differences before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "> 2%",
      "> 2%  ~ neuroHarm",
      "> 2% ~ ComBat",
      "PD",
      "PD  ~ neuroHarm",
      "PD ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; > 2% = share of percent differences > 2%; ",
      "PD = percent difference;  ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume;"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all[, c("measure", "p", "p_nh", "p_combat")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "p-value  ~ neuroHarm",
      "p-value  ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

\FloatBarrier

## Sensitivity and exploratory analyses

Results after excluding manually edited scans were consistent with those of the full sample across all statistical tests (see @tbl-update-icc-no-correction to @tbl-life-signif-no-correction).\
The tests for remaining additive or multiplicative batch effects after harmonization showed that the additive batch effects were fully removed but multiplicative effects remained significant after harmonization in the longitudinal data (see @tbl-add-mult).\
Using measurements based on cross-sectional pipelines in the LIFE-Adult dataset did not lead to substantially different results (see @tbl-icc-cross).\
Harmonizing change scores directly led to moderately higher ICCs for some variables.
Yet, all ROIs except HCV and LVV remained below the threshold of 0.75 (see @tbl-icc-change).\
In contrast, if we consider the agreement between the harmonized absolute values at baseline and follow-up rather than the changes, ICCs for all volumes but WMHV are markedly higher.
All exceed 0.75 for neuroHarmonize and all but baseline AV for longitudinal ComBat (see @tbl-life-bl-icc and @tbl-life-fu-icc).
In addition, all PDs except for WMHV are substantially reduced after harmonization with lower values for neuroHarmonize at baseline and for longitudinal ComBat at follow-up (see @tbl-life-bl-pd and @tbl-life-fu-pd).
Similarly, there are no significant differences between batches at baseline after the application of neuroHarmonize (except WMHV).
Decreased but still significant differences at follow-up remain for all ROIs after both harmonization methods (see @tbl-life-bl-signif and @tbl-life-fu-signif).
We provide the corresponding Bland-Altman plots for observed values at each timepoint in the supplement (see @fig-ba-bl-tgmv to @fig-ba-fu-av).\
The agreement between WMHV using the same segmentation method on  different MRI machines showed an ICC of `r paste0(round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC"], digits = 3), " (95% CI: ", round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI"], digits = 3), "-", round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI"], digits = 3), ")")` before and of `r paste0(round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_adj"], digits = 3), " (95% CI: ", round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI_adj"], digits = 3), "-", round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI_adj"], digits = 3), ")")` after harmonization.
Measurements on the SKYRA were on average `r round(outcomes$update_pd_matrices$WMHV_norm_mean_pd_mat[1,2], digits = 1)`% larger before harmonization but only `r abs(round(outcomes$update_pd_matrices$WMHV_adj_mean_pd_mat_adj[1,2], digits = 1))`% smaller after the application of neuroHarmonize.
A paired t-test of the batches was significant prior to harmonization (p-value = `r formatC(outcomes$update_pwcs$WMHV_norm_pwc$p, digits = 2, format = "e")`; t-statistic = `r formatC(outcomes$update_pwcs$WMHV_norm_pwc$statistic, digits = 2, format = "g")`; df = `r outcomes$update_pwcs$WMHV_norm_pwc$df`) but ceased to be so afterwards (p-value = `r outcomes$update_pwcs$WMHV_adj_pwc_adj$p`; t-statistic = `r formatC(outcomes$update_pwcs$WMHV_adj_pwc_adj$statistic, digits = 2, format = "g")`; df = `r outcomes$update_pwcs$WMHV_adj_pwc_adj$df`).

\newpage

# Discussion

Integrative data analyses in neuroimaging require strategies to account for batch effects arising from heterogeneity in scanners and software.
Therefore, we evaluated the performance of ComBat methods in removing such effects across various ROIs using complementary statistical tests in cross-sectional and longitudinal designs.\
Our analyses provided fully satisfying results for all ROIs in cross-sectional analyses.
In contrast, ComBat did not achieve sufficient agreement of change scores.
This was no artifact resulting from template-generation in longitudinal pipelines or manual editing of scans.
Furthermore, the direct harmonization of change scores was no solution to this problem.
We attribute this result to two factors.\
Firstly, ComBat has been more effective in removing additive than multiplicative batch effects (@tbl-add-mult).
Additive effects affect measurements in the same batch uniformly and thus cancel out when computing change scores.
Hence, batch differences in change scores solely result from the more challenging multiplicative effects.\
Secondly, even over a follow-up period of six years, the individual variance in volumetric changes is small compared to the measurement error for many ROIs.
The ICED framework helps us to elucidate this constraint on harmonizability.
The results show that systematic differences between batches can be removed using ComBat(@tbl-iced-means).
Unsystematic measurement error is unamenable to harmonization, though (@tbl-iced-variances).
Hence, even in the absence of batch effects, high reliability is only attainable when true variance sufficiently exceeds error variance.
Therefore, acceptable ICCs were only observed for HCV and LVV, two structures that substantially atrophy or enlarge during ageing.
Analogously, longitudinal harmonizations were much more successful when evaluated on the scale of observed variables.
Only then can we successfully differentiate the signal of true variance from noise.
We have illustrated this by adding the extent of one standard deviation of the measured variable as a scale to Bland-Altman plots (@fig-ba-dtgmv to @fig-ba-fu-av).\
WMHV consistently exhibited the poorest agreement.
Even in the absence of segmentation software differences the ICC only narrowly exceeded our threshold after harmonization.
As differences between batches were small and statistically non-significant after harmonization, this must be attributed to the substantial measurement error visible in @fig-update4.

Our results confirm previous positive evaluations of ComBat methods on cross-sectional data[@tassi_assessment_2024; @sun_comparison_2022; @eshaghzadeh_torbati_multi-scanner_2021; @fortin_harmonization_2018; @pomponio_harmonization_2020; @yu_statistical_2018].
They provide particularly solid evidence by leveraging the large *travelling brain* scanner-update dataset to establish a ground truth scenario.
Importantly, this high reliability could be achieved without modelling an overlap sample between batches [@dewey_deepharmony_2019; @maikusa_comparison_2021; @yamashita_harmonization_2019; @tian_deep_2022].
Moreover, it was attained in spite of variable post-processing in the form of manual editing.\ 
However, our results also indicate concerningly low reliability when pooling data for the study of within-participant changes.
They thus highlight previously unrecognized limitations in longitudinal settings[@beer_longitudinal_2020].

## Limitations

We purposively chose statistical methods to make our results generalizable to other contexts.
However, we cannot assume with certainty that the results extend to all scanners, segmentations and their combinations.\
The high WMHV measurement error may be less pronounced in supervised algorithms like BIANCA[@torres-simon_decoding_2024].
However, samseg has been found to have high reliability when compared to manual segmentations on the volumetric level and to be more robust in handling multi-site data than BIANCA[@ferris_optimizing_2023].\
Furthermore, we cannot rule out that the observed problems in longitudinal harmonization would persist in the absence of segmentation software differences.
Given the success of ComBat in the cross-sectional setting this seems unlikely, though.
An empirical clarification would require a dataset of participants that underwent scanning in multiple scanners repeatedly with sufficient intervals for measurable longitudinal changes.
\
Lastly, it should be noted that harmonization techniques like ComBat exclusively focus on differences between scanners.
However, drift within scanners over time can also affect results[@takao_effect_2011].

## Conclusions

From our preregistered analyses, we derive the following conclusions: 1.
ComBat can be used to harmonize cross-sectional datasets with compounded scanner and segmentation effects.
2.
ComBat does not successfully harmonize longitudinal data to allow the study of individual changes across ROIs.

Our results highlight the importance of the size of true variance relative to measurement error in integrative data analyses.
We would thus recommend the use of models that allow modelling observed values at each timepoint as dependent variables in longitudinal settings.
Researchers could use linear mixed effects models to distinguish within- and between-participant effects of age[@van_de_pol_simple_2009].
If used at all, change scores should be restricted to datasets with extensive follow-up periods and for ROIs with substantial change.
Alternatively, harmonization involving *travelling brains* might be an option[@maikusa_comparison_2021] but require further evaluation in longitudinal data.\
In addition, the issue of measurement error is of particular significance for WMHV.
As this puts a consequential upper limit on the attainable reliability of measurements, we recommend only pooling homogeneously segmented data to minimise batch effects.
Implementing more laborious harmonization methods might be an option, too[@bordin_integrating_2021].

Integrating heterogenous neuroimaging data is increasingly necessary to push the frontiers of neuroscience[@zhu_statistical_2023].
Our results help to clarify when ComBat methods can be successfully employed to facilitate this.

# References {#sec-references}

::: {#refs}
:::

\newpage
\appendix

<!-- Switch to S-prefixed numbering -->

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\setcounter{figure}{0}

# Supplement {#sec-supplement}

## Supplementary tables

```{r scanner_software_comp}
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp <- data.frame(matrix(nrow = 7, ncol = 3))
colnames(scanner_software_comp) <- c("ROI", "within_scanner_diff", "within_software_diff")
scanner_software_comp$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
# start with scanner update
pd <- outcomes$update_pd_matrices
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  mat <- pd[[paste0(outcome,"_mean_pd_mat")]]
  transf_mat <- abs(mat) %>%
    as.data.frame() %>%
    mutate(row_name = rownames(.)) %>%
    pivot_longer(-row_name, names_to = "col_name", values_to = "value") %>%
    filter(row_name != col_name) %>%
    mutate(pair_id = map2_chr(row_name, col_name, ~ paste(sort(c(.x, .y)), collapse = "_"))) %>%
    distinct(pair_id, .keep_all = TRUE) %>%
    separate(row_name, into = c("metric1","scanner1","suffix1"), sep = "_") %>%
    separate(col_name, into = c("metric2","scanner2","suffix2"), sep = "_") 
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_scanner_diff"] <- transf_mat %>%
    filter(scanner1 == scanner2) %>%
    filter(suffix1 != suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_software_diff"] <- transf_mat %>%
    filter(scanner1 != scanner2) %>%
    filter(suffix1 == suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
}
scanner_software_comp[,2:3] <- round(scanner_software_comp[,2:3], digits = 2)
```

```{r}
#| label: tbl-between-within
#| tbl-cap: "Average percent differences between softwares and between scanners"
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "between software", "between scanners"),
    booktabs  = TRUE,
    linesep   = "",
    format    = "latex",
    align     = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; between software = average percent difference comparing two segmentations with different software of the same image;",
      "between scanners = average percent difference comparing two segmentations with identical software of images from different scanners"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

::: landscape
```{r}
#| label: tbl-iced-icc
#| tbl-cap: "ICED ICCs before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- lapply(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], as.numeric)
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- round(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], digits = 2)
iced[, c("ROI", paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "              = 1,
    "SKYRA cat"      = 2,
    "VERIO cat"      = 2,
    "SKYRA recon6"   = 2,
    "VERIO recon6"   = 2,
    "SKYRA recon7"   = 2,
    "VERIO recon7"   = 2,
    "SKYRA samseg"   = 2,
    "VERIO samseg"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; harm. = harmonized ICC; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-variances
#| tbl-cap: "ICED variances before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truevar", "SKYRAvar", "VERIOvar", "catvar", "recon6var", "recon7var", "samsegvar", "obsvar"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "SKYRA"    = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2,
    "samseg"   = 2,
    "residual"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; var = latent variable variance; harm. = harmonized latent variable variance; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-means
#| tbl-cap: "ICED means before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truemean", "VERIOmean", "catmean", "recon6mean", "recon7mean"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 10))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; mean = latent variable mean; harm. = harmonized latent variable mean; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```
:::

```{r update_prep_no_correction}
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_mat <- outcomes$update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- outcomes$update_anova_res
pwcs <- outcomes$update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc_no_correction")]]
  post <- pwcs[[paste0(outcome, "_pwc_no_correction_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p_no_correction <- ifelse(anovas$p_no_correction < 0.001, formatC(anovas$p_no_correction, format = "e", digits = 2), anovas$p_no_correction)
anovas$p_no_correction_adj <- ifelse(anovas$p_no_correction_adj < 0.001, formatC(anovas$p_no_correction_adj, format = "e", digits = 2), anovas$p_no_correction_adj)
```

```{r}
#| label: tbl-update-icc-no-correction
#| tbl-cap: "ICCs before and after harmonization without manually edited scans - Scanner-update dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction, ICC_lower_95_CI_no_correction, ICC_upper_95_CI_no_correction),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction_adj, ICC_lower_95_CI_no_correction_adj, ICC_upper_95_CI_no_correction_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-pd-no-correction
#| tbl-cap: "Percent differences before and after harmonization without manually edited scans - Scanner-update dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percentage difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif-no-correction
#| tbl-cap: "Statistical significance of differences before and after harmonization without manually edited scans - Scanner-update dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p_no_correction", "p_no_correction_adj", "n_sig_pre", "n_sig_post")] |>
  kbl(
    col.names = c(
      "ROI",
      "p-value",
      "p-value harm.",
      "sig. PWCs",
      "sig. PWCs harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; p-value = p-value from ANOVA; harm. = harmonized with neuroHarmonize; ",
      "sig. PWCs = share of pairwise comparisons' (t-tests) p-values < 0.05; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r life_prep_no_correction}
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_mat <- outcomes$life_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 5, ncol = 7))
colnames(pd_sub2) <- c("ROI", "pre", "nh", "combat", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_sub2$ROI <- c("\u0394TGMV", "\u0394TCV", "\u0394LVV",  "\u0394HCV", "\u0394AV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_no_correction")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_no_correction")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_nh <- length(which(nh >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_combat <- length(which(combat >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "nh"] <- paste0(n_exceeding_nh, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "combat"] <- paste0(n_exceeding_combat, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pwcs <- outcomes$life_pwcs
pwc <- pwcs$pwc_no_correction[,!grepl("adj", colnames(pwcs$pwc_no_correction))]
pwc_combat <- pwcs$pwc_combat_no_correction[,!grepl("adj", colnames(pwcs$pwc_combat_no_correction))]
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- pwcs$pwc_nh[,!grepl("adj", colnames(pwcs$pwc_nh_no_correction))]
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all$p <- ifelse(pwc_all$p < 0.001, formatC(pwc_all$p, format = "e", digits = 2), pwc_all$p)
pwc_all$p_nh <- ifelse(pwc_all$p_nh < 0.001, formatC(pwc_all$p_nh, format = "e", digits = 2), pwc_all$p_nh)
pwc_all$p_combat <- ifelse(pwc_all$p_combat < 0.001, formatC(pwc_all$p_combat, format = "e", digits = 2), pwc_all$p_combat)
pwc_all$measure <- gsub("d", "\u0394", pwc_all$measure)
```

```{r }
#| label: tbl-life-icc-no-correction
#| tbl-cap: "ICCs before and after harmonization without manually edited scans - LIFE-Adult dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
df2 <- life_lme_icc_res %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_no_correction, ICC_lower_95_CI_no_correction, ICC_upper_95_CI_no_correction
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_no_correction, ICC_lower_95_CI_nh_no_correction, ICC_upper_95_CI_nh_no_correction
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_no_correction, ICC_lower_95_CI_combat_no_correction, ICC_upper_95_CI_combat_no_correction
    )
  ) 
df2 <- df2[df2$ROI != "WMHV",]
df2[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "\u0394 = change in",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; "
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-pd-no-correction
#| tbl-cap: "Percent differences before and after harmonization without manually edited scans - LIFE-Adult dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "> 2%",
      "> 2%  ~ neuroHarm",
      "> 2% ~ ComBat",
      "PD",
      "PD  ~ neuroHarm",
      "PD ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; > 2% = share of percentage differences > 2%; ",
      "PD = percent difference;  ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume;"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-signif-no-correction
#| tbl-cap: "Statistical significance of differences before and after harmonization without manually edited scans - LIFE-Adult dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all[, c("measure", "p", "p_nh", "p_combat")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "p-value  ~ neuroHarm",
      "p-value  ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; "
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-add-mult
#| tbl-cap: "Statistical significance of additive and multiplicative batch effects before and after harmonization"
#| echo: false
#| warning: false
#| message: false
#| error: false
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/longcombat_tests.RData")
longCombat_results <- data.frame(matrix(ncol = 5, nrow =6))
colnames(longCombat_results) <- c("ROI", "add_pre", "add_post", "mult_pre", "mult_post")
longCombat_results$ROI <- c(longcombat_tests$addTestTable$Feature, "WMHV")
for (roi in longCombat_results$ROI){
  if (roi != "WMHV") {
    longCombat_results[longCombat_results$ROI == roi, "add_pre"] <- longcombat_tests$addTestTable[longcombat_tests$addTestTable$Feature == roi, "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "add_post"] <- longcombat_tests$addTestTable_combat[longcombat_tests$addTestTable_combat$Feature == paste0(roi, ".combat"), "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_pre"] <- longcombat_tests$multTestTable[longcombat_tests$multTestTable$Feature == roi, "p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_post"] <- longcombat_tests$multTestTable_combat[longcombat_tests$multTestTable_combat$Feature == paste0(roi, ".combat"), "p-value"]
  } else {
    longCombat_results[longCombat_results$ROI == roi, "add_pre"] <- longcombat_tests$addTestTable_wmhv[longcombat_tests$addTestTable_wmhv$Feature == "WMHV_norm", "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "add_post"] <- longcombat_tests$addTestTable_wmhv_combat[longcombat_tests$addTestTable_wmhv_combat$Feature == "WMHV_norm.combat", "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_pre"] <- longcombat_tests$multTestTable_wmhv[longcombat_tests$multTestTable_wmhv$Feature == "WMHV_norm", "p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_post"] <- longcombat_tests$multTestTable_wmhv_combat[longcombat_tests$multTestTable_wmhv_combat$Feature == "WMHV_norm.combat", "p-value"]
  }
}
longCombat_results$add_pre <- ifelse(as.numeric(longCombat_results$add_pre) < 0.001, formatC(as.numeric(longCombat_results$add_pre), format = "e", digits = 2), round(as.numeric(longCombat_results$add_pre), digits = 3))
longCombat_results$add_post <- ifelse(as.numeric(longCombat_results$add_post) < 0.001, formatC(as.numeric(longCombat_results$add_post), format = "e", digits = 2), round(as.numeric(longCombat_results$add_post), digits = 3))
longCombat_results$mult_pre <- ifelse(as.numeric(longCombat_results$mult_pre) < 0.001, formatC(as.numeric(longCombat_results$mult_pre), format = "e", digits = 2), round(as.numeric(longCombat_results$mult_pre), digits = 3))
longCombat_results$mult_post <- ifelse(as.numeric(longCombat_results$mult_post) < 0.001, formatC(as.numeric(longCombat_results$mult_post), format = "e", digits = 2), round(as.numeric(longCombat_results$mult_post), digits = 3))
longCombat_results %>%
  kbl(
    col.names = c("ROI", "p add.", "p add. harm.", "p mult.", "p mult. harm."),
    align = c("l", "r", "r", "r", "r"),
    format = "latex",
    booktabs = TRUE
  ) %>%
  kable_styling(
    latex_options = c("hold_position", "scale_down")
  ) %>%
  footnote(
    general = "ROI = region of interest; p add = p-value for additive effects; harm. = harmonized with longitudinal ComBat; p mult. = p-value for multiplicative effects; TGMV = total grey matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r life_cross_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$life_lme_icc_res
df$ROI <- gsub("d", "\u0394", df$ROI)
df2 <- df %>%
  mutate(
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_cross_nh, ICC_lower_95_CI_cross_nh, ICC_upper_95_CI_cross_nh
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_cross_combat, ICC_lower_95_CI_cross_combat, ICC_upper_95_CI_cross_combat
    )
  ) 
```

```{r }
#| label: tbl-icc-cross
#| tbl-cap: "ICCs before and after harmonization of results of cross-sectional pipelines"
#| echo: false
#| warning: false
#| message: false
#| error: false
df2[1:5, c("ROI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; \u0394 = change in; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r change_life_icc_table}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$change_life_lme_icc_res
colnames(df) <- paste0(colnames(df), "_change")
df2 <- cbind(df, outcomes$life_lme_icc_res[,c("ICC", "ICC_lower_95_CI", "ICC_upper_95_CI")])
df3 <- df2 %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC, ICC_lower_95_CI, ICC_upper_95_CI
    ),
    ICC_change_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_change, ICC_lower_95_CI_change, ICC_upper_95_CI_change
    )
  ) 
df3$ROI[nrow(df3)] <- "WMHV"
df3$ROI <- paste0("\u0394", df3$ROI)
```

```{r }
#| label: tbl-icc-change
#| tbl-cap: "ICCs before and after harmonization of directly harmonized change scores"
#| echo: false
#| warning: false
#| message: false
#| error: false
df3[, c("ROI", "ICC_CI", "ICC_change_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; \u0394 = change in; ",
      "TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r tp_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$tp_life_lme_icc_res
df_bl <- df %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_bl, ICC_lower_95_CI_bl, ICC_upper_95_CI_bl
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_bl, ICC_lower_95_CI_nh_bl, ICC_upper_95_CI_nh_bl
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_bl, ICC_lower_95_CI_combat_bl, ICC_upper_95_CI_combat_bl
    )
  ) 
df_bl$ROI[nrow(df_bl)] <- "WMHV"
df_fu <- df %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_fu, ICC_lower_95_CI_fu, ICC_upper_95_CI_fu
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_fu, ICC_lower_95_CI_nh_fu, ICC_upper_95_CI_nh_fu
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_fu, ICC_lower_95_CI_combat_fu, ICC_upper_95_CI_combat_fu
    )
  ) 
df_fu$ROI[nrow(df_fu)] <- "WMHV"

pd_mat <- outcomes$tp_life_pd_matrices

pd_bl <- data.frame(matrix(nrow = 6, ncol = 4))
colnames(pd_bl) <- c("ROI", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_bl$ROI <- c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_bl")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_bl")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_bl")]]
  pd_bl[pd_bl$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_bl[pd_bl$ROI == outcome, "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_bl[pd_bl$ROI == outcome, "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pd_bl[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")] <- round(pd_bl[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")], digits = 3)
pd_fu <- data.frame(matrix(nrow = 6, ncol = 4))
colnames(pd_fu) <- c("ROI", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_fu$ROI <- c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_fu")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_fu")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_fu")]]
  pd_fu[pd_fu$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_fu[pd_fu$ROI == outcome, "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_fu[pd_fu$ROI == outcome, "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pd_fu[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")] <- round(pd_fu[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")], digits = 3)
pwcs <- outcomes$tp_life_pwcs
pwcs$pwc_wmhv_bl$measure <- "WMHV"
pwcs$pwc_wmhv_combat_bl$measure <- "WMHV"
pwcs$pwc_wmhv_nh_bl$measure <- "WMHV"
# ignore "adjusted" cols (there was no adjustment)
pwc <- rbind(pwcs$pwc_bl[,!grepl("adj", colnames(pwcs$pwc_bl))], pwcs$pwc_wmhv_bl)
pwc_combat <- rbind(pwcs$pwc_combat_bl[,!grepl("adj", colnames(pwcs$pwc_combat_bl))], pwcs$pwc_wmhv_combat_bl)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh_bl[,!grepl("adj", colnames(pwcs$pwc_nh_bl))], pwcs$pwc_wmhv_nh_bl)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all_bl <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all_bl$p <- ifelse(pwc_all_bl$p < 0.001, formatC(pwc_all_bl$p, format = "e", digits = 2), pwc_all_bl$p)
pwc_all_bl$p_nh <- ifelse(pwc_all_bl$p_nh < 0.001, formatC(pwc_all_bl$p_nh, format = "e", digits = 2), pwc_all_bl$p_nh)
pwc_all_bl$p_combat <- ifelse(pwc_all_bl$p_combat < 0.001, formatC(pwc_all_bl$p_combat, format = "e", digits = 2), pwc_all_bl$p_combat)
pwcs$pwc_wmhv_fu$measure <- "WMHV"
pwcs$pwc_wmhv_combat_fu$measure <- "WMHV"
pwcs$pwc_wmhv_nh_fu$measure <- "WMHV"
# ignore "adjusted" cols (there was no adjustment)
pwc <- rbind(pwcs$pwc_fu[,!grepl("adj", colnames(pwcs$pwc_fu))], pwcs$pwc_wmhv_fu)
pwc_combat <- rbind(pwcs$pwc_combat_fu[,!grepl("adj", colnames(pwcs$pwc_combat_fu))], pwcs$pwc_wmhv_combat_fu)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh_fu[,!grepl("adj", colnames(pwcs$pwc_nh_fu))], pwcs$pwc_wmhv_nh_fu)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all_fu <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all_fu$p <- ifelse(pwc_all_fu$p < 0.001, formatC(pwc_all_fu$p, format = "e", digits = 2), pwc_all_fu$p)
pwc_all_fu$p_nh <- ifelse(pwc_all_fu$p_nh < 0.001, formatC(pwc_all_fu$p_nh, format = "e", digits = 2), pwc_all_fu$p_nh)
pwc_all_fu$p_combat <- ifelse(pwc_all_fu$p_combat < 0.001, formatC(pwc_all_fu$p_combat, format = "e", digits = 2), pwc_all_fu$p_combat)
```

```{r }
#| label: tbl-life-bl-icc
#| tbl-cap: "Baseline ICCs before and after harmonization in the LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df_bl[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] %>%
  kable(
    col.names = c("ROI", "ICC (95% CI)", "ICC (95% CI) ~ neuroHarm", "ICC (95% CI) ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; ICC = intraclass correlation coefficient; CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-fu-icc
#| tbl-cap: "Follow-up ICCs before and after harmonization in the LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df_fu[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] %>%
  kable(
    col.names = c("ROI", "ICC (95% CI)", "ICC (95% CI) ~ neuroHarm", "ICC (95% CI) ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; ICC = intraclass correlation coefficient; CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-bl-pd
#| tbl-cap: "Baseline percent differences before and after harmonization in the LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_bl %>%
  kable(
    col.names = c("ROI", "PD", "PD ~ neuroHarm", "PD ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(which(colnames(pd_sub2) %in% c("pre", "nh", "combat")), extra_css = "text-align: right;") %>%
  footnote(
    general = "ROI = region of interest; PD = percent difference; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; TGMV = total grey matter volume; TWMV = total white matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume",
    general_title = "",
    threeparttable = TRUE
    )
```

```{r }
#| label: tbl-life-fu-pd
#| tbl-cap: "Follow-up percent differences before and after harmonization in the LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_fu %>%
  kable(
    col.names = c("ROI", "PD", "PD ~ neuroHarm", "PD ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(which(colnames(pd_sub2) %in% c("pre", "nh", "combat")), extra_css = "text-align: right;") %>%
  footnote(
    general = "ROI = region of interest; PD = percent difference; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; TGMV = total grey matter volume; TWMV = total white matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-bl-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in the baseline LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all_bl[, c("measure", "p", "p_nh", "p_combat")] %>%
  kable(
    col.names = c("ROI", "p", "p ~ neuroHarm", "p ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; p = p-value; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-fu-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in the follow-up LIFE-Adult data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all_fu[, c("measure", "p", "p_nh", "p_combat")] %>%
  kable(
    col.names = c("ROI", "p", "p ~ neuroHarm", "p ~ ComBat"),
    format = "latex",
    booktabs = TRUE,
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; p = p-value; ~ neuroHarm = harmonized with neuroHarmonize; ~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

\FloatBarrier

## Supplementary figures

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dtgmv
#| fig-cap: "Bland-Altman plots of TGMV changes. Blue vertical lines indicate the size of one standard deviationof TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/ba_plots.RData")
print(ba_plots$wrapped_life_ba_plots[[paste0("dTGMV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dtcv
#| fig-cap: "Bland-Altman plots of TCV changes. Blue vertical lines indicate the size of one standard deviation of TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dTCV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dlvv
#| fig-cap: "Bland-Altman plots of LVV changes. Blue vertical lines indicate the size of one standard deviation of LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dLVV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dhcv
#| fig-cap: "Bland-Altman plots of HCV changes. Blue vertical lines indicate tghe size of one standard deviation of HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dHCV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dav
#| fig-cap: "Bland-Altman plots of AV changes. Blue vertical lines indicate the size of one standard deviation of AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dAV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-tgmv
#| fig-cap: "Bland-Altman plots of baseline TGMV. Blue vertical lines indicate the size of one standard deviation of baseline TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TGMV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-tcv
#| fig-cap: "Bland-Altman plots of baseline TCV. Blue vertical lines indicate the size of one standard deviation of baseline TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TCV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-lvv
#| fig-cap: "Bland-Altman plots of baseline LVV. Blue vertical lines indicate the size of one standard deviation of baseline LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("LVV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-hcv
#| fig-cap: "Bland-Altman plots of baseline HCV. Blue vertical lines indicate the size of one standard deviation of baseline HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("HCV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-av
#| fig-cap: "Bland-Altman plots of baseline AV. Blue vertical lines indicate tghe size of one standard deviation of baseline AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("AV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-tgmv
#| fig-cap: "Bland-Altman plots of follow-up TGMV. Blue vertical lines indicate the size of one standard deviation of follow-up TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TGMV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-tcv
#| fig-cap: "Bland-Altman plots of follow-up TCV. Blue vertical lines indicate the size of one standard deviation of follow-up TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TCV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-lvv
#| fig-cap: "Bland-Altman plots of follow-up LVV. Blue vertical lines indicate the size of one standard deviation of follow-up LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("LVV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-hcv
#| fig-cap: "Bland-Altman plots of follow-up HCV. Blue vertical lines indicate tghe size of follow-up HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("HCV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-av
#| fig-cap: "Bland-Altman plots of follow-up AV. Blue vertical lines indicate tghe size of one standard deviation of follow-up AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("AV", "_fu")]])
```
