---
title: ""
author: " "
date: ""
format:
  pdf:
    papersize: a4
    keep-tex: true
    template: custom-template.tex
    cite-method: biblatex
    header-includes:
      - \usepackage[backend=biber,style=apa]{biblatex}
      - \DeclareLanguageMapping{english}{english-apa}
      - \addbibresource{references.bib}
      - \usepackage{sectsty}
      - \allsectionsfont{\normalfont}
      - \usepackage{flafter}
      - \usepackage[section]{placeins}
      - \usepackage{tikz}
      - \usetikzlibrary{arrows, positioning, shapes, fit, backgrounds, calc, shapes.geometric}
      - \usepackage{authblk}
      - \renewcommand\Affilfont{\small}
      - \usepackage{orcidlink}
      - \date{}
    include-before-body:
      text: |
        \title{Potential and Limitations of ComBat for the Harmonization of Multi-Site Neuroimaging Studies}
        \author[1,2,3]{Laurenz Lammer\orcidlink{0000-0001-5612-4924}}
        \author[1,4]{Frauke Beyer\orcidlink{0000-0001-5401-852X}}
        \author[1,4,5]{A. Veronica Witte\orcidlink{0000-0001-9054-6688}}
        \author[1,3,4,5,6]{Arno Villringer\orcidlink{0000-0003-2604-2404}}
        \author[2,3,7]{Ulman Lindenberger\orcidlink{0000-0001-8428-6453}}
        \author[8,9]{Denis Gerstorf\orcidlink{0000-0002-2133-9498}}
        \author[2,7,10]{Andreas M. Brandmaier\orcidlink{0000-0001-8765-6982}}
        \affil[1]{Department of Neurology, Max Planck Institute for Human Cognitive and Brain Sciences, Leipzig, Germany}
        \affil[2]{Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany}
        \affil[3]{Max Planck School of Cognition, Leipzig, Germany}
        \affil[4]{Day Clinic for Cognitive Neurology, University Hospital Leipzig, Leipzig, Germany}
        \affil[5]{Leipzig Research Centre for Civilization Diseases (LIFE), Leipzig University, Leipzig, Germany}
        \affil[6]{MindBrainBody Institute, Berlin School of Mind and Brain, Humboldt University, Berlin, Germany}
        \affil[7]{Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany, and, London, UK}
        \affil[8]{Department of Psychology, Humboldt-Universität zu Berlin, Berlin, Germany}
        \affil[9]{German Institute for Economic Research (DIW), Socio-Economic Panel, Berlin, Germany}
        \affil[10]{Department of Psychology, MSB Medical School Berlin, Berlin, Germany}
        \maketitle
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
crossref:
  tbl-prefix: "Table"
  fig-prefix: "Figure"
bibliography: references.bib
editor:
  markdown:
    wrap: sentence
---

## Corresponding author

[Laurenz Lammer](mailto:lammer@cbs.mpg.de) (lammer\@cbs.mpg.de)

## Author contributions

Author roles were classified using the Contributor Role Taxonomy ([CRediT](https://credit.niso.org/) (`https://credit.niso.org/`)) as follows:

**Conceptualization**: Laurenz Lammer, Frauke Beyer, Ulman Lindenberger, Denis Gerstorf, Andreas M. Brandmaier\
**Methodology**: Laurenz Lammer, Frauke Beyer, Andreas M. Brandmaier\
**Investigation**: Frauke Beyer\
**Software, Formal analysis, Visualization**: Laurenz Lammer\
**Data Curation**: Laurenz Lammer, Frauke Beyer\
**Resources, Supervision**: A. Veronica Witte, Arno Villringer, Ulman Lindenberger\
**Funding acquisition**: A.
Veronica Witte, Arno Villringer\
**Writing – original draft**: Laurenz Lammer\
**Writing – review & editing**: all authors

\newpage
\pagenumbering{arabic}

# Abstract

Integrating multiple neuroimaging datasets is a key strategy to obtain the increasing sample sizes needed to advance neuroscience.
However, this requires addressing known biases due to differences in scanners and segmentations.
ComBat is the most widely used solution to address this problem.
Existing validations are promising but largely limited to scanner effects and mostly rely on indirect indicators of success.
Thus, we assessed ComBat's performance with compounded scanner and segmentation effects in this preregistered study.
We leveraged a travelling participant dataset (121 individuals scanned on 2 scanners and processed with 4 segmentation pipelines) and used another large dataset (n = \~1000 with 2 timepoints \~6 years apart) to evaluate harmonization in longitudinal data.
To assess the agreement between batches, we calculated intraclass correlation coefficients, percent differences and tests for the significance of differences.
Furthermore, we employed intraclass effect decomposition (ICED) to gauge the contributions of scanners, segmentations, individual differences and measurement error to total variance.
ICED illustrated how the upper limit of the attainable agreement is determined by the amount of random measurement error.
In the cross-sectional data, harmonization met all preregistered criteria (ICC \> 0.75; p-value \> 0.05; percent difference \< 2%).
In contrast, the harmonization of change scores was not satisfactory across all regions of interests.
When evaluated as observed values rather than changes, the results of the longitudinal harmonization with the exception of white matter hyperintensity volumes were acceptable, though.
The discrepancy of the results is primarily attributable to the size of true between-individual variance relative to measurement error.
Even with a follow-up period of six years, changes in most brain volumes are small compared to measurement error.
Overall, our results support the utility of ComBat for the harmonization of heterogenous neuroimaging data but also delineate when harmonization fails.
Further, our analyses demonstrate that the relative size of measurement error is a crucial determinant of the achievable agreement.
We conclude that ComBat can be applied confidently in cross-sectional settings but caution is warranted in longitudinal designs.

\newpage

# Introduction

In the pursuit of a deeper understanding of the causes, correlates and consequences of variations in brain structure and function ever greater samples are required [@poldrack_brain_2016].
Larger and more representative samples are needed to mitigate concerns of insufficient power [@button_power_2013], enhance the precision and reliability of our estimates [@cremers_relation_2017; @grady_influence_2021; @marek_reproducible_2022] and increase generalizability [@lewinn_sample_2017; @marek_replicability_2025].
Hence, numerous multi-site neuroimaging studies have been conducted to facilitate the recruitment and scanning of larger numbers of participants [@van_horn_multisite_2009].
Even larger sample sizes can be attained by conducting integrative data analyses (IDA) in which data from multiple studies are pooled [@curran_integrative_2009].
For example, Bethlehem et al. recently aggregated over 100,000 brain MRIs to develop brain charts across the human lifespan [@bethlehem_brain_2022].

Integrating data from multiple sites and studies also introduces new challenges, though.
A central concern in multi-site studies is that images acquired on different scanners differ systematically [@medawar_estimating_2021].
To address this, several data-harmonization approaches have been established.
Many rely on *travelling brains*, i.e. participants that undergo imaging on multiple scanners [@maikusa_comparison_2021; @yamashita_harmonization_2019; @tian_deep_2022].
Under the assumption that that the true parameters of interest do not change during these travels, this allows the immediate measurement of between-scanner differences.
However, this is a resource-intensive logistical challenge for any study and impossible when datasets are integrated post hoc.

In contrast, ComBat does not require travelling participants to remove additive and multiplicative batch effects of different scanners [@fortin_harmonization_2018].
Thus, it was readily adopted by the neuroimaging community [@gebre_crossscanner_2023; @horien_hitchhikers_2020].
For example, the ENIGMA consortium advises all researchers using its data to employ it [@radua_increased_2020].
Multiple investigations concluded that ComBat successfully harmonizes data from multiple scanners while preserving the results of meaningful differences like distinct age and gender compositions of samples [@tassi_assessment_2024; @sun_comparison_2022; @eshaghzadeh_torbati_multi-scanner_2021].
Ongoing developments have further advanced the method to handle longitudinal data and to model nonlinear relationships more flexibly [@pomponio_harmonization_2020; @beer_longitudinal_2020].

Although these assessments are very promising, they lack a ground truth against which ComBat's performance could be evaluated.
Instead, quality was appraised based on ComBat's ability to detect significant differences between cases and controls [@sun_comparison_2022; @radua_increased_2020] or the significance of established associations of age and cortical thickness [@beer_longitudinal_2020; @fortin_harmonization_2018].
Measurements of the same participant from multiple sites in close temporal proximity would provide a ground truth.
They should agree perfectly within the limits of measurement error if no systematic differences are present.
Yet, such designs have seldomly been employed in datasets of more than 50 participants [@gebre_crossscanner_2023; @richter_validation_2022].

Moreover, if cerebral measurements pooled in an IDA have been derived using different segmentation algorithms, an additional complication arises.
Volumetrics estimated using different software exhibit even larger discrepancies than those derived using the same software on different scanners [@burkle_scan-rescan_2025].
Even using cross-sectional or longitudinal segmentation pipelines of the same software version affects the results [@vidal-pineiro_reliability_2025].
Haddad et al. showed that significance patterns of age-cortical thickness associations are more homogenous after harmonization but did not evaluate the effect on agreement between batches [@haddad_multisite_2023].

Therefore, we investigated whether ComBat successfully harmonizes parameters of interest (such as hippocampal volume) in the presence of both scanner and software heterogeneity in a large *travelling brain* dataset.
Moreover, we provide an evaluation of the performance of ComBat in longitudinal neuroimaging data and its application to white matter hyperintensity volume.

# Methods {#sec-methods}

## Preregistration and code

We preregistered this study at [OSF](https://osf.io/7dcuh/overview) (`https://osf.io/7dcuh/overview`).
The only deviation from the preregistration was the addition of absolute value bars in the percent difference formula.
We also conducted multiple additional clearly labelled exploratory analyses.
All code used to process, analyze and visualize the data is available on [github](https://github.com/formal-methods-mpi/SegmentationHarmonizability) (`https://github.com/formal-methods-mpi/SegmentationHarmonizability`).
All analyses were performed in R version 4.5.0 except for neuroHarmonize which was applied in Python version 3.11.2.

```{r life_age_gender}
#| echo: false
#| warning: false
#| message: false
#| error: false
life_age_gender <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/life_age_gender.csv")
```

## Samples

To test ComBat’s performance in the case of compounded scanner and segmentation software effects, we used the scanner-update dataset by Medawar et al. 2021.
The data were originally collected to study the effects of a scanner update between waves of a longitudinal study [@medawar_estimating_2021].
This sample consists of 121 healthy participants (median age = 28 years, range = 19--54 years; 61 women) that were each scanned on two different 3 Tesla MRI scanners.
The time between sessions ranged from 0.5 to 18 weeks (median = 7.7 weeks).\
Additionally, we used data from the longitudinal LIFE-Adult neuroimaging sample [@engel_cohort_2023] to evaluate the performance of ComBat in longitudinal data.
At baseline, \~2600 participants underwent MR imaging (median age = `{r} round(median(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)` years, range = `{r} round(min(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)`--`{r} round(max(life_age_gender[life_age_gender$timepoint == "BL", "AGE"], na.rm = T), digits = 0)` years; `{r} round(mean(life_age_gender[life_age_gender$timepoint == "BL", "GENDER"], na.rm = T), digits = 2)*100`% women) of which \~1600 returned for follow-up imaging after \~6 years.
Our study's requirements allowed us to include \~1000 participants with two timepoints.
The samples are hereafter referred to as the cross-sectional and the longitudinal dataset.
Both studies were approved by the ethics committee of the Faculty of Medicine at the University of Leipzig and all participants gave written informed consent for being included in the study according to the Declaration of Helsinki.\

## Neuroimaging

All scans were obtained on 3 Tesla MRI scanners.
We used T1-weighted MPRAGE and fluid-attenuated inversion-recovery (FLAIR) sequences.
In the cross-sectional study, participants were scanned on two different 3 Tesla Siemens MRI scanners: MAGNETOM VERIO syngo MR B17A and MAGNETOM SKYRA fit syngo MR E11C. Five participants did not undergo scanning on the SKYRA scanner.
In the longitudinal study, all participants were scanned on the same 3 Tesla Siemens VERIO scanner.
Detailed acquisition parameters have been reported previously [@medawar_estimating_2021; @lammer_generalized_2025].

## Segmentation

In addition to the original FreeSurfer version 6.0.0p1 longitudinal recon-all pipeline [@reuter_within-subject_2012] employed by Medawar et al. 2021, we segmented the MRIs using FreeSurfer 7.4.0 recon-all (cross-sectional) [@fischl_whole_2002], FreeSurfer 7.4.0 samseg (cross-sectional) [@cerri_open-source_2023], and SPM 12.7771 CAT-12.9 (cross-sectional) [@gaser_cat_2024].
These four pipelines across two scanners resulted in eight batches of scanner-segmentation combinations.
All processing was performed with default settings on T1-weighted images.
For an exploratory analysis of the harmonizability of white matter hyperintensity volumes (WMHV) from different scanners, we additionally used FLAIR images as inputs to FreeSurfer's version 7.4.0 cross-sectional samseg.\
In the longitudinal dataset, we reused existing segmentations from FreeSurfer’s version 5.3.0 longitudinal recon-all pipeline (using only T1-weighted images) and FreeSurfer’s version 7.4.1 longitudinal samseg (using both T1-weighted and FLAIR images) [@lammer_impact_2023; @lammer_generalized_2025].
To investigate the harmonizability of WMHV from different software, we compared the measurements of samseg with those of the longitudinal pipeline of the Lesion Segmentation Toolbox version 3.0.0 [@beyer_progression_2024].
For further exploratory analyses, we additionally analysed data from FreeSurfer's cross-sectional recon-all and samseg pipeline with the same versions as above.

## Derivation of composite volumes

We derived volumetric measures of the following regions of interest (ROIs):

-   Total intracranial volume (TIV) - unit: $dm^3$

-   Total grey matter volume (TGMV) - unit: $100 cm^3$

-   Total white matter volume (TWMV) - unit: $100 cm^3$

-   Total cortical volume (TCV) - unit: $100 cm^3$

-   Lateral ventricle volume (LVV) - unit: $asinh(mm^3)$

-   Hippocampal volume (HCV) - unit: $cm^3$

-   Amygdala volume (AV) - unit: $cm^3$

-   White matter hyperintensity volume (WMHV) - unit: $log_{10}$($mm^3$)

We summed all bilateral measures over both hemispheres.
If no composite value for an ROI was provided by an atlas, we calculated it by summing all relevant substructures.
Further details on the derivation of composite volumes are available in the [preregistration](https://osf.io/7dcuh/overview) (`https://osf.io/7dcuh/overview`).
TIV was not analyzed in longitudinal data as this measure was not expected to change over time.
Furthermore, in longitudinal recon-all, TIV is always identical in all observations.
Additionally, TWMV is not available in FreeSurfer's version 5.3.0 segmentation.

## Harmonization

ComBat is a harmonization method developed to remove additive and multiplicative batch effects while preserving the effects of meaningful biological covariates in microarray expression data [@johnson_adjusting_2007].
Observed data are adjusted as:

$$
Y^*_{ijg} = \frac{Y_{ijg} - \hat{\alpha_g} - X\hat{\beta_g} - \hat{\gamma_{jg}}}{\hat{\delta_{jg}}} + \hat{\alpha_g} + X\hat{\beta_g} 
$$

where $Y^*_{ijg}$ are the adjusted values for gene *g* of sample *i* in batch *j*.
$\hat{\alpha_g}$ (the grand mean), $\hat{\beta_g}$ (the regression coefficients of biological covariates), $\hat{\gamma_{ig}}$ (the additive effect of a batch on a gene) and $\hat{\delta_{ig}}$ (the multiplicative effect of a batch on a gene) are estimated from the data.
$X$ is the matrix of covariate values.

The method was later adopted to harmonize MRI data where genes are replaced by neuroimaging metrics and each scanner is a batch [@fortin_harmonization_2018].
We harmonized the cross-sectional data using ComBat as implemented in `neuroHarmonize` [@pomponio_harmonization_2020].
The confounders' effects were estimated using a generalized additive model with gender as a dichotomous covariate and a smooth function for age.
In the longitudinal dataset, we used both `longComBat()` [@beer_longitudinal_2020] with age and gender as linear covariates and `neuroHarmonize()` with the same covariate adjustment as above.
As suggested by the authors of `neuroHarmonize`, we first trained the model on our baseline data and then also applied it to follow-up observations [@pomponio_neuroharmonize_2024].
The observations in the batches of each dataset came from the same participants.
As ComBat is intended to work when there is no overlap sample, we intentionally did not model this dependency between observations in the different batches.
In an exploratory analysis of the longitudinal data, we used both methods to harmonize change scores rather than the observed values.

## Exclusion criteria, sampling weights and imputation

We included all participants of the cross-sectional dataset.
In the longitudinal dataset, we included all participants with available segmentations if their scans were labelled as usable by a trained clinician.
For WMHV values, we excluded participants whose scans had been flagged in visual quality control.
This included motion-artifacts, ventricular expansion which led to regression of lesion voxels, or brain pathologies such as stroke or congenital lesions.
We did not exclude outliers, use sampling weights nor imputed missing data.
In structural equation models (SEMs), we used full information maximum likelihood estimation.

## Transformations

A visual inspection of histograms of the ROI volumes showed that LVV and WMHV did not follow a normal distribution.
LVV could be modified to follow an approximately normal distribution with an inverse hyperbolic sine transformation.
The same was not successful for WMHV.
Therefore, we used `bestnormalize()` [@peterson_finding_2021] to identify the optimal non-standardizing transformation excluding the ordered quantile technique.
This was a Box-Cox transformation with lambda = -0.44 in the longitudinal data and the decadic logarithm in the cross-sectional data.
We provide raincloud plots of original measurements for all ROIs and after transformations if applicable in @fig-update-tiv-raincloud to @fig-life-wmhv-raincloud.

## Analyses

We employed a combination of statistical methods to evaluate the agreement and the magnitude and significance of differences between batches before and after harmonization.
In the cross-sectional study, we drew inferences based on absolute volume measures.
We used the longitudinal study primarily to investigate ComBat's performance for developmental investigations with repeated measures.
Hence, we evaluated the agreement on change scores calculated before and after harmonization in this dataset.

### Intraclass correlation coefficients

The intraclass correlation coefficient indicates how much of the total variance is attributable to true between-participant differences rather than the scanner, segmentation, or measurement error.
A high ICC is desirable because it indicates good agreement of measurements across batches.
We calculated ICCs for each ROI using the absolute agreement two-way random-effects model without averaging across raters.
Absolute agreement rather than mere consistency is warranted because measurements from different batches should be approximately equal and not just proportional.
Averaging across raters is adequate if multiple measurements of the same participant (at the same timepoint) are available.
This is not the case in real data and would thus produce overly confident results if employed here.
We used the two-way random effects model as we assumed that participants, scanners and segmentations were drawn at random from underlying larger distributions.
This allows us to generalize the results to other scanner-segmentation combinations than the ones present in this study [@mcgraw_forming_1996].
As all ICC models, the model assumes that each scanner and segmentation contributes a different offset to the true value while the remaining noise is identically distributed.
Whether interactions between batches and individual participants are modelled or not is inconsequential to the obtained ICC value [@liljequist_intraclass_2019].
Classical ICC calculations assume independent batches.
This is violated in the cross-sectional dataset when batches share scanners or segmentations.
We accounted for this by estimating separate random effects for scanner and segmentation.
For each ROI we used the following model:

$$
volume = \beta_0 + id + scanner + segmentation + \varepsilon
$$

$$
\begin{aligned}
id \sim \mathcal{N}(0, \sigma^2_{id}), \\
scanner \sim \mathcal{N}(0, \sigma^2_{scanner}), \\
segmentation \sim \mathcal{N}(0, \sigma^2_{segmentation}), \\
\varepsilon \sim \mathcal{N}(0, \sigma^2_{\varepsilon})
\end{aligned}
$$

For each ROI, we estimated an overall intercept $\beta_0$ and three random intercepts (individual, scanner and segmentation) and an error term with `lmer()` from `lme4` [@bates_fitting_2015].
Then we calculated the ICC as:

$$
ICC = \frac{\sigma^2_{id}}{\sigma^2_{id} + \sigma^2_{scanner} + \sigma^2_{segmentation} + \sigma^2_{\varepsilon}}
$$

We derived confidence intervals by 1000 parametric bootstraps using `bootMer()` on the fitted model.
For the longitudinal dataset we followed the same procedure but only included the random effect of segmentation as the scanner was identical.

### Percent differences

We calculated the percent difference (PD) to quantify the size and direction of the difference between batches for each ROI j across n participants according to:

$$
PD_g = 100 \cdot \frac{2}{n} \sum_{i=1}^{n} \frac{V_{ig} - S_{ig}}{|V_{ig}| + |S_{ig}|}
$$

where $V_{ig}$ and $S_{ig}$ are the volumes of ROI *g* of participant *i* in two different batches.
The absolute differences are scaled per individual so that the percent difference is a measure of mismatch relative to each participant's volume and not the sample's average volume.
Absolute value bars are necessary to avoid that changes of different signs cancel each other out.

### Significance tests

To test for significant differences, we calculated repeated-measures ANOVAs for the cross-sectional dataset and two-tailed paired t-tests for the longitudinal data.
The ANOVAs were followed up with two-tailed paired t-tests with Hommel correction for multiple testing.

### Intraclass effect decomposition

Brandmaier et al. 2018 have introduced intraclass effect decomposition (ICED).
ICED is based on Generalizability Theory [@cronbach_dependability_1972] and uses structural equation modelling (SEM) to discern individual sources of variation [@brandmaier_assessing_2018].
This approach allowed us to identify the contributions of individual differences, scanners, segmentations and residual error to the variance in the cross-sectional dataset.
We specified a SEM with eight observed variables corresponding to each unique scanner-segmentation combination.
We assumed that they all measured one common latent variable representing the true effect of interest.
In addition, we added latent variables representing common method variance of those observed variables that came from the same scanner and those that used the same segmentation approach.
In addition, we assumed homoscedastic residual error variances across all observations.
All variances were orthogonal to one another.
We constrained all factor loadings to 1 and estimated a mean for the *truth*, the second scanner and the second to fourth segmentation latent variables.
Figure \ref{fig-model} provides a path diagram of the model.\
We calculated the ICED-ICC for the combination of scanner $j$ with segmentation $k$ as:

$$
ICC_{jk} = \frac{\sigma^2_{truth}}{\sigma^2_{truth} + \sigma^2_{scanner_j} + \sigma^2_{segmentation_k} + \sigma^2_{\varepsilon}}
$$

If a variance estimate was negative with a standard error larger than its distance to 0, we set it to 0 for the ICC calculation.

```{=latex}
\begin{figure}[htbp]
\centering

\begin{tikzpicture}
    [manifest/.style={rectangle, draw=black!80, semithick, minimum width=1cm,
                      inner sep=4pt, text centered},
     latent/.style={circle, draw=black!80, thick,
                    minimum size=1.6cm,
                    inner sep=0pt,
                    text centered},
     error/.style={circle, draw=black!80, thick,   
                   minimum size=0.725cm,
                   inner sep=0pt,
                   text centered},
     from/.style={<-, >=stealth', semithick},
     text height=1.75ex, text depth=.5ex]

\begin{scope}

% ── Manifest nodes ────────────────────────────────────────────────────────
\node[manifest] (m1) {$m_1$};
\node[manifest, node distance=.6] (m2) [right=of m1] {$m_2$};
\node[manifest, node distance=.6] (m3) [right=of m2] {$m_3$};
\node[manifest, node distance=.6] (m4) [right=of m3] {$m_4$};
\node[manifest, node distance=.6] (m5) [right=of m4] {$m_5$};
\node[manifest, node distance=.6] (m6) [right=of m5] {$m_6$};
\node[manifest, node distance=.6] (m7) [right=of m6] {$m_7$};
\node[manifest, node distance=.6] (m8) [right=of m7] {$m_8$};

% ── Latent nodes ──────────────────────────────────────────────────────────
\coordinate (mid) at ($(m4)!0.5!(m5)$);
\node[latent] (truth) at ($(mid) + (0, 4)$) {$\mathrm{truth}$};

\node[latent] (SKYRA) at ($(m2) + (0, 2.5)$) {$\mathrm{SKYRA}$};

\node[latent] (VERIO) at ($(m7) + (0, 2.5)$) {$\mathrm{VERIO}$};

\coordinate (samsegmid) at ($(m1)!0.5!(m2)$);
\node[latent] (samseg) at ($(samsegmid) + (0, -3)$) {$\mathrm{samseg}$};

\coordinate (recon6mid) at ($(m3)!0.5!(m4)$);
\node[latent] (recon6) at ($(recon6mid) + (0, -3)$) {$\mathrm{recon6}$};

\coordinate (recon7mid) at ($(m5)!0.5!(m6)$);
\node[latent] (recon7) at ($(recon7mid) + (0, -3)$) {$\mathrm{recon7}$};

\coordinate (catmid) at ($(m7)!0.5!(m8)$);
\node[latent] (cat) at ($(catmid) + (0, -3)$) {$\mathrm{cat}$};

% ── residual errors ─────────────────────────────────────────
\node[error] (error1) at ($(m1) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error2) at ($(m2) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error3) at ($(m3) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error4) at ($(m4) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error5) at ($(m5) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error6) at ($(m6) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error7) at ($(m7) + (0, -1.75)$) {$\epsilon_1$};
\node[error] (error8) at ($(m8) + (0, -1.75)$) {$\epsilon_1$};

% ── truth → m1:m8 (fan) ─────────────────────────────────────────
\foreach \m in {m1, m2, m3, m4, m5, m6, m7, m8} {
     \draw[->, >=stealth', semithick] (truth) -- (\m);
}

% ── SKYRA → m1, m3, m5, m7  (fan) ────────────────────────────────────────
\foreach \m in {m1, m3, m5, m7} {
    \draw[->, >=stealth', semithick] (SKYRA) -- (\m);
}

% ── VERIO → m2, m4, m6, m8  (fan) ────────────────────────────────────────
\foreach \m in {m2, m4, m6, m8} {
    \draw[->, >=stealth', semithick] (VERIO) -- (\m);
}

% ── samseg → m1, m2  (fan) ────────────────────────────────────────
\foreach \m in {m1,m2} {
    \draw[->, >=stealth', semithick] (samseg) -- (\m);
}

% ── recon6 → m3, m4  (fan) ────────────────────────────────────────
\foreach \m in {m3, m4} {
    \draw[->, >=stealth', semithick] (recon6) -- (\m);
}

% ── recon7 → m5, m6  (fan) ────────────────────────────────────────
\foreach \m in {m5, m6} {
    \draw[->, >=stealth', semithick] (recon7) -- (\m);
}

% ── cat → m7, m8  (fan) ────────────────────────────────────────
\foreach \m in {m7, m8} {
    \draw[->, >=stealth', semithick] (cat) -- (\m);
}

% ── error → manifest ──────────────────────────────────────────────────────
\foreach \e/\m in {error1/m1, error2/m2, error3/m3, error4/m4,
                   error5/m5, error6/m6, error7/m7, error8/m8} {
    \draw[->, >=stealth', semithick] (\e) -- (\m);
}

% ── Mean nodes ─────────────────────────────────────────────────────────────
\node[isosceles triangle,
      draw,
      isosceles triangle apex angle=60,
      rotate=90,
      minimum size=1cm] (mean1) at ($(VERIO) + (0, 2.5)$) {};

% Place text independently at the visual centre
\node at ($(mean1.center) + (0, 0.1)$) {$1$};

\node[isosceles triangle,
      draw,
      isosceles triangle apex angle=60,
      rotate=90,
      minimum size=1cm] (mean2) at ($(recon7) + (0, -3)$) {};

% Place text independently at the visual centre
\node at ($(mean2.center) + (0, 0.1)$) {$1$};


% ── mean1 → VERIO, truth ───────────────────────────────────────────────────
\draw[->, >=stealth', semithick] (mean1) -- node[pos=0.5, fill=white, inner sep=2pt] {$\mu_2$} (VERIO);
\draw[->, >=stealth', semithick] (mean1) -- node[pos=0.5, fill=white, inner sep=2pt] {$\mu_1$} (truth);

% ── mean2 → recon6, recon7, cat ───────────────────────────────────────────────────
\draw[->, >=stealth', semithick] (mean2) -- node[pos=0.5, fill=white, inner sep=2pt] {$\mu_3$} (recon6);
\draw[->, >=stealth', semithick] (mean2) -- node[pos=0.5, fill=white, inner sep=2pt] {$\mu_4$} (recon7);
\draw[->, >=stealth', semithick] (mean2) -- node[pos=0.5, fill=white, inner sep=2pt] {$\mu_5$} (cat);

% ── Latent variances (curved self-arrows) ─────────────────────────────────
% Top latents: loop curves upward
\foreach \n/\dir/\lbl in {truth/90/$\sigma^2_{tr}$, SKYRA/135/$\sigma^2_{sc1}$, VERIO/45/$\sigma^2_{sc2}$} {
    \draw[<->, >=stealth', semithick,
          looseness=3, out=\dir+25, in=\dir-25]
        (\n) to node[pos=0.5, fill=white, inner sep=0.5pt, font=\scriptsize] {\lbl} (\n);
}

% Bottom latents: loop curves leftward
\foreach \n/\dir/\lbl in {error1/180/$\sigma^2_{\epsilon}$, error3/180/$\sigma^2_{\epsilon}$,
                           error5/180/$\sigma^2_{\epsilon}$, error7/180/$\sigma^2_{\epsilon}$} {
    \draw[<->, >=stealth', semithick,
          looseness=3, out=\dir+40, in=\dir-40]
        (\n) to node[pos=0.5, fill=white, inner sep=0.5pt, font=\tiny] {\lbl} (\n);
}

% Bottom latents: loop curves rightward
\foreach \n/\dir/\lbl in {error2/0/$\sigma^2_{\epsilon}$, error4/0/$\sigma^2_{\epsilon}$,
                           error6/0/$\sigma^2_{\epsilon}$, error8/0/$\sigma^2_{\epsilon}$} {
    \draw[<->, >=stealth', semithick,
          looseness=3, out=\dir+40, in=\dir-40]
        (\n) to node[pos=0.5, fill=white, inner sep=0.5pt, font=\tiny] {\lbl} (\n);
}

% latent errors: loop curves downward
\foreach \n/\dir/\lbl in {samseg/270/$\sigma^2_{seg1}$, recon6/270/$\sigma^2_{seg2}$,
                           recon7/330/$\sigma^2_{seg3}$, cat/270/$\sigma^2_{seg4}$} {
    \draw[<->, >=stealth', semithick,
          looseness=3, out=\dir+25, in=\dir-25]
        (\n) to node[pos=0.5, fill=white, inner sep=0.5pt, font=\scriptsize] {\lbl} (\n);
}

\end{scope}
\end{tikzpicture}
\caption{Path diagram of the structural equation model for intraclass effect decomposition. All factor loadings are fixed to 1. All residual error variances are fixed to equality. The latent variables' variances are freely estimated. As SKYRA and samseg do not have a mean they are the references relative to which the other latent variables' means are estimated. $m_1$-$m_8$ are the measurements of a region of interest with each of the scanner-segmentation combinations. tr = truth; sc = scanner; seg = segmentation.}
\label{fig-model}
\end{figure}
```

## Effect Size

We preregistered that an ICC \> 0.75 would be our primary criterion for sufficient harmonization.
Moreover, we considered a percent difference \< 2% and a p-value \>= 5% to be negligible.

## Statistical Power

We conducted a priori power analyses using `power.t.student()` and `power.f.mixed.anova()` from `pwrss` [@bulus_pwrss_2025] as described in our [preregistration](https://osf.io/7dcuh/overview) (`https://osf.io/7dcuh/overview`).
The results indicated that we had \>99% power to detect small effects (*d* = 0.2 / $\eta^2$ = 0.01) for all our samples.

## Sensitivity analyses

As a sensitivity analyses, we tested whether harmonization success is sensitive to manual edits.
We excluded manually edited segmentations (cross-sectional: FreeSurfer 6.0.0p; longitudinal: FreeSurfer 5.3.0), harmonized the reduced datasets, and re-ran our analyses.

## Exploratory analyses

We conducted five exploratory analyses:

1.  To further probe the effects of harmonization in longitudinal data, we tested for additive and multiplicative batch effects before and after harmonization with longitudinal ComBat.
    To that end, we used `longCombat`'s `addTest()` and `multTest()` functions with random intercepts per participant and controlling for age and gender.

2.  We tested if ComBat is sensitive to the template generation employed in longitudinal pipelines.
    Hence, we evaluated its performance on the cross-sectional pipelines in the longitudinal data.

3.  We tested the behaviour of `neuroHarmonize` when harmonizing change scores directly rather than absolute values from which change scores are calculated subsequently.

4.  We evaluated the success of ComBat in harmonizing absolute values (rather than changes).

5.  We investigated whether we could harmonize WMHV if only scanner differences are present.

# Results {#sec-results}

## Preregistered analyses

### Cross-sectional *travelling brain* dataset

```{r update_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
# set font for figures
library(showtext)
library(ggplot2)
font_add_google("Fira Sans", "firasans")
showtext_auto()
theme_set(theme_classic(base_family = "firasans", base_size = 11))
# load and prepare results 
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/outcomes.RData")
update_lme_icc_res <- outcomes$update_lme_icc_res
update_lme_icc_res_no_wmhv <- update_lme_icc_res[update_lme_icc_res$ROI != "WMHV",]
pd_mat <- outcomes$update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- outcomes$update_anova_res
pwcs <- outcomes$update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc")]]
  post <- pwcs[[paste0(outcome, "_pwc_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p <- ifelse(anovas$p < 0.001, formatC(anovas$p, format = "e", digits = 2), anovas$p)
anovas$p_adj <- ifelse(anovas$p_adj < 0.001, formatC(anovas$p_adj, format = "e", digits = 2), anovas$p_adj)
anovas$ges <- ifelse(anovas$ges < 0.001, formatC(anovas$ges, format = "e", digits = 2), anovas$ges)
anovas$ges_adj <- ifelse(anovas$ges_adj < 0.001, formatC(anovas$ges_adj, format = "e", digits = 2), anovas$ges_adj)
iced <- outcomes$update_iced_res
iced <- iced[iced$ROI != "TIV",]
iced[, 2:ncol(iced)] <- apply(iced[, 2:ncol(iced)], 2, function(x) ifelse(abs(as.numeric(x)) < 0.001 & as.numeric(x) != 0, formatC(as.numeric(x), format = "e", digits = 0), round(as.numeric(x), digits = 3)))
```

After harmonization, all our preregistered criteria for harmonization success were met for all ROIs.
Prior to harmonization, ICCs varied widely across ROIs (range: `r round(min(update_lme_icc_res_no_wmhv$ICC), digits = 2)` -- `r round(max(update_lme_icc_res_no_wmhv$ICC), digits = 2)`).
The disagreement was particularly strong for total cortical volume, for which the neuromorphometrics atlas of CAT did not provide a composite volume, but even more pronounced for amygdala volume.
ANOVAs returned statistically significant batch differences for all volumes (all generalized $\eta^2 \geq$ `r min(anovas$ges)`).
Likewise, average percent differences in batch comparisons exceeded 2% in all ROIs.
Differences between segmentations were greater than those between scanners for all ROIs (see @tbl-between-within).\
After harmonization, ICCs improved substantially (range: `r round(min(update_lme_icc_res_no_wmhv$ICC_adj), digits = 2)` - `r round(max(update_lme_icc_res_no_wmhv$ICC_adj), digits = 2)`, see @tbl-update-icc) and there were no significant differences between the mean volumetric estimates across ROIs (all generalized $\eta^2 \leq$ `r max(anovas$ges_adj)`).
Likewise, no pairwise comparison of two batches was significant after harmonization (see @tbl-update-signif).
The highest average percent difference was `r pd_sub2[which.max(pd_sub2$mean_pd_post), "ROI"]` with `r round(pd_sub2[which.max(pd_sub2$mean_pd_post), "mean_pd_post"], digits = 2)`% (see @tbl-update-pd).\
The ICED analyses produced ICCs \> 0.75 for all scanner-segmentation combinations across all ROIs.
Neither scanner showed markedly higher reliability.
This SEM-based metric was largely unaltered after harmonization (see @tbl-iced-icc).\
ComBat did not shift variances.
However, scanner and segmentation means were all close to 0 after harmonization indicating a reduction of systematic batch differences (see @tbl-iced-variances and @tbl-iced-means).\
@fig-update to @fig-update4 illustrate the distribution of measurements in the batches before and after harmonization.

```{r}
#| label: tbl-update-icc
#| tbl-cap: "ICCs before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC, ICC_lower_95_CI, ICC_upper_95_CI),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_adj, ICC_lower_95_CI_adj, ICC_upper_95_CI_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricles volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p", "ges", "n_sig_pre", "p_adj", "ges_adj", "n_sig_post")] |>
  kbl(
    escape = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r", "r", "r")
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 3,
    "harmonized" = 3
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
  general = paste0(
    "ROI = region of interest; p-value = p-value from ANOVA; ",
    "$\\\\eta^2$", " = generalized ", "$\\\\eta^2$", " from ANOVA;",
    "harm. = harmonized with neuroHarmonize; ",
    "sig. PWCs = share of significant pairwise comparisons; ",
    "TIV = total intracranial volume; TGMV = total grey matter volume; ",
    "TWMV = total white matter volume; TCV = total cortical volume; ",
    "LVV = bilateral lateral ventricle volume; ",
    "HCV = bilateral hippocampal volume; ",
    "AV = bilateral amygdala volume"
  ),
  general_title  = "",
  threeparttable = TRUE,
  escape = FALSE
  )
```

```{r}
#| label: tbl-update-pd
#| tbl-cap: "Percent differences before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percent difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

\FloatBarrier
\newpage

::: landscape
```{r}
#| label: fig-update
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline TIV = Total Intracranial Volume; TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
library(tidyverse)
library(MetBrewer)
library(patchwork)
library(showtext)
showtext_auto()
font_add_google("EB Garamond")
colors = met.brewer(name="Java", n=5, type="discrete")
colors <- colors[c(1,2,4,5)]
# load the unadjusted data from the scanner update study 
df_update <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/df_update.csv")
# break each scanner-segmentation combination down to a scanner and a segmentation variable
df_update$scanner <- ifelse(grepl("VER", df_update$SITE), "VERIO", "SKYRA")
df_update <- df_update %>%
  mutate(segmentation = case_when(
    grepl("samseg", SITE) ~ "samseg",
    grepl("cat", SITE) ~ "cat",
    grepl("recon6", SITE) ~ "recon6",
    grepl("recon7", SITE) ~ "recon7"
  ))
# for unscaled age data
QA <- read.csv("/data/pt_nro186_lifeupdate/Data/FREESURFER/QA/QA_list_short.csv")
# remove empty rows
# remove empty rows
QA <- QA[1:121,]
# create id col, add 0 that might got lost at the beginning/end, remove ., add "sub-"
# first add 0 lost at the end
QA$ID <- ifelse(nchar(sub(".*\\.", "", QA$ID)) == 1, paste0(QA$ID, "0"), QA$ID)
QA$id <- ifelse(nchar(QA$ID) != 7, gsub("\\.", "", paste0("sub-" ,QA$ID)), gsub("\\.", "", paste0("sub-", "0", QA$ID)))

# trim trailing spaces
QA$id <- trimws(QA$id)
QA$correction.necessary <- ifelse(is.na(QA$correction.necessary), 0, QA$correction.necessary)
# rename German colnames
QA <- QA %>%
  rename(
    "age" = "Alter_aktuell",
    "gender" = "Geschlecht"
  )
df_update <- df_update %>%
  select(id, TIV, TGMV, TWMV, TCV, LVV, HCV, AV, SITE) %>%
  left_join(QA[,c("id", "correction.necessary", "age", "gender")], by = "id")

# load the harmonized data from the scanner update study 
df_update_adj <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/df_update_adj.csv", col.names = colnames(df_update)[2:8])
# fill in the missing data from the unadjusted df (these cols must not be included during harmonization) 
df_update_adj[,c("id", "SITE", "age", "gender")] <- df_update[,c("id", "SITE", "age", "gender")]
df_update$SITE <- gsub("_", " ", df_update$SITE)
df_update_adj$SITE <- gsub("_", " ", df_update_adj$SITE)
df_update$SITE <- factor(df_update$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_update_adj$SITE <- factor(df_update_adj$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_update_wmhv <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/wmhv_update.csv")
df_update_wmhv_adj <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/wmhv_update_adj.csv")
df_update_wmhv$SITE <- ifelse(df_update_wmhv$scanner == "skyra", "SKYRA samseg", "VERIO samseg")
df_update_wmhv$SITE <- factor(df_update_wmhv$SITE, levels = c("SKYRA samseg", "VERIO samseg"))
df_update_wmhv_adj[,c("id", "SITE")] <- df_update_wmhv[,c("id", "SITE")]
df_update_wmhv$WMHV <- df_update_wmhv$WMHV_norm
df_update_wmhv_adj$WMHV <- df_update_wmhv_adj$X..WMHV_norm  
plots <- list()
comparisons <- list()
dfs <- list(df_update = df_update, df_update_adj = df_update_adj)
wmhv_dfs <- list(df_update_wmhv = df_update_wmhv, df_update_wmhv_adj = df_update_wmhv_adj)
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV", "HCV", "AV")){
  for (name in names(dfs)){
    df <- dfs[[name]]
    suffix <- ""
    y_axis_title <- outcome
    if (grepl("adj", name)){
      suffix <- "_adj"
      y_axis_title <- paste0(outcome, " harmonized")
    }
    ordered_ids <- df %>%
      group_by(id) %>%
      summarise(mean_val = mean(.data[[outcome]], na.rm = TRUE)) %>%
      arrange(mean_val) %>%
      pull(id)
    df$id <- factor(df$id, levels = ordered_ids)
    plt <- ggplot(data = df, aes(x = factor(id), y = .data[[outcome]], color = SITE, group = SITE, shape = SITE)) +
      geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
      scale_shape_manual(values = c(16, 1, 16, 1, 16, 1, 16, 1)) +
      theme_classic(base_family = "garamond") +
      guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) + 
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.0, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.5, "cm"), legend.key.spacing.y = unit(0.0, "cm"), legend.key.spacing.x = unit(0, "cm"),
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
      ylab(y_axis_title) + scale_color_manual(values = rep(colors, each = 2))
    if (grepl("adj", name)){
      plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
    }
    plots[[paste0(outcome, suffix)]] <- plt
  }
  y_range <- range(c(plots[[outcome]]$data[[outcome]], plots[[paste0(outcome, "_adj")]]$data[[outcome]]), na.rm = TRUE)
  plots[[outcome]] <- plots[[outcome]] + coord_cartesian(ylim = y_range)
  if(outcome %in% c("TGMV", "TCV", "HCV")){
    plots[[outcome]] <- plots[[outcome]] + labs(x = "individuals (sorted by mean across batches)") + theme(axis.title.x = element_text(size = 5, margin = margin(t = 6)))
  }
  plots[[paste0(outcome, "_adj")]] <- plots[[paste0(outcome, "_adj")]] + coord_cartesian(ylim = y_range)
  comparison <- wrap_plots(list(plots[[outcome]], plots[[paste0(outcome, "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
  comparisons[[outcome]] <- comparison
}
for (name in names(wmhv_dfs)){
  df <- wmhv_dfs[[name]]
  suffix <- ""
  y_axis_title <- "WMHV"
  if (grepl("adj", name)){
    suffix <- "_adj"
    y_axis_title <- paste0("WMHV harmonized")
  }
  ordered_ids <- df %>%
    group_by(id) %>%
    summarise(mean_val = mean(.data[["WMHV"]], na.rm = TRUE)) %>%
    arrange(mean_val) %>%
    pull(id)
  df$id <- factor(df$id, levels = ordered_ids)
  plt <- ggplot(data = df, aes(x = factor(id), y = .data[["WMHV"]], color = SITE, group = SITE, shape = SITE)) +
    geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
    scale_shape_manual(values = c(16, 1)) +
    theme_classic(base_family = "garamond") +
    guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.0, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.5, "cm"), legend.key.spacing.y = unit(0.0, "cm"), legend.key.spacing.x = unit(0, "cm"),
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
    ylab(y_axis_title) + scale_color_manual(values = rep(colors[3], 2))
  if (grepl("adj", name)){
    plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
  }
  plots[[paste0("WMHV", suffix)]] <- plt
}
y_range <- range(c(plots[["WMHV"]]$data[["WMHV"]], plots[[paste0("WMHV", "_adj")]]$data[["WMHV"]]), na.rm = TRUE)
plots[["WMHV"]] <- plots[["WMHV"]] + coord_cartesian(ylim = y_range) + labs(x = "individuals (sorted by mean across batches)") + theme(axis.title.x = element_text(size = 5, margin = margin(t = 6)))
plots[[paste0("WMHV", "_adj")]] <- plots[[paste0("WMHV", "_adj")]] + coord_cartesian(ylim = y_range)
comparison <- wrap_plots(list(plots[["WMHV"]], plots[[paste0("WMHV", "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
comparisons[["WMHV"]] <- comparison

legend_only_plot <- ggplot() + 
  theme_void() +
  annotation_custom(cowplot::get_legend(comparisons[[1]]))
p_combined <- wrap_plots(comparisons[[1]], comparisons[[2]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update2
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline TWMV = Total White Matter Volume; TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[3]], comparisons[[4]],
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update3
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[5]], comparisons[[6]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update4
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[7]], comparisons[[8]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

\newpage
:::

\FloatBarrier

### Longitudinal dataset

```{r life_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
life_lme_icc_res <- outcomes$life_lme_icc_res
life_lme_icc_res$ROI <- gsub("d", "\u0394", life_lme_icc_res$ROI)
pd_mat <- outcomes$life_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 6, ncol = 7))
colnames(pd_sub2) <- c("ROI", "pre", "nh", "combat", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_sub2$ROI <- c("\u0394TGMV", "\u0394TCV", "\u0394LVV",  "\u0394HCV", "\u0394AV", "\u0394WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_nh <- length(which(nh >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_combat <- length(which(combat >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "nh"] <- paste0(n_exceeding_nh, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "combat"] <- paste0(n_exceeding_combat, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pwcs <- outcomes$life_pwcs
pwcs$pwc_wmhv$measure <- "WMHV"
pwcs$pwc_wmhv_combat$measure <- "WMHV"
pwcs$pwc_wmhv_nh$measure <- "WMHV"
pwc <- rbind(pwcs$pwc[,!grepl("adj", colnames(pwcs$pwc))], pwcs$pwc_wmhv)
pwc_combat <- rbind(pwcs$pwc_combat[,!grepl("adj", colnames(pwcs$pwc_combat))], pwcs$pwc_wmhv_combat)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh[,!grepl("adj", colnames(pwcs$pwc_nh))], pwcs$pwc_wmhv_nh)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all$p <- ifelse(pwc_all$p < 0.001, formatC(pwc_all$p, format = "e", digits = 2), pwc_all$p)
pwc_all$p_nh <- ifelse(pwc_all$p_nh < 0.001, formatC(pwc_all$p_nh, format = "e", digits = 2), pwc_all$p_nh)
pwc_all$p_combat <- ifelse(pwc_all$p_combat < 0.001, formatC(pwc_all$p_combat, format = "e", digits = 2), pwc_all$p_combat)
pwc_all$measure <- gsub("d", "\u0394", pwc_all$measure)
pwc_all[pwc_all$measure == "WMHV", "measure"] <- paste0("\u0394", pwc_all[pwc_all$measure == "WMHV", "measure"])
# calculate cohens d
pwc_all$d <- pwc_all$statistic/(sqrt(pwc_all$n1))
pwc_all$d_combat <- pwc_all$statistic_combat/(sqrt(pwc_all$n1_combat))
pwc_all$d_nh <- pwc_all$statistic_nh/(sqrt(pwc_all$n1_nh))
pwc_all[,c("d", "d_nh", "d_combat")] <- abs(round(pwc_all[,c("d", "d_nh", "d_combat")], digits = 2))
```

ICCs for all ROIs except for change in HCV and LVV were below 0.75 both before and after harmonization (see @tbl-life-icc).
Likewise, percent differences remained markedly above the 2% threshold (see @tbl-life-pd) and paired t-tests comparing the batches stayed highly significant for all ROIs except amygdala volume (see @tbl-life-signif).
This was the case both for data harmonized with `longComBat` and `neuroHarmonize`.
Bland-Altman plots depicting the differences between the two batches before and after harmonization are provided in @fig-ba-dtgmv to @fig-ba-dav.

```{r }
#| label: tbl-life-icc
#| tbl-cap: "ICCs before and after harmonization in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df2 <- life_lme_icc_res %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC, ICC_lower_95_CI, ICC_upper_95_CI
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh, ICC_lower_95_CI_nh, ICC_upper_95_CI_nh
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat, ICC_lower_95_CI_combat, ICC_upper_95_CI_combat
    )
  ) 
df2[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "\u0394 = change in",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all[, c("measure", "p", "d", "p_nh", "d_nh", "p_combat", "d_combat")] |>
  kbl(
    row.names = FALSE,
    escape = FALSE,
    col.names = c(
    "ROI",
    "p-value",
    "$d_z$",
    "p-value",
    "$d_{z}$",
    "p-value",
    "$d_{z}$"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c", "c", "c", "c")
  ) |>
  add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; neuroHarm = harmonized with neuroHarmonize; ",
      "ComBat = harmonized with longitudinal ComBat; $d_z$ = Cohen's d for paired samples; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE,
    escape = FALSE
  )
```

```{r }
#| label: tbl-life-pd
#| tbl-cap: "Percent differences before and after harmonization in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2[,c("ROI", "pre", "mean_pd_pre", "nh", "mean_pd_nh", "combat", "mean_pd_combat")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "> 2%",
      "PD",
      "> 2%",
      "PD",
      "> 2%",
      "PD"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "c", "r", "c", "r", "c"),
    format = "latex"
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>

  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; > 2% = share of percent differences > 2%; ",
      "PD = percent difference;  ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

\FloatBarrier

## Sensitivity and exploratory analyses

After excluding manually edited scans, results were consistent with those of the full sample across all statistical tests (see @tbl-update-icc-no-correction to @tbl-life-signif-no-correction).\
The tests for remaining additive or multiplicative batch effects after harmonization showed that the additive batch effects were fully removed but multiplicative effects remained significant after harmonization (see @tbl-add-mult).\
Using measurements based on cross-sectional pipelines in the longitudinal dataset did not lead to substantially different results (see @tbl-icc-cross).\
Harmonizing change scores directly led to moderately higher ICCs for some variables.
Still, all ROIs except HCV and LVV remained below the threshold of 0.75 (see @tbl-icc-change).\
In contrast, if we consider the agreement between the harmonized absolute values at baseline and follow-up rather than the changes, ICCs for all volumes but WMHV are markedly higher.
All exceed 0.75 for neuroHarmonize and all but baseline AV for longitudinal ComBat (see @tbl-life-bl-icc and @tbl-life-fu-icc).
In addition, all percent differences except for WMHV are substantially reduced after harmonization with lower values for neuroHarmonize at baseline and for longitudinal ComBat at follow-up (see @tbl-life-bl-pd and @tbl-life-fu-pd).
Similarly, there are no significant differences between batches at baseline after the application of neuroHarmonize (except WMHV).
Decreased but still significant differences at follow-up remain for all ROIs after both harmonization methods (see @tbl-life-bl-signif and @tbl-life-fu-signif).
We provide the corresponding Bland-Altman plots for observed values at each timepoint in the supplement (see @fig-ba-bl-tgmv to @fig-ba-fu-av).\
The agreement between WMHV using the same segmentation method on different MRI machines in the cross-sectional data showed an ICC of `r round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC"], digits = 3)` (95% CI: `r  round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI"], digits = 3)`--`r  round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI"], digits = 3)`) before and of `r round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_adj"], digits = 3)` (95% CI: `r  round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI_adj"], digits = 3)`--`r  round(outcomes$update_lme_icc_res[outcomes$update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI_adj"], digits = 3)`) after harmonization.
Measurements on the SKYRA were on average `r round(outcomes$update_pd_matrices$WMHV_norm_mean_pd_mat[1,2], digits = 1)`% larger before harmonization but only `r abs(round(outcomes$update_pd_matrices$WMHV_adj_mean_pd_mat_adj[1,2], digits = 1))`% smaller after the application of neuroHarmonize.
A paired t-test of the batches was significant prior to harmonization (p = `r formatC(outcomes$update_pwcs$WMHV_norm_pwc$p, digits = 2, format = "e")`; t(`r outcomes$update_pwcs$WMHV_adj_pwc_adj$df`) = `r formatC(outcomes$update_pwcs$WMHV_norm_pwc$statistic, digits = 2, format = "g")`; d = `r  round(abs((outcomes$update_pwcs$WMHV_norm_pwc$statistic)/(sqrt(outcomes$update_pwcs$WMHV_norm_pwc$n1))), digits = 2)`) but ceased to be so afterwards (p = `r  outcomes$update_pwcs$WMHV_adj_pwc_adj$p`; t(`r outcomes$update_pwcs$WMHV_adj_pwc_adj$df`) = `r formatC(outcomes$update_pwcs$WMHV_adj_pwc_adj$statistic, digits = 2, format = "g")`; d = `r round(abs((outcomes$update_pwcs$WMHV_adj_pwc_adj$statistic)/(sqrt(outcomes$update_pwcs$WMHV_adj_pwc_adj$n1))), digits = 2)`).

\newpage

# Discussion

Integrating heterogenous neuroimaging data is increasingly necessary to push the frontiers of neuroscience [@zhu_statistical_2023].
This in turn requires strategies to account for batch effects arising from heterogeneity in scanners and software.
Therefore, we evaluated the performance of ComBat methods in removing such effects across various ROIs in cross-sectional and longitudinal designs.\
Our analyses provided fully satisfactory results for all ROIs in cross-sectional analyses.
In contrast, ComBat did not achieve sufficient agreement of change scores.
This was not an artifact of template-generation in longitudinal pipelines or manual editing of scans.
Furthermore, the direct harmonization of change scores was no solution to this problem.
We attribute this result to two factors.\
Firstly, ComBat has been more effective in removing additive than multiplicative batch effects (@tbl-add-mult).
Additive effects affect measurements in the same batch uniformly and thus cancel out when computing change scores.
Hence, batch differences in change scores solely result from the more challenging multiplicative effects.\
Secondly, even over a follow-up period of six years, the individual variance in volumetric changes is small compared to the measurement error for many ROIs.
The ICED framework helps us to elucidate this constraint on harmonizability.
The results show that systematic differences between batches can be removed using ComBat (@tbl-iced-means).
Unsystematic measurement error is by nature unamenable to harmonization, though (@tbl-iced-variances).
ICED-ICCs express how much of the total variability (within a scanner-segmentation combination) is explained by true variance and were thus unaffected by harmonization.
Hence, even in the absence of batch effects, high reliability is only attainable when true variance sufficiently exceeds error variance.
Therefore, acceptable standard ICCs were observed only for HCV and LVV, two structures that undergo substantial age-related atrophy or enlargement.
Analogously, longitudinal harmonizations were much more successful when evaluated on the scale of observed variables.
Only then can we successfully differentiate the signal of true variance from noise.
We have illustrated this by adding the extent of one standard deviation of the measured variable as a scale to Bland-Altman plots (@fig-ba-dtgmv to @fig-ba-fu-av).\
WMHV consistently exhibited the poorest agreement.
Even in the absence of segmentation software differences the ICC only narrowly exceeded our threshold after harmonization.
As differences between batches were small and statistically non-significant after harmonization, this must be attributed to the substantial measurement error visible in @fig-update4.
The plot also shows that this is particularly pronounced in smaller WMHV values while agreement improves as volumes increase.

Our results confirm previous positive evaluations of ComBat methods on cross-sectional data [@tassi_assessment_2024; @sun_comparison_2022; @eshaghzadeh_torbati_multi-scanner_2021; @fortin_harmonization_2018; @pomponio_harmonization_2020; @yu_statistical_2018].
They provide particularly solid evidence by leveraging a large *travelling brain* dataset to establish a ground truth scenario.
Importantly, this high reliability could be achieved without modelling an overlap sample between batches [@dewey_deepharmony_2019; @maikusa_comparison_2021; @yamashita_harmonization_2019; @tian_deep_2022].
Moreover, it was attained in spite of variable post-processing in the form of manual editing.\
Previous cross-sectional evaluations with a *travelling brain* design and uniform segmentation found high agreement in unharmonized structural measures and no benefit in harmonization [@gebre_crossscanner_2023; @richter_validation_2022].
In contrast, our results showed improved agreement in ROIs with both high and low agreement prior to harmonization.\
However, our results also indicate low reliability when pooling data for the study of within-participant changes.
Studies of the harmonizability of change scores from different scanners with the same segmentation have found similar problems [@gebre_crossscanner_2023; @richter_validation_2022].
Our analyses illustrate that this is partly due to problems with correcting multiplicative batch effects.
The low agreement is primarily attributable to measurement error, which cannot be remedied by harmonization, though.

## Limitations

We purposively chose statistical methods to make our results generalizable to other scanners and segmentations.
However, we cannot assume with certainty that the results extend to all scanners, segmentations and their combinations.
The two scanners in this study were 3T scanners from the same manufacturer with highly similar pulse sequences.
Between-scanner differences might be more pronounced and more intricate to adjust for when scans are more dissimilar.\
The high WMHV measurement error may be less pronounced in supervised algorithms like BIANCA [@torres-simon_decoding_2024].
However, samseg has been found to have high reliability when compared to manual segmentations on the volumetric level and to be more robust in handling multi-site data than BIANCA [@ferris_optimizing_2023].\
Furthermore, it should be noted that ComBat will conflate actual batch effects with sample differences if these are not specified as covariates.
This could lead to unintended consequences in subsequent analyses.
Therefore, results based on data harmonised with ComBat should always be interpreted relative to the adjustment set.\
Moreover, ComBat does not address all batch differences of potential relevance.
For example, there might be interactions between sample differences like age composition and measurement quality [@radua_increased_2020].\
ComBat can be used to harmonise data and thus allows researchers to conduct an IDA with higher power than a meta-analysis would have [@radua_increased_2020].
However, this depends on the availability of individual participant data.
If aggregate data can be obtained for multiple additional studies, a meta-analysis might be more suitable aside from being much less laborious [@cooper_relative_2009].\
Lastly, it should be noted that ComBat is used to adjust for differences between scanners.
However, drift within scanners over time can also affect results [@takao_effect_2011] and is not corrected for by the methods presented here.

## Conclusions

From our preregistered analyses, we derive the following conclusions:

1.  ComBat can be used to harmonize cross-sectional datasets with compounded scanner and segmentation effects.

2.  ComBat does not successfully harmonize longitudinal data to allow the study of individual changes.

Our results highlight the importance of the size of true variance relative to measurement error in integrative data analyses.
IDAs address random sampling error but can not mitigate measurement error.
Importantly, large measurement errors substantially shrink effect estimates [@brandmaier_rethinking_2026].
Hence, IDA researchers should bear measurement error in mind when deciding on their methodology.\
Improvements in measurement reliability through increased field strengths [@calabro_leveraging_2025] hold promise.
Alternatively, cluster scanning [@elliott_precision_2026] at regular field strengths can mitigate measurement error by providing multiple indicators per timepoint [@von_oertzen_effect_2010].
Quantitative multi-parameter mapping offers another solution to obtain parameter measures that are comparable across sites by providing absolute values of the longitudinal relaxation time T1 and longitudinal relaxation rate [@weiskopf_quantitative_2013; @weiskopf_unified_2011].
However, none of these solutions is yet available on a large scale.
We would thus recommend the use of models that allow modelling observed values at each timepoint as dependent variables in longitudinal settings.
Researchers could use linear mixed effects models to distinguish within- and between-participant effects of age [@van_de_pol_simple_2009].
If at all used in integrative data analyses, change scores should be restricted to datasets with extensive follow-up periods and for ROIs with substantial change.
Alternatively, harmonization involving *travelling brains* might be an option [@maikusa_comparison_2021] but require further evaluation in longitudinal data.\
In addition, the issue of measurement error is of particular significance for WMHV.
As this puts a consequential upper limit on the attainable reliability of measurements, we recommend only pooling data segmented with the same software to minimise batch effects.

*In nuce*, our results highlight the relative size of measurement error as a major determinant of the success of ComBat harmonization.
This translates to satisfactory results for volumetric measurements of the studied brain structures but not for intra-individual change in these measurements.

# Code and data availability

All code is published on [github](https://github.com/formal-methods-mpi/SegmentationHarmonizability) (`https://github.com/formal-methods-mpi/SegmentationHarmonizability`).
The cross-sectional data is published on [OSF](https://osf.io/n3eg8/overview) (`https://osf.io/n3eg8/overview`).
If you execute [this script](https://github.com/formal-methods-mpi/SegmentationHarmonizability/blob/main/scripts/reproduction.qmd) (https://github.com/formal-methods-mpi/SegmentationHarmonizability/blob/main/scripts/reproduction.qmd), you can reproduce all results based on this dataset.
We were granted access to the longitudinal data by LIFE (Leipziger Forschungszentrum für Zivilisationserkrankungen) under project agreement PV-894.
All data will exclusively be shared by [LIFE](https://www.uniklinikum-leipzig.de/einrichtungen/life) (`https://www.uniklinikum-leipzig.de/einrichtungen/life`) based on a project proposal under data protection rules according to the University of Leipzig.
Thus, we can not share this dataset.

# Funding

This research was supported by LIFE Leipzig Research Center for Civilization Diseases, Leipzig University (LIFE is funded by the EU, the European Social Fund, the European Regional Development Fund, and Free State Saxony’s excellence initiative; project numbers: 713-241202, 14505/2470, 14575/2470).

# Competing interests

The authors declare no competing interests.

# References {#sec-references}

::: {#refs}
:::

\newpage
\appendix

<!-- Switch to S-prefixed numbering -->

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\setcounter{figure}{0}

# Supplement {#sec-supplement}

## Supplementary tables

```{r scanner_software_comp}
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp <- data.frame(matrix(nrow = 7, ncol = 3))
colnames(scanner_software_comp) <- c("ROI", "within_scanner_diff", "within_software_diff")
scanner_software_comp$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
# start with scanner update
pd <- outcomes$update_pd_matrices
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  mat <- pd[[paste0(outcome,"_mean_pd_mat")]]
  transf_mat <- abs(mat) %>%
    as.data.frame() %>%
    mutate(row_name = rownames(.)) %>%
    pivot_longer(-row_name, names_to = "col_name", values_to = "value") %>%
    filter(row_name != col_name) %>%
    mutate(pair_id = map2_chr(row_name, col_name, ~ paste(sort(c(.x, .y)), collapse = "_"))) %>%
    distinct(pair_id, .keep_all = TRUE) %>%
    separate(row_name, into = c("metric1","scanner1","suffix1"), sep = "_") %>%
    separate(col_name, into = c("metric2","scanner2","suffix2"), sep = "_") 
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_scanner_diff"] <- transf_mat %>%
    filter(scanner1 == scanner2) %>%
    filter(suffix1 != suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_software_diff"] <- transf_mat %>%
    filter(scanner1 != scanner2) %>%
    filter(suffix1 == suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
}
scanner_software_comp[,2:3] <- round(scanner_software_comp[,2:3], digits = 2)
```

```{r}
#| label: tbl-between-within
#| tbl-cap: "Average percent differences between softwares and between scanners in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "between software", "between scanners"),
    booktabs  = TRUE,
    linesep   = "",
    format    = "latex",
    align     = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; between software = average percent difference comparing two segmentations with different software of the same image;",
      "between scanners = average percent difference comparing two segmentations with identical software of images from different scanners"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

::: landscape
```{r}
#| label: tbl-iced-icc
#| tbl-cap: "ICED ICCs before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- lapply(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], as.numeric)
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- round(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], digits = 2)
iced[, c("ROI", paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "              = 1,
    "SKYRA cat"      = 2,
    "VERIO cat"      = 2,
    "SKYRA recon6"   = 2,
    "VERIO recon6"   = 2,
    "SKYRA recon7"   = 2,
    "VERIO recon7"   = 2,
    "SKYRA samseg"   = 2,
    "VERIO samseg"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; harm. = harmonized ICC; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-variances
#| tbl-cap: "ICED variances before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truevar", "SKYRAvar", "VERIOvar", "catvar", "recon6var", "recon7var", "samsegvar", "obsvar"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "SKYRA"    = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2,
    "samseg"   = 2,
    "residual"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; var = latent variable variance; harm. = harmonized latent variable variance; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-means
#| tbl-cap: "ICED means before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truemean", "VERIOmean", "catmean", "recon6mean", "recon7mean"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 10))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9), border_right = TRUE) |>
  footnote(
    general = paste0(
      "true = true score with SKYRA and samseg as reference scanner and segmentation, respectively; ",
      "ROI = region of interest; mean = latent variable mean; harm. = harmonized latent variable mean; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```
:::

```{r update_prep_no_correction}
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_mat <- outcomes$update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- outcomes$update_anova_res
pwcs <- outcomes$update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc_no_correction")]]
  post <- pwcs[[paste0(outcome, "_pwc_no_correction_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p_no_correction <- ifelse(anovas$p_no_correction < 0.001, formatC(anovas$p_no_correction, format = "e", digits = 2), anovas$p_no_correction)
anovas$p_no_correction_adj <- ifelse(anovas$p_no_correction_adj < 0.001, formatC(anovas$p_no_correction_adj, format = "e", digits = 2), anovas$p_no_correction_adj)
```

```{r}
#| label: tbl-update-icc-no-correction
#| tbl-cap: "ICCs before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction, ICC_lower_95_CI_no_correction, ICC_upper_95_CI_no_correction),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction_adj, ICC_lower_95_CI_no_correction_adj, ICC_upper_95_CI_no_correction_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-pd-no-correction
#| tbl-cap: "Percent differences before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percentage difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif-no-correction
#| tbl-cap: "Statistical significance of differences before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p_no_correction", "ges_no_correction", "n_sig_pre", "p_no_correction_adj", "ges_no_correction_adj", "n_sig_post")] |>
  kbl(
    escape = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r", "r", "r")
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 3,
    "harmonized" = 3
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
  general = paste0(
    "ROI = region of interest; p-value = p-value from ANOVA; ",
    "$\\\\eta^2$", " = generalized ", "$\\\\eta^2$", " from ANOVA;",
    "harm. = harmonized with neuroHarmonize; ",
    "sig. PWCs = share of significant pairwise comparisons; ",
    "TIV = total intracranial volume; TGMV = total grey matter volume; ",
    "TWMV = total white matter volume; TCV = total cortical volume; ",
    "LVV = bilateral lateral ventricle volume; ",
    "HCV = bilateral hippocampal volume; ",
    "AV = bilateral amygdala volume"
  ),
  general_title  = "",
  threeparttable = TRUE,
  escape = FALSE
  )
```

```{r life_prep_no_correction}
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_mat <- outcomes$life_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 5, ncol = 7))
colnames(pd_sub2) <- c("ROI", "pre", "nh", "combat", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_sub2$ROI <- c("\u0394TGMV", "\u0394TCV", "\u0394LVV",  "\u0394HCV", "\u0394AV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_no_correction")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_no_correction")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_nh <- length(which(nh >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_combat <- length(which(combat >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "nh"] <- paste0(n_exceeding_nh, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "combat"] <- paste0(n_exceeding_combat, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_sub2[pd_sub2$ROI == paste0("\u0394", outcome), "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pwcs <- outcomes$life_pwcs
pwc <- pwcs$pwc_no_correction[,!grepl("adj", colnames(pwcs$pwc_no_correction))]
pwc_combat <- pwcs$pwc_combat_no_correction[,!grepl("adj", colnames(pwcs$pwc_combat_no_correction))]
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- pwcs$pwc_nh[,!grepl("adj", colnames(pwcs$pwc_nh_no_correction))]
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all$p <- ifelse(pwc_all$p < 0.001, formatC(pwc_all$p, format = "e", digits = 2), pwc_all$p)
pwc_all$p_nh <- ifelse(pwc_all$p_nh < 0.001, formatC(pwc_all$p_nh, format = "e", digits = 2), pwc_all$p_nh)
pwc_all$p_combat <- ifelse(pwc_all$p_combat < 0.001, formatC(pwc_all$p_combat, format = "e", digits = 2), pwc_all$p_combat)
pwc_all$measure <- gsub("d", "\u0394", pwc_all$measure)
# calculate cohens d
pwc_all$d <- pwc_all$statistic/(sqrt(pwc_all$n1))
pwc_all$d_combat <- pwc_all$statistic_combat/(sqrt(pwc_all$n1_combat))
pwc_all$d_nh <- pwc_all$statistic_nh/(sqrt(pwc_all$n1_nh))
pwc_all[,c("d", "d_nh", "d_combat")] <- abs(round(pwc_all[,c("d", "d_nh", "d_combat")], digits = 2))
```

```{r }
#| label: tbl-life-icc-no-correction
#| tbl-cap: "ICCs before and after harmonization without manually edited scans - longitudinal dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
life_lme_icc_res_no_wmhv <- life_lme_icc_res[!grepl("WMHV", life_lme_icc_res$ROI), ]
df2 <- life_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_no_correction, ICC_lower_95_CI_no_correction, ICC_upper_95_CI_no_correction
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_no_correction, ICC_lower_95_CI_nh_no_correction, ICC_upper_95_CI_nh_no_correction
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_no_correction, ICC_lower_95_CI_combat_no_correction, ICC_upper_95_CI_combat_no_correction
    )
  ) 
df2 <- df2[df2$ROI != "WMHV",]
df2[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "c", "c", "c"),
    format = "latex"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TGMV = total grey matter volume; ",
      "\u0394 = change in",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-pd-no-correction
#| tbl-cap: "Percent differences before and after harmonization without manually edited scans - longitudinal dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2[,c("ROI", "pre", "mean_pd_pre", "nh", "mean_pd_nh", "combat", "mean_pd_combat")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "> 2%",
      "PD",
      "> 2%",
      "PD",
      "> 2%",
      "PD"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "c", "r", "c", "r", "c"),
    format = "latex"
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>

  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; > 2% = share of percent differences > 2%; ",
      "PD = percent difference;  ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-signif-no-correction
#| tbl-cap: "Statistical significance of differences before and after harmonization without manually edited scans - longitudinal dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all[, c("measure", "p", "d", "p_nh", "d_nh", "p_combat", "d_combat")] |>
  kbl(
    row.names = FALSE,
    escape = FALSE,
    col.names = c(
    "ROI",
    "p-value",
    "$d_z$",
    "p-value",
    "$d_{z}$",
    "p-value",
    "$d_{z}$"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c", "c", "c", "c")
  ) |>
  add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; neuroHarm = harmonized with neuroHarmonize; ",
      "ComBat = harmonized with longitudinal ComBat; $d_z$ = Cohen's d for paired samples; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE,
    escape = FALSE
  )
```

```{r }
#| label: tbl-add-mult
#| tbl-cap: "Statistical significance of additive and multiplicative batch effects before and after harmonization in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/longcombat_tests.RData")
longCombat_results <- data.frame(matrix(ncol = 5, nrow =6))
colnames(longCombat_results) <- c("ROI", "add_pre", "add_post", "mult_pre", "mult_post")
longCombat_results$ROI <- c(longcombat_tests$addTestTable$Feature, "WMHV")
for (roi in longCombat_results$ROI){
  if (roi != "WMHV") {
    longCombat_results[longCombat_results$ROI == roi, "add_pre"] <- longcombat_tests$addTestTable[longcombat_tests$addTestTable$Feature == roi, "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "add_post"] <- longcombat_tests$addTestTable_combat[longcombat_tests$addTestTable_combat$Feature == paste0(roi, ".combat"), "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_pre"] <- longcombat_tests$multTestTable[longcombat_tests$multTestTable$Feature == roi, "p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_post"] <- longcombat_tests$multTestTable_combat[longcombat_tests$multTestTable_combat$Feature == paste0(roi, ".combat"), "p-value"]
  } else {
    longCombat_results[longCombat_results$ROI == roi, "add_pre"] <- longcombat_tests$addTestTable_wmhv[longcombat_tests$addTestTable_wmhv$Feature == "WMHV_norm", "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "add_post"] <- longcombat_tests$addTestTable_wmhv_combat[longcombat_tests$addTestTable_wmhv_combat$Feature == "WMHV_norm.combat", "KR p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_pre"] <- longcombat_tests$multTestTable_wmhv[longcombat_tests$multTestTable_wmhv$Feature == "WMHV_norm", "p-value"]
    longCombat_results[longCombat_results$ROI == roi, "mult_post"] <- longcombat_tests$multTestTable_wmhv_combat[longcombat_tests$multTestTable_wmhv_combat$Feature == "WMHV_norm.combat", "p-value"]
  }
}
longCombat_results$add_pre <- ifelse(as.numeric(longCombat_results$add_pre) < 0.001, formatC(as.numeric(longCombat_results$add_pre), format = "e", digits = 2), round(as.numeric(longCombat_results$add_pre), digits = 3))
longCombat_results$add_post <- ifelse(as.numeric(longCombat_results$add_post) < 0.001, formatC(as.numeric(longCombat_results$add_post), format = "e", digits = 2), round(as.numeric(longCombat_results$add_post), digits = 3))
longCombat_results$mult_pre <- ifelse(as.numeric(longCombat_results$mult_pre) < 0.001, formatC(as.numeric(longCombat_results$mult_pre), format = "e", digits = 2), round(as.numeric(longCombat_results$mult_pre), digits = 3))
longCombat_results$mult_post <- ifelse(as.numeric(longCombat_results$mult_post) < 0.001, formatC(as.numeric(longCombat_results$mult_post), format = "e", digits = 2), round(as.numeric(longCombat_results$mult_post), digits = 3))
longCombat_results %>%
  kbl(
    col.names = c("ROI", "p add.", "p add. harm.", "p mult.", "p mult. harm."),
    align = c("l", "r", "r", "r", "r"),
    format = "latex",
    booktabs = TRUE,
    linesep  = ""
  ) %>%
  kable_styling(
    latex_options = c("hold_position", "scale_down")
  ) %>%
  footnote(
    general = "ROI = region of interest; p add = p-value for additive effects; harm. = harmonized with longitudinal ComBat; p mult. = p-value for multiplicative effects; TGMV = total grey matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r life_cross_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$life_lme_icc_res
df$ROI <- gsub("d", "\u0394", df$ROI)
df2 <- df %>%
  mutate(
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_cross_nh, ICC_lower_95_CI_cross_nh, ICC_upper_95_CI_cross_nh
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_cross_combat, ICC_lower_95_CI_cross_combat, ICC_upper_95_CI_cross_combat
    )
  ) 
```

```{r }
#| label: tbl-icc-cross
#| tbl-cap: "ICCs before and after harmonization of results of cross-sectional pipelines in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df2[1:5, c("ROI", "ICC_nh_CI", "ICC_combat_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI) ~ neuroHarm",
      "ICC (95% CI) ~ ComBat"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; \u0394 = change in; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r change_life_icc_table}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$change_life_lme_icc_res
colnames(df) <- paste0(colnames(df), "_change")
df2 <- cbind(df, outcomes$life_lme_icc_res[,c("ICC", "ICC_lower_95_CI", "ICC_upper_95_CI")])
df3 <- df2 %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC, ICC_lower_95_CI, ICC_upper_95_CI
    ),
    ICC_change_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_change, ICC_lower_95_CI_change, ICC_upper_95_CI_change
    )
  ) 
df3$ROI[nrow(df3)] <- "WMHV"
df3$ROI <- paste0("\u0394", df3$ROI)
```

```{r }
#| label: tbl-icc-change
#| tbl-cap: "ICCs before and after harmonization of directly harmonized change scores in longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df3[, c("ROI", "ICC_CI", "ICC_change_CI")] |>
  kbl(
    row.names = FALSE,
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) ~ neuroHarm"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ",
      "~ ComBat = harmonized with longitudinal ComBat; \u0394 = change in; ",
      "TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r tp_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- outcomes$tp_life_lme_icc_res
df$ROI[nrow(df)] <- "WMHV"
df_bl <- df %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_bl, ICC_lower_95_CI_bl, ICC_upper_95_CI_bl
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_bl, ICC_lower_95_CI_nh_bl, ICC_upper_95_CI_nh_bl
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_bl, ICC_lower_95_CI_combat_bl, ICC_upper_95_CI_combat_bl
    )
  ) 
df_fu <- df %>%
  mutate(
    ICC_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_fu, ICC_lower_95_CI_fu, ICC_upper_95_CI_fu
    ),
    ICC_nh_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_nh_fu, ICC_lower_95_CI_nh_fu, ICC_upper_95_CI_nh_fu
    ),
    ICC_combat_CI = sprintf(
      "%.2f (%.2f, %.2f)",
      ICC_combat_fu, ICC_lower_95_CI_combat_fu, ICC_upper_95_CI_combat_fu
    )
  ) 

pd_mat <- outcomes$tp_life_pd_matrices

pd_bl <- data.frame(matrix(nrow = 6, ncol = 4))
colnames(pd_bl) <- c("ROI", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_bl$ROI <- c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_bl")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_bl")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_bl")]]
  pd_bl[pd_bl$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_bl[pd_bl$ROI == outcome, "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_bl[pd_bl$ROI == outcome, "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pd_bl[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")] <- round(pd_bl[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")], digits = 3)
pd_fu <- data.frame(matrix(nrow = 6, ncol = 4))
colnames(pd_fu) <- c("ROI", "mean_pd_pre", "mean_pd_nh", "mean_pd_combat")
pd_fu$ROI <- c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")
for (outcome in c("TGMV", "TCV", "LVV",  "HCV", "AV", "WMHV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_fu")]]
  nh <- pd_mat[[paste0(outcome, "_mean_pd_mat_nh_fu")]]
  combat <- pd_mat[[paste0(outcome, "_mean_pd_mat_combat_fu")]]
  pd_fu[pd_fu$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_fu[pd_fu$ROI == outcome, "mean_pd_nh"] <- mean(abs(nh), na.rm = T)
  pd_fu[pd_fu$ROI == outcome, "mean_pd_combat"] <- mean(abs(combat), na.rm = T)
}
pd_fu[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")] <- round(pd_fu[,c("mean_pd_pre", "mean_pd_nh", "mean_pd_combat")], digits = 3)
pwcs <- outcomes$tp_life_pwcs
pwcs$pwc_wmhv_bl$measure <- "WMHV"
pwcs$pwc_wmhv_combat_bl$measure <- "WMHV"
pwcs$pwc_wmhv_nh_bl$measure <- "WMHV"
# ignore "adjusted" cols (there was no adjustment)
pwc <- rbind(pwcs$pwc_bl[,!grepl("adj", colnames(pwcs$pwc_bl))], pwcs$pwc_wmhv_bl)
pwc_combat <- rbind(pwcs$pwc_combat_bl[,!grepl("adj", colnames(pwcs$pwc_combat_bl))], pwcs$pwc_wmhv_combat_bl)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh_bl[,!grepl("adj", colnames(pwcs$pwc_nh_bl))], pwcs$pwc_wmhv_nh_bl)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all_bl <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all_bl$p <- ifelse(pwc_all_bl$p < 0.001, formatC(pwc_all_bl$p, format = "e", digits = 2), pwc_all_bl$p)
pwc_all_bl$p_nh <- ifelse(pwc_all_bl$p_nh < 0.001, formatC(pwc_all_bl$p_nh, format = "e", digits = 2), pwc_all_bl$p_nh)
pwc_all_bl$p_combat <- ifelse(pwc_all_bl$p_combat < 0.001, formatC(pwc_all_bl$p_combat, format = "e", digits = 2), pwc_all_bl$p_combat)
# calculate cohens d
pwc_all_bl$d <- pwc_all_bl$statistic/(sqrt(pwc_all_bl$n1))
pwc_all_bl$d_combat <- pwc_all_bl$statistic_combat/(sqrt(pwc_all_bl$n1_combat))
pwc_all_bl$d_nh <- pwc_all_bl$statistic_nh/(sqrt(pwc_all_bl$n1_nh))
pwc_all_bl[,c("d", "d_nh", "d_combat")] <- abs(round(pwc_all_bl[,c("d", "d_nh", "d_combat")], digits = 2))
pwcs$pwc_wmhv_fu$measure <- "WMHV"
pwcs$pwc_wmhv_combat_fu$measure <- "WMHV"
pwcs$pwc_wmhv_nh_fu$measure <- "WMHV"
# ignore "adjusted" cols (there was no adjustment)
pwc <- rbind(pwcs$pwc_fu[,!grepl("adj", colnames(pwcs$pwc_fu))], pwcs$pwc_wmhv_fu)
pwc_combat <- rbind(pwcs$pwc_combat_fu[,!grepl("adj", colnames(pwcs$pwc_combat_fu))], pwcs$pwc_wmhv_combat_fu)
colnames(pwc_combat) <- paste0(colnames(pwc_combat), "_combat")
pwc_nh <- rbind(pwcs$pwc_nh_fu[,!grepl("adj", colnames(pwcs$pwc_nh_fu))], pwcs$pwc_wmhv_nh_fu)
colnames(pwc_nh) <- paste0(colnames(pwc_nh), "_nh")
pwc_all_fu <- cbind(pwc, pwc_combat, pwc_nh)
pwc_all_fu$p <- ifelse(pwc_all_fu$p < 0.001, formatC(pwc_all_fu$p, format = "e", digits = 2), pwc_all_fu$p)
pwc_all_fu$p_nh <- ifelse(pwc_all_fu$p_nh < 0.001, formatC(pwc_all_fu$p_nh, format = "e", digits = 2), pwc_all_fu$p_nh)
pwc_all_fu$p_combat <- ifelse(pwc_all_fu$p_combat < 0.001, formatC(pwc_all_fu$p_combat, format = "e", digits = 2), pwc_all_fu$p_combat)
pwc_all_fu$d <- pwc_all_fu$statistic / (sqrt(pwc_all_fu$n1))
pwc_all_fu$d_combat <- pwc_all_fu$statistic_combat / (sqrt(pwc_all_fu$n1_combat))
pwc_all_fu$d_nh <- pwc_all_fu$statistic_nh / (sqrt(pwc_all_fu$n1_nh))
pwc_all_fu[, c("d", "d_nh", "d_combat")] <- abs(round(pwc_all_fu[, c("d", "d_nh", "d_combat")], 2))
```

```{r }
#| label: tbl-life-bl-icc
#| tbl-cap: "Baseline ICCs before and after harmonization in the longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df_bl[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] %>%
  kable(
    col.names = c("ROI", "ICC (95% CI)", "ICC (95% CI) ~ neuroHarm", "ICC (95% CI) ~ longCB"),
    format = "latex",
    booktabs = TRUE,
    linesep  = "",
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; ICC = intraclass correlation coefficient; CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ~ longCB = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-fu-icc
#| tbl-cap: "Follow-up ICCs before and after harmonization in the longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
df_fu[, c("ROI", "ICC_CI", "ICC_nh_CI", "ICC_combat_CI")] %>%
  kable(
    col.names = c("ROI", "ICC (95% CI)", "ICC (95% CI) ~ neuroHarm", "ICC (95% CI) ~ longCB"),
    format = "latex",
    booktabs = TRUE,
    linesep  = "",
    align = "c"
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  footnote(
    general = "ROI = region of interest; ICC = intraclass correlation coefficient; CI = confidence interval; ~ neuroHarm = harmonized with neuroHarmonize; ~ longCB = harmonized with longitudinal ComBat; TGMV = total grey matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-bl-pd
#| tbl-cap: "Baseline percent differences before and after harmonization in the longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_bl %>%
  kable(
    col.names = c("ROI", "PD", "PD ~ neuroHarm", "PD ~ longCB"),
    format = "latex",
    booktabs = TRUE,
    linesep  = ""
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(which(colnames(pd_sub2) %in% c("pre", "nh", "combat")), extra_css = "text-align: right;") %>%
  footnote(
    general = "ROI = region of interest; PD = percent difference; ~ neuroHarm = harmonized with neuroHarmonize; ~ longCB = harmonized with longitudinal ComBat; TIV = total intracranial volume; TGMV = total grey matter volume; TWMV = total white matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume",
    general_title = "",
    threeparttable = TRUE
    )
```

```{r }
#| label: tbl-life-fu-pd
#| tbl-cap: "Follow-up percent differences before and after harmonization in the longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_fu %>%
  kable(
    col.names = c("ROI", "PD", "PD ~ neuroHarm", "PD ~ longCB"),
    format = "latex",
    booktabs = TRUE,
    linesep  = ""
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  column_spec(which(colnames(pd_sub2) %in% c("pre", "nh", "combat")), extra_css = "text-align: right;") %>%
  footnote(
    general = "ROI = region of interest; PD = percent difference; ~ neuroHarm = harmonized with neuroHarmonize; ~ longCB = harmonized with longitudinal ComBat; TIV = total intracranial volume; TGMV = total grey matter volume; TWMV = total white matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume",
    general_title = "",
    threeparttable = TRUE
  )
```

```{r }
#| label: tbl-life-bl-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in the baseline longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all_bl[, c("measure", "p", "d", "p_nh", "d_nh", "p_combat", "d_combat")] |>
  kbl(
    row.names = FALSE,
    escape = FALSE,
    col.names = c(
    "ROI",
    "p-value",
    "$d_z$",
    "p-value",
    "$d_{z}$",
    "p-value",
    "$d_{z}$"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c", "c", "c", "c")
  ) |>
  add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; neuroHarm = harmonized with neuroHarmonize; ",
      "ComBat = harmonized with longitudinal ComBat; $d_z$ = Cohen's d for paired samples; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE,
    escape = FALSE
  )

```

```{r }
#| label: tbl-life-fu-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in the follow-up longitudinal data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pwc_all_fu[, c("measure", "p", "d", "p_nh", "d_nh", "p_combat", "d_combat")] |>
  kbl(
    row.names = FALSE,
    escape = FALSE,
    col.names = c(
    "ROI",
    "p-value",
    "$d_z$",
    "p-value",
    "$d_{z}$",
    "p-value",
    "$d_{z}$"
    ),
    booktabs = TRUE,
    linesep  = "",
    format   = "latex",
    align    = c("l", "c", "c", "c", "c", "c", "c")
  ) |>
  add_header_above(c(
    " " = 1,
    "unadjusted" = 2,
    "neuroHarm" = 2,
    "ComBat" = 2
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; neuroHarm = harmonized with neuroHarmonize; ",
      "ComBat = harmonized with longitudinal ComBat; $d_z$ = Cohen's d for paired samples; TGMV = total grey matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume; ",
      "WMHV = white matter hyperintensity volume"
    ),
    general_title  = "",
    threeparttable = TRUE,
    escape = FALSE
  )
```

\FloatBarrier

## Supplementary figures

```{r update_raincloud_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
library(ggdist)
df <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/df_update.csv")
df_wmhv <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/scanner_update/Data/wmhv_update.csv")
df$SITE <- gsub("_", " ", df$SITE)
df$SITE <- factor(df$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_wmhv$asinh <- asinh(df_wmhv$Lesions)
# retransform LVV
df$LVV_asinh <- df$LVV
df$LVV <- sinh(df$LVV)
colors = met.brewer(name="Java", n=5, type="discrete")
colors <- colors[c(1,2,4,5)]
colors <- rep(colors, each = 2)
make_raincloud <- function(data, col, group_val, fill_color, dotsize = 1.25) {
  data_filtered <- data %>% filter(SITE == group_val)
  
  ggplot(data_filtered, aes(x = 1, y = .data[[col]])) +
    stat_halfeye(adjust = 0.5, width = 0.6, justification = -0.2,
                 .width = 0, point_colour = NA, fill = fill_color) +
    geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.5) +
    stat_dots(side = "left", justification = 1.1, dotsize = dotsize) +
    coord_flip() +
    labs(title = as.character(group_val), x = NULL, y = col) +
    theme_classic() +
    theme(axis.text.y = element_blank(), axis.title.x = element_blank(), axis.ticks.y = element_blank())
}
# Create one plot per column
cols <- c("TIV", "TGMV", "TWMV", "TCV", "LVV", "LVV_asinh", "HCV", "AV")
group_vals <- c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat")
plots_by_col <- lapply(cols, function(col) {
  
  group_plots <- mapply(make_raincloud,
                        group_val = group_vals,
                        fill_color = colors,
                        MoreArgs = list(data = df, col = col),
                        SIMPLIFY = FALSE)
  
  wrap_plots(group_plots, ncol = 2)
})

# Name the list by column for easy access
names(plots_by_col) <- cols

# create plots for WMHV
# remove highest values to enable visualisation
df_wmhv_cut <- df_wmhv[df_wmhv$Lesions < 5000,] 
add_row_label <- function(plot_list, label) {
  plot_list[[1]] <- plot_list[[1]] + 
    labs(x = label) +
    theme(axis.title.y = element_text(angle = 90, vjust = 0.5))
  plot_list
}
unharmonized <- mapply(make_raincloud,
       group_val = c("SKYRA", "VERIO"),
       fill_color = colors[5:6],
       MoreArgs = list(data = df_wmhv_cut, col = "Lesions"),
       SIMPLIFY = FALSE)
asinh  <- mapply(make_raincloud,
                 group_val = c("SKYRA", "VERIO"),
                 fill_color = colors[5:6],
                 MoreArgs = list(data = df_wmhv_cut, col = "asinh"),
                 SIMPLIFY = FALSE)
norm  <- mapply(make_raincloud,
                 group_val = c("SKYRA", "VERIO"),
                 fill_color = colors[5:6],
                 MoreArgs = list(data = df_wmhv_cut, col = "WMHV_norm"),
                 SIMPLIFY = FALSE)
unharmonized <- add_row_label(unharmonized, "original")
asinh  <- add_row_label(asinh, "asinh")
norm   <- add_row_label(norm, expression("log"[10]))
wmhv_plots <- wrap_plots(unname(c(unharmonized, asinh, norm)), nrow = 3)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tiv-raincloud
#| fig-cap: "Raincloud plots of total intracranial volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TIV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tgmv-raincloud
#| fig-cap: "Raincloud plots of total grey matter volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TGMV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-twmv-raincloud
#| fig-cap: "Raincloud plots of total white matter volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TWMV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tcv-raincloud
#| fig-cap: "Raincloud plots of total cortical volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TCV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-lvv-raincloud
#| fig-cap: "Raincloud plots of bilateral lateral ventricle volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$LVV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-lvv-asinh-raincloud
#| fig-cap: "Raincloud plots of bilateral lateral ventricle volume by batch after inverse hyperbolic sine transformation in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$LVV_asinh)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-hcv-raincloud
#| fig-cap: "Raincloud plots of bilateral hippocampal volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$HCV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-av-raincloud
#| fig-cap: "Raincloud plots of bilateral amygdala volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$AV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-wmhv-raincloud
#| fig-cap: "Raincloud plots of white matter hyperintensity volume by batch in cross-sectional data. 1st row: original data. 2nd row: after inverse hyperbolic sine transformation. 3rd row: after decadic logarithm transformation. The maximum value is not depicted."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(wmhv_plots)
```

```{r life_raincloud_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
df <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/life/Data/df_life.csv")
df_wmhv <- read.csv("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/life/Data/df_life_wmhv.csv")
# retransform LVV
df$LVV_asinh <- df$LVV
df$LVV <- sinh(df$LVV)
df_wmhv$WMHV_asinh <- asinh(df_wmhv$WMHV)
df_wmhv$WMHV_boxcox <- df_wmhv$WMHV_norm
# how many were < 20k
n_over_10k <- length(which(df_wmhv$WMHV >= 10000))
df_wmhv <- df_wmhv[df_wmhv$WMHV < 10000,]
group_vals <- c("samseg", "recon5")
cols <- c("TGMV", "TCV", "LVV", "LVV_asinh", "HCV", "AV")
plots_by_col <- lapply(cols, function(col) {
  
  group_plots <- mapply(make_raincloud,
                        group_val = group_vals,
                        fill_color = c(colors[5], colors[3]),
                        MoreArgs = list(data = df, col = col),
                        SIMPLIFY = FALSE)
  group_plots[[1]] <- group_plots[[1]] + labs(x = col)
  
  wrap_plots(group_plots, ncol = 2)
})
plots_by_col <- wrap_plots(plots_by_col, nrow = 6)

group_vals <- c("samseg", "lst")
cols <- c("WMHV", "WMHV_asinh", "WMHV_boxcox")
plots_by_col_wmhv <- lapply(cols, function(col) {
  
  group_plots <- mapply(make_raincloud,
                        group_val = group_vals,
                        fill_color = c(colors[5], colors[3]),
                        MoreArgs = list(data = df_wmhv, col = col),
                        SIMPLIFY = FALSE)
  group_plots[[1]] <- group_plots[[1]] + labs(x = col)
  
  wrap_plots(group_plots, ncol = 2)
})

wmhv_plots <- wrap_plots(plots_by_col_wmhv, nrow = 3)
wmhv_caption <- paste0("Raincloud plots of white matter hyperintensity volume by batch in longitudinal data. 1st row: original data. 2nd row: after inverse hyperbolic sine transformation. 3rd row: after decadic logarithm transformation. The highest ",  n_over_10k, " values are not depicted.")
```

```{r fig.width=6, fig.height=12}
#| label: fig-life-raincloud
#| fig-cap: "Raincloud plots of measurements by batch in longitudinal data. TGMV = total grey matter volume; TCV = total cortical volume; LVV = bilateral lateral ventricle volume; LVV_asinh = bilateral lateral ventricle volume after inverse hyperbolic sine transformation; HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col)
```

```{r fig.width=6, fig.height=12}
#| label: fig-life-wmhv-raincloud
#| fig-cap: !expr wmhv_caption
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(wmhv_plots)
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dtgmv
#| fig-cap: "Bland-Altman plots of TGMV changes in longitudinal data. Blue vertical lines indicate the size of one standard deviation of change in TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
load("/data/pt_life/ResearchProjects/LLammer/intergeneration/segmentation_harmonization/Results/ba_plots.RData")
print(ba_plots$wrapped_life_ba_plots[[paste0("dTGMV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dtcv
#| fig-cap: "Bland-Altman plots of TCV changes in longitudinal data. Blue vertical lines indicate the size of one standard deviation of change in TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dTCV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dlvv
#| fig-cap: "Bland-Altman plots of LVV changes in longitudinal data. Blue vertical lines indicate the size of one standard deviation of change in LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dLVV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dhcv
#| fig-cap: "Bland-Altman plots of HCV changes in longitudinal data. Blue vertical lines indicate the size of one standard deviation of change in HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dHCV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-dav
#| fig-cap: "Bland-Altman plots of AV changes in longitudinal data. Blue vertical lines indicate the size of one standard deviation of change in AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$wrapped_life_ba_plots[[paste0("dAV")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-tgmv
#| fig-cap: "Bland-Altman plots of baseline TGMV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of baseline TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TGMV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-tcv
#| fig-cap: "Bland-Altman plots of baseline TCV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of baseline TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TCV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-lvv
#| fig-cap: "Bland-Altman plots of baseline LVV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of baseline LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("LVV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-hcv
#| fig-cap: "Bland-Altman plots of baseline HCV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of baseline HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("HCV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-bl-av
#| fig-cap: "Bland-Altman plots of baseline AV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of baseline AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("AV", "_bl")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-tgmv
#| fig-cap: "Bland-Altman plots of follow-up TGMV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of follow-up TGMV. TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TGMV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-tcv
#| fig-cap: "Bland-Altman plots of follow-up TCV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of follow-up TCV. TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("TCV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-lvv
#| fig-cap: "Bland-Altman plots of follow-up LVV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of follow-up LVV. LVV = bilateral lateral ventricle volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("LVV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-hcv
#| fig-cap: "Bland-Altman plots of follow-up HCV in longitudinal data. Blue vertical lines indicate the size of follow-up HCV. HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("HCV", "_fu")]])
```

```{r fig.width=6, fig.height=12}
#| label: fig-ba-fu-av
#| fig-cap: "Bland-Altman plots of follow-up AV in longitudinal data. Blue vertical lines indicate the size of one standard deviation of follow-up AV. AV = bilateral amygdala volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(ba_plots$tp_wrapped_life_ba_plots[[paste0("AV", "_fu")]])
```
