---
title: "Cross-sectional results reproduction"
author: "Laurenz Lammer"
format:
  pdf:
    papersize: a4
    keep-tex: true
    # uncomment the following line if you have downloaded the custom template from github
    # this template will ensure that the renderings are analogous to the ones in the manuscript
    #template: custom-template.tex
    header-includes:
      - \usepackage{flafter}
      - \usepackage[section]{placeins}
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
crossref:
  tbl-prefix: "Table"
  fig-prefix: "Figure"
editor:
  markdown:
    wrap: sentence
---

# Purpose of the script

This script serves to allow the reproduction of the results of our study. 
We were not able to publish the longitudinal dataset but the data from the cross-sectional scanner update study are uploaded to osf.\
Please note that the confidence intervals for ICCs are bootstrapped so that slight differences to the values in the manuscript are to be expected. 

# Reproduction of result presentation in the results section

```{r analyses}
#| echo: false
#| warning: false
#| message: false
#| error: false
# load required packages
# versions used in our analyses in respective comments
library(tidyverse) # version 2.0.0
library(lme4) # version 1.1-38
library(rstatix) # version 0.7.3
library(glue) # version 1.8.0
library(lavaan) # version 0.6-21
library(BlandAltmanLeh) # version 0.3.1
library(patchwork) # version 1.3.1
library(showtext) # version 0.9-7
library(ggplot2) # version 4.0.2

# load data from osf
df_update <- read.csv("https://osf.io/deru3/?action=download", sep = ",", header=TRUE)
# break each scanner-segmentation combination down to a scanner and a segmentation variable
df_update$scanner <- ifelse(grepl("VER", df_update$SITE), "VERIO", "SKYRA")
df_update <- df_update %>%
  mutate(segmentation = case_when(
    grepl("samseg", SITE) ~ "samseg",
    grepl("cat", SITE) ~ "cat",
    grepl("recon6", SITE) ~ "recon6",
    grepl("recon7", SITE) ~ "recon7"
  ))
# load the adjusted data from the scanner update study 
df_update_adj <- read.csv("https://osf.io/6ynfj/?action=download", sep = ",", header=TRUE)

df_update_no_correction <- read.csv("https://osf.io/58g3m/?action=download", sep = ",", header=TRUE)

df_update_no_correction_adj <- read.csv("https://osf.io/4rbt7/?action=download", sep = ",", header=TRUE)

# store the dfs in a list
update_dfs <- list(df_update = df_update, df_update_adj = df_update_adj, df_update_no_correction = df_update_no_correction,
            df_update_no_correction_adj = df_update_no_correction_adj)

# some analyses will need the dfs in the wide format
df_update_wide <- df_update %>%
  pivot_wider(
    names_from = c(SITE),
    values_from = c(TIV, TGMV, TWMV, TCV, LVV, HCV, AV),
    id_cols = id  
  )
df_update_adj_wide <- df_update_adj %>%
  pivot_wider(
    names_from = c(scanner, segmentation),
    values_from = c(TIV, TGMV, TWMV, TCV, LVV, HCV, AV),
    id_cols = id  
  )

df_update_no_correction_wide <- df_update_no_correction %>%
  pivot_wider(
    names_from = c(SITE),
    values_from = c(TIV, TGMV, TWMV, TCV, LVV, HCV, AV),
    id_cols = id  
  )
df_update_no_correction_adj_wide <- df_update_no_correction_adj %>%
  pivot_wider(
    names_from = c(scanner, segmentation),
    values_from = c(TIV, TGMV, TWMV, TCV, LVV, HCV, AV),
    id_cols = id  
  )

# store wide dfs in a list
update_wide_dfs <- list(df_update_wide = df_update_wide, df_update_adj_wide = df_update_adj_wide, 
                 df_update_no_correction_wide = df_update_no_correction_wide, df_update_no_correction_adj_wide = df_update_no_correction_adj_wide)

# produce dfs without NAs for pairise t-tests
update_pairwise_dfs <- map(
  update_dfs,
  ~ .x %>%
    group_by(id) %>%
    filter(any(scanner == "SKYRA") & any(scanner == "VERIO")) %>%
    ungroup())

# prepare a results df to be filled
update_lme_icc_res <- data.frame(matrix(ncol = 61, nrow = 7))
colnames(update_lme_icc_res) <- c("ROI", paste0(rep(paste0(rep(paste0(rep(c("id_var", "scanner_var", "segmentation_var", "resdiual_var", "ICC"), each = 3), 
                                       c("", "_lower_95_CI", "_upper_95_CI")), each = 2), c("", "_no_correction")), times = 2), 
                                       rep(c("", "_adj"), each = 30)))
update_lme_icc_res$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")

# formula for bootstrapped CIs for ICC
icc_ci <- function(fit) {
  variances <- as.data.frame(VarCorr(fit))
  id_var = variances[variances$grp == "id", "vcov"]
  scan_var = variances[variances$grp == "scanner", "vcov"]
  seg_var = variances[variances$grp == "segmentation", "vcov"] 
  residual_var = variances[variances$grp == "Residual", "vcov"]
  icc = id_var / (id_var + scan_var + seg_var + residual_var)
  c(id_var, scan_var, seg_var, residual_var, icc)
}

# calculate LMEs
# iterate over outcomes
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")) {
  # iterate over dfs
  for (df_name in names(update_dfs)){
    df <- update_dfs[[df_name]] 
    suffix <- substring(df_name, 10)
    # remove SKYRA-recon6 rows if outcome is TIV (uninformative compared to VERIO-recon6)
    if (outcome == "TIV"){
      df <- df[!(df$scanner == "SKYRA" & df$segmentation == "recon6"),]
    }
    # create outcome specific LME formula
    formula <- as.formula(paste0(outcome, " ~ (1|id) + (1|scanner) + (1|segmentation)"))
    # estimate model and bootstrap confidence intervals
    res <- lmer(formula = formula, data = df, control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
    variances <- as.data.frame(VarCorr(res))
    lme_icc <- (variances[variances$grp == "id", "vcov"])/
      (variances[variances$grp == "scanner", "vcov"] + variances[variances$grp == "id", "vcov"] + 
         variances[variances$grp == "segmentation", "vcov"] + variances[variances$grp == "Residual", "vcov"])
    boot_icc <- bootMer(res, FUN  = icc_ci, nsim = 1000, type = "parametric")
    boot_res <- apply(boot_icc$t, 2, quantile, probs = c(0.025, 0.975))
    # store the results in the appropriate columns (unadjusted/adjusted data)
    update_lme_icc_res[update_lme_icc_res$ROI == outcome, paste0(c("id_var", "scanner_var", "segmentation_var", "resdiual_var"), suffix)] <- 
      c(variances[variances$grp == "id", "vcov"], variances[variances$grp == "scanner", "vcov"], 
        variances[variances$grp == "segmentation", "vcov"], variances[variances$grp == "Residual", "vcov"])
    update_lme_icc_res[update_lme_icc_res$ROI == outcome, paste0("ICC", suffix)] <- lme_icc
    update_lme_icc_res[update_lme_icc_res$ROI == outcome, paste0(paste0(rep(c("id_var", "scanner_var", "segmentation_var", "resdiual_var", "ICC"),
                                                              each = 2), c("_lower_95_CI", "_upper_95_CI")), suffix)] <- as.vector(boot_res)
  }
}

############ Section 1.2: Calculation of PDs #######################
# prepare a list to store the pd_matrices
update_pd_matrices <- list()
# iterate over ROIs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")) {
  # iterate over datasets
  for (df_name in names(update_wide_dfs)){
    df <- update_wide_dfs[[df_name]]
    suffix <- substr(df_name,10,nchar(df_name)-5)
    # drop the SKYRA-recon6 TIV col
    df$TIV_SKYRA_recon6 <- NULL
    # get names of relevant cols
    roi_cols <- grep(outcome, colnames(df), value = TRUE)
    # initialize empty matrix
    mean_pd_mat <- matrix(
      NA_real_,
      nrow = length(roi_cols),
      ncol = length(roi_cols),
      dimnames = list(roi_cols, roi_cols)
    )
    # fill upper triangle
    for (i in seq_along(roi_cols)) {
      for (j in seq_along(roi_cols)) {
        if (i < j) {
          # extract all values in the 2 relevant cols
          col1 <- df[[roi_cols[i]]]
          col2 <- df[[roi_cols[j]]]
          # calculate a vector of PDs
          # a negative value indicates that the values of the batch given by the rowname is smaller 
          pd <- (2 * (col1 - col2)) / (abs(col1) + abs(col2))
          # calculate the mean PD between the two batches (NAs exist because 5 participants were not scanned on both scanners)
          mean_pd_mat[i, j] <- mean(pd, na.rm = TRUE)
        }
      }
    }
    # mirror the inverse to lower triangle
    mean_pd_mat[lower.tri(mean_pd_mat)] <- t(mean_pd_mat)[lower.tri(mean_pd_mat)] * (-1)
    # turn into percentage values
    mean_pd_mat <- mean_pd_mat*100
    # save variables to list
    update_pd_matrices[[paste0(outcome, "_mean_pd_mat",suffix)]] <- mean_pd_mat
  }
}

# prepare a results df to be filled
update_anova_res <- data.frame(matrix(ncol = 25, nrow = 7))
colnames(update_anova_res) <- c("ROI", paste0(rep(c("DFn",  "DFd",  "F",  "p",  "p<.05",  "ges"), times = 4), 
                                       rep(c("", "_adj", "_no_correction", "_no_correction_adj"), each = 6)))
update_anova_res$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
#prepare list to store pairwise comparisons
update_pwcs <- list()
# iterate over ROIs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")) {
  # iterate over datasets
  for (df_name in names(update_dfs)){
    df <- update_dfs[[df_name]]
    suffix <- substring(df_name, 10)
    pairwise_df <- update_pairwise_dfs[[df_name]]
    # remove SKYRA-recon6 rows if outcome is TIV (uninformative compared to VERIO-recon6)
    if (outcome == "TIV"){
      df <- df[!(df$scanner == "SKYRA" & df$segmentation == "recon6"),]
      pairwise_df <- pairwise_df[!(pairwise_df$scanner == "SKYRA" & pairwise_df$segmentation == "recon6"),]
    }
    res.aov <- anova_test(data = df, dv = outcome, wid = id, within = SITE)
    table <- get_anova_table(res.aov)
    update_anova_res[update_anova_res$ROI == outcome, paste0(c("DFn",  "DFd",  "F",  "p",  "p<.05",  "ges"), suffix)] <- as.data.frame(table)[1,2:7]
    # perform pairwise comparisons of the individual scanner-segmentation comparisons
    formula <- as.formula(paste0(outcome, " ~ SITE"))
    pwc <- pairwise_df %>%
      pairwise_t_test(
        formula = formula, paired = TRUE, id = "id",
        p.adjust.method = "hommel"
      )
    update_pwcs[[paste0(outcome, "_pwc", suffix)]] <- pwc
  }
}

# prepare a results df to be filled
update_iced_res <- data.frame(matrix(ncol = 85, nrow = 7))
parameters <- c("obsvar", "truevar", "SKYRAvar",  "VERIOvar", "samsegvar", "recon6var", "recon7var", "catvar",
                "VERIOmean",  "recon6mean", "recon7mean", "catmean", "truemean")
col_names <- c(parameters, paste0(rep(c("ICC_SKYRA_", "ICC_VERIO_"), each = 4), rep(c("samseg", "recon6", "recon7", "cat"), times = 2))) 
                
colnames(update_iced_res) <- c("ROI", paste0(rep(col_names, times = 4), rep(c("", "_adj", "_no_correction", "_no_correction_adj"), each = 21)))
update_iced_res$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
# iterate over ROIs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")) {
  # create colname vectors to adapt the model to the respective ROI
  obsvars <- grep(outcome, colnames(df_update_wide), value = T)
  skyravars <- grep("SKYRA", obsvars, value = T)
  veriovars <- grep("VERIO", obsvars, value = T)
  samsegvars <- grep("samseg", obsvars, value = T)
  recon6vars <- grep("recon6", obsvars, value = T)
  recon7vars <- grep("recon7", obsvars, value = T)
  catvars <- grep("cat", obsvars, value = T)
  # write the lavvan model
  model <- glue("
   ! regressions 
   truth=~1.0*{obsvars[1]}
   truth=~1.0*{obsvars[2]}
   truth=~1.0*{obsvars[3]}
   truth=~1.0*{obsvars[4]}
   truth=~1.0*{obsvars[5]}
   truth=~1.0*{obsvars[6]}
   truth=~1.0*{obsvars[7]}
   truth=~1.0*{obsvars[8]}
   SKYRA=~1.0*{skyravars[1]}
   SKYRA=~1.0*{skyravars[2]}
   SKYRA=~1.0*{skyravars[3]}
   SKYRA=~1.0*{skyravars[4]}
   VERIO=~1.0*{veriovars[1]}
   VERIO=~1.0*{veriovars[2]}
   VERIO=~1.0*{veriovars[3]}
   VERIO=~1.0*{veriovars[4]}
   samseg=~1.0*{samsegvars[1]}
   samseg=~1.0*{samsegvars[2]}
   recon6=~1.0*{recon6vars[1]}
   recon6=~1.0*{recon6vars[2]}
   recon7=~1.0*{recon7vars[1]}
   recon7=~1.0*{recon7vars[2]}
   cat=~1.0*{catvars[1]}
   cat=~1.0*{catvars[2]}
! residuals, variances and covariances
   {obsvars[1]} ~~ start(0.01)*obsvar*{obsvars[1]}
   {obsvars[2]} ~~ start(0.01)*obsvar*{obsvars[2]}
   {obsvars[3]} ~~ start(0.01)*obsvar*{obsvars[3]}
   {obsvars[4]} ~~ start(0.01)*obsvar*{obsvars[4]}
   {obsvars[5]} ~~ start(0.01)*obsvar*{obsvars[5]}
   {obsvars[6]} ~~ start(0.01)*obsvar*{obsvars[6]}
   {obsvars[7]} ~~ start(0.01)*obsvar*{obsvars[7]}
   {obsvars[8]} ~~ start(0.01)*obsvar*{obsvars[8]}
   truth ~~ start(0.1)*truevar*truth
   SKYRA ~~ start(0.0001)*SKYRAvar*SKYRA
   VERIO ~~ start(0.0001)*VERIOvar*VERIO
   samseg ~~ start(0.001)*samsegvar*samseg
   recon6 ~~ start(0.001)*recon6var*recon6
   recon7 ~~ start(0.001)*recon7var*recon7
   cat ~~ start(0.001)*catvar*cat
   truth ~~ 0.0*SKYRA
   truth ~~ 0.0*VERIO
   truth ~~ 0.0*samseg
   truth ~~ 0.0*recon6
   truth ~~ 0.0*recon7
   truth ~~ 0.0*cat
   SKYRA ~~ 0.0*VERIO
   SKYRA ~~ 0.0*samseg
   SKYRA ~~ 0.0*recon6
   SKYRA ~~ 0.0*recon7
   SKYRA ~~ 0.0*cat
   VERIO ~~ 0.0*samseg
   VERIO ~~ 0.0*recon6
   VERIO ~~ 0.0*recon7
   VERIO ~~ 0.0*cat
   samseg ~~ 0.0*recon6
   samseg ~~ 0.0*recon7
   samseg ~~ 0.0*cat
   recon6 ~~ 0.0*recon7
   recon6 ~~ 0.0*cat
   recon7 ~~ 0.0*cat
! means
   # the means are relative to samseg-SKYRA
   SKYRA~0*1;
   samseg~0*1;
   VERIO~VERIOmean*1
   recon6~recon6mean*1
   recon7~recon7mean*1
   cat~catmean*1
   truth~truemean*1
   {obsvars[1]}~0*1;
   {obsvars[2]}~0*1;
   {obsvars[3]}~0*1;
   {obsvars[4]}~0*1;
   {obsvars[5]}~0*1;
   {obsvars[6]}~0*1;
   {obsvars[7]}~0*1;
   {obsvars[8]}~0*1;
                ")
  # adapt model if the ROI is TIV (remove SKYRA-recon6)
  if (outcome == "TIV"){
    drop_lines <- c(
      "TIV_SKYRA_recon6~0*1;",
      "TIV_SKYRA_recon6 ~~ obsvar*TIV_SKYRA_recon6",
      "SKYRA=~1.0*TIV_SKYRA_recon6",
      "truth=~1.0*TIV_SKYRA_recon6",
      "recon6=~1.0*TIV_SKYRA_recon6")
    pattern <- paste0(
      "(?m)^\\s*(",
      paste0("\\Q", drop_lines, "\\E", collapse = "|"),
      ")\\s*$")
    model <- str_remove_all(model, pattern)
  }
  # iterate over datasets
  for (df_name in names(update_wide_dfs)){
    df <- update_wide_dfs[[df_name]]
    suffix <- substr(df_name,10,nchar(df_name)-5)
    # estimate the lavaan model
    fit <- lavaan(model = model, data=df, fixed.x=FALSE, missing="FIML", control = list(iter.max = 5000))
    pe <- parameterEstimates(fit)
    pe <- pe[!duplicated(pe$label, na.rm = T), ]
    # set negative variances to 0 (might occur if true variance is very close to 0)
    pe$est[pe$est < 0 & grepl("var", pe$label)] <- 0
    # prepare iced icc list
    iced_iccs <- list(
      SKYRA_samseg = NULL,
      SKYRA_recon6 = NULL,
      SKYRA_recon7 = NULL,
      SKYRA_cat = NULL,
      VERIO_samseg = NULL,
      VERIO_recon6 = NULL,
      VERIO_recon7 = NULL,
      VERIO_cat = NULL
    )
    # calculate ICED-ICC for each batch
    for (scanner in c("SKYRA", "VERIO")) {
      for (segmentation in c("samseg", "recon6", "recon7", "cat")){
       icc <- pe[pe$label == "truevar", "est"] / (pe[pe$label == "truevar", "est"] + pe[pe$label == "obsvar", "est"] +
                                                    pe[pe$label == paste0(scanner, "var"), "est"] + pe[pe$label == paste0(segmentation, "var"), "est"])
       iced_iccs[[paste0(scanner, "_", segmentation)]] <- icc
       }
    }
    # store the results
    update_iced_res[update_iced_res$ROI == outcome, paste0(parameters, suffix)] <- pe[pe$label %in% parameters, "est"]
    update_iced_res[update_iced_res$ROI == outcome, c(paste0("ICC_SKYRA_", c("samseg", "recon6", "recon7", "cat"), suffix), 
                                        paste0("ICC_VERIO_", c("samseg", "recon6", "recon7", "cat"), suffix))] <- iced_iccs
    # check if non-convergences occured (lavaan just gives starting values in this case)
    if (!lavaan::lavInspect(fit, "converged")) {
      update_iced_res[update_iced_res$ROI == outcome, paste0(parameters, suffix)] <- "uconverged"
      update_iced_res[update_iced_res$ROI == outcome, c(paste0("ICC_SKYRA_", c("samseg", "recon6", "recon7", "cat"), suffix), 
                      paste0("ICC_VERIO_", c("samseg", "recon6", "recon7", "cat"), suffix))] <- "unconverged"
    }
  }
}

showtext_auto()
update_ba_plots <- list()
update_ba_plot_comparisons <- list()
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")) {
  # iterate over datasets
  for (df_name in names(update_wide_dfs)){
    plot_list <- list()
    df <- update_wide_dfs[[df_name]]
    suffix <- substring(df_name, 10)
    cols <- grep(outcome, colnames(df), value = T)
    # get all combinations of cols
    combs <- combn(cols, 2, simplify = FALSE)
    for (combination in combs){
      plt <- bland.altman.plot(group1 = df[[combination[1]]], group2 = df[[combination[2]]], 
                               graph.sys = "ggplot2", conf.int=.95, pch=19)
      plt <- plt + labs(title = paste0(gsub("_", " ", combination[1]), " vs. ", gsub("_", " ", combination[2]))) + theme_bw(base_family = "garamond") + 
        theme(plot.title = element_text(hjust = 0.5, size = 12))
      plot_list[[paste0(combination[1], " vs. ", combination[2])]] <- plt
    }
    plots <- wrap_plots(plot_list, ncol(6))
    update_ba_plots[[paste0(outcome, suffix)]] <- plots
    update_ba_plots[[paste0(outcome, suffix, "_indiv_plots")]] <- plot_list
  }
  for (i in seq(1, 28, by = 4)) {
    plots_left  <- update_ba_plots[[paste0(outcome, "_wide_indiv_plots")]][i:(i+3)]
    plots_right <- update_ba_plots[[paste0(outcome, "_adj_wide_indiv_plots")]][i:(i+3)]
    left_col  <- wrap_plots(plots_left, ncol = 1)
    right_col <- wrap_plots(plots_right, ncol = 1)
    combined <- wrap_plots(left_col, right_col, ncol = 2) + plot_annotation(title = paste0(outcome, " before (left) and after (right) harmonization"), 
                                                                            theme = theme(plot.title = element_text(hjust = 0.5)))
    update_ba_plot_comparisons[[paste0(outcome, (((i-1)/4)+1))]] <- combined
  }
}

############ Section 3.3 Scanner update WMHV ################
df_update_wmhv <- read.csv("https://osf.io/mxde5/?action=download", sep = ",", header=TRUE)
df_update_wmhv_wide <- df_update_wmhv[,c("id", "SITE", "WMHV_norm", "WMHV_adj")] %>%
  pivot_wider(
    names_from = SITE,
    values_from = c(WMHV_norm, WMHV_adj)
  )
############ Section 3.3.1: Calculation of ICCs ####################### 

# add row to update_lme_icc_res
update_lme_icc_res[nrow(update_lme_icc_res) + 1, "ROI"] <- "WMHV"

# formula for bootstrapped CIs for ICC
icc_ci <- function(fit) {
  variances <- as.data.frame(VarCorr(fit))
  id_var = variances[variances$grp == "id", "vcov"]
  scan_var = variances[variances$grp == "scanner", "vcov"]
  residual_var = variances[variances$grp == "Residual", "vcov"]
  icc = id_var / (id_var + scan_var + residual_var)
  c(id_var, scan_var, residual_var, icc)
}

# calculate LMEs
# iterate over outcomes
for (outcome in c("WMHV_norm", "WMHV_adj")) {
  # get suffix
  suffix <- ""
  if (outcome == "WMHV_adj"){
    suffix <- "_adj"
  }
  formula <- as.formula(paste0(outcome, " ~ (1|id) + (1|scanner)"))
  # estimate model and bootstrap confidence intervals
  res <- lmer(formula = formula, data = df_update_wmhv, control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
  variances <- as.data.frame(VarCorr(res))
  lme_icc <- (variances[variances$grp == "id", "vcov"])/
    (variances[variances$grp == "scanner", "vcov"] + variances[variances$grp == "id", "vcov"] + 
       variances[variances$grp == "Residual", "vcov"])
  boot_icc <- bootMer(res, FUN  = icc_ci, nsim = 1000, type = "parametric")
  boot_res <- apply(boot_icc$t, 2, quantile, probs = c(0.025, 0.975))
  # store the results in the appropriate columns (unadjusted/adjusted data)
  update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", paste0(c("id_var", "scanner_var", "resdiual_var"), suffix)] <- 
    c(variances[variances$grp == "id", "vcov"], variances[variances$grp == "scanner", "vcov"], 
      variances[variances$grp == "Residual", "vcov"])
  update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", paste0("ICC", suffix)] <- lme_icc
  update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", paste0(paste0(rep(c("id_var", "scanner_var", "resdiual_var", "ICC"),
                                                                          each = 2), c("_lower_95_CI", "_upper_95_CI")), suffix)] <- as.vector(boot_res)
}

############ Section 3.3.2: Calculation of PDs #######################

# iterate over ROIs
for (outcome in c("WMHV_norm", "WMHV_adj")) {
  suffix <- ""
  if (outcome == "WMHV_adj"){
    suffix <- "_adj"
  }  
  # get names of relevant cols
  roi_cols <- grep(outcome, colnames(df_update_wmhv_wide), value = TRUE)
  # initialize empty matrix
  mean_pd_mat <- matrix(
    NA_real_,
    nrow = length(roi_cols),
    ncol = length(roi_cols),
    dimnames = list(roi_cols, roi_cols)
  )
  # fill upper triangle
  for (i in seq_along(roi_cols)) {
    for (j in seq_along(roi_cols)) {
      if (i < j) {
        # extract all values in the 2 relevant cols
        col1 <- df_update_wmhv_wide[[roi_cols[i]]]
        col2 <- df_update_wmhv_wide[[roi_cols[j]]]
        # calculate a vector of PDs
        # a negative value indicates that the values of the batch given by the rowname is smaller 
        pd <- (2 * (col1 - col2)) / (abs(col1) + abs(col2))
        # calculate the mean PD between the two batches (NAs exist because 5 participants were not scanned on both scanners)
        mean_pd_mat[i, j] <- mean(pd, na.rm = TRUE)
      }
    }
  }
  # mirror the inverse to lower triangle
  mean_pd_mat[lower.tri(mean_pd_mat)] <- t(mean_pd_mat)[lower.tri(mean_pd_mat)] * (-1)
  # turn into percentage values
  mean_pd_mat <- mean_pd_mat*100
  # save variables to list
  update_pd_matrices[[paste0(outcome, "_mean_pd_mat",suffix)]] <- mean_pd_mat
}


############ Section 3.3.3: Calculation of pairwise t-tests #######################

# iterate over ROIs
for (outcome in c("WMHV_norm", "WMHV_adj")) {
  suffix <- ""
  if (outcome == "WMHV_adj"){
    suffix <- "_adj"
  }  
  # perform pairwise comparisons 
  formula <- as.formula(paste0(outcome, " ~ SITE"))
  pwc <- df_update_wmhv %>%
    group_by(id) %>%
    filter(n() == 2) %>%
    ungroup() %>%
    pairwise_t_test(
      formula = formula, paired = TRUE, id = "id"
    )
  update_pwcs[[paste0(outcome, "_pwc", suffix)]] <- pwc
}
############ Section 3.3.4 Bland-Altman plots ################
for (outcome in c("WMHV_norm", "WMHV_adj")) {
  cols <- grep(outcome, colnames(df_update_wmhv_wide), value = T)
  # add a vertical line to show the size of 1 SD
  # determine the x-coordinate (should be left of the smallest value)
  x_coord = min(rowMeans(df_update_wmhv_wide[,cols]), na.rm = T)
  # determine the y-coordinates 
  sd_val <- sd(unlist(df_update_wmhv_wide[cols]), na.rm = TRUE)
  y_coord_a <- sd_val/2 
  y_coord_b <- -sd_val/2
  plt <- bland.altman.plot(group1 = df_update_wmhv_wide[[cols[1]]], group2 = df_update_wmhv_wide[[cols[2]]], 
                           graph.sys = "ggplot2", conf.int=.95, pch=19)
  plt <- plt + labs(title = paste0(cols[1], " vs. ", cols[2])) + theme(plot.title = element_text(size = 12), base_family = "garamond")
  # add line to idicate size of SD
  plt <- plt + 
    geom_segment(
      x = x_coord, xend = x_coord,
      y = y_coord_a, yend = y_coord_b,
      colour = "blue",
      linewidth = 1
    )
  update_ba_plots[[paste0(outcome)]] <- plt
}
unharm_plot <- update_ba_plots[["WMHV_norm"]] + labs(title = paste0("WMHV SKYRA vs. WMHV VERIO unharmonized")) 
nh_plot <- update_ba_plots[["WMHV_adj"]] + labs(title = paste0("WMHV SKYRA vs. WMHV VERIO harmonized")) 
plots <- wrap_plots(list(unharm_plot, nh_plot), nrow = 2, )
update_ba_plot_comparisons[["WMHV"]] <- plots 

```

```{r update_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
# set font for figures
font_add_google("Fira Sans", "firasans")
showtext_auto()
theme_set(theme_classic(base_family = "firasans", base_size = 11))

update_lme_icc_res_no_wmhv <- update_lme_icc_res[update_lme_icc_res$ROI != "WMHV",]
pd_mat <- update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- update_anova_res
pwcs <- update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc")]]
  post <- pwcs[[paste0(outcome, "_pwc_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p <- ifelse(anovas$p < 0.001, formatC(anovas$p, format = "e", digits = 2), anovas$p)
anovas$p_adj <- ifelse(anovas$p_adj < 0.001, formatC(anovas$p_adj, format = "e", digits = 2), anovas$p_adj)
anovas$ges <- ifelse(anovas$ges < 0.001, formatC(anovas$ges, format = "e", digits = 2), anovas$ges)
anovas$ges_adj <- ifelse(anovas$ges_adj < 0.001, formatC(anovas$ges_adj, format = "e", digits = 2), anovas$ges_adj)
iced <- update_iced_res
iced <- iced[iced$ROI != "TIV",]
iced[, 2:ncol(iced)] <- apply(iced[, 2:ncol(iced)], 2, function(x) ifelse(abs(as.numeric(x)) < 0.001 & as.numeric(x) != 0, formatC(as.numeric(x), format = "e", digits = 0), round(as.numeric(x), digits = 3)))
```

```{r}
#| label: tbl-update-icc
#| tbl-cap: "ICCs before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC, ICC_lower_95_CI, ICC_upper_95_CI),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_adj, ICC_lower_95_CI_adj, ICC_upper_95_CI_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricles volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif
#| tbl-cap: "Statistical significance of differences before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p", "ges", "n_sig_pre", "p_adj", "ges_adj", "n_sig_post")] |>
  kbl(
    escape = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r", "r", "r")
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 3,
    "harmonized" = 3
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
  general = paste0(
    "ROI = region of interest; p-value = p-value from ANOVA; ",
    "$\\\\eta^2$", " = generalized ", "$\\\\eta^2$", " from ANOVA;",
    "harm. = harmonized with neuroHarmonize; ",
    "sig. PWCs = share of significant pairwise comparisons; ",
    "TIV = total intracranial volume; TGMV = total grey matter volume; ",
    "TWMV = total white matter volume; TCV = total cortical volume; ",
    "LVV = bilateral lateral ventricle volume; ",
    "HCV = bilateral hippocampal volume; ",
    "AV = bilateral amygdala volume"
  ),
  general_title  = "",
  threeparttable = TRUE,
  escape = FALSE
  )
```

```{r}
#| label: tbl-update-pd
#| tbl-cap: "Percent differences before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percent difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

\FloatBarrier
\newpage

::: landscape
```{r}
#| label: fig-update
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline TIV = Total Intracranial Volume; TGMV = Total Grey Matter Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
library(tidyverse)
library(MetBrewer)
library(patchwork)
library(showtext)
showtext_auto()
font_add_google("EB Garamond")
colors = met.brewer(name="Java", n=5, type="discrete")
colors <- colors[c(1,2,4,5)]
# break each scanner-segmentation combination down to a scanner and a segmentation variable
df_update$scanner <- ifelse(grepl("VER", df_update$SITE), "VERIO", "SKYRA")
df_update <- df_update %>%
  mutate(segmentation = case_when(
    grepl("samseg", SITE) ~ "samseg",
    grepl("cat", SITE) ~ "cat",
    grepl("recon6", SITE) ~ "recon6",
    grepl("recon7", SITE) ~ "recon7"
  ))

# load the harmonized data from the scanner update study 
df_update_adj <- read.csv("https://osf.io/6ynfj/?action=download", sep = ",", header=TRUE)

# fill in the missing data from the unadjusted df (these cols must not be included during harmonization) 
df_update$SITE <- gsub("_", " ", df_update$SITE)
df_update$SITE <- factor(df_update$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))
df_update_adj[,c("id", "SITE", "age", "gender")] <- df_update[,c("id", "SITE", "age", "gender")]
df_update_wmhv$SITE <- ifelse(df_update_wmhv$scanner == "skyra", "SKYRA samseg", "VERIO samseg")
df_update_wmhv$SITE <- factor(df_update_wmhv$SITE, levels = c("SKYRA samseg", "VERIO samseg"))
df_update_wmhv$WMHV <- df_update_wmhv$WMHV_norm
df_update_wmhv_adj <- df_update_wmhv
df_update_wmhv_adj$WMHV <- df_update_wmhv$WMHV_adj
plots <- list()
comparisons <- list()
dfs <- list(df_update = df_update, df_update_adj = df_update_adj)
wmhv_dfs <- list(df_update_wmhv = df_update_wmhv, df_update_wmhv_adj = df_update_wmhv_adj)
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV", "HCV", "AV")){
  for (name in names(dfs)){
    df <- dfs[[name]]
    suffix <- ""
    y_axis_title <- outcome
    if (grepl("adj", name)){
      suffix <- "_adj"
      y_axis_title <- paste0(outcome, " harmonized")
    }
    ordered_ids <- df %>%
      group_by(id) %>%
      summarise(mean_val = mean(.data[[outcome]], na.rm = TRUE)) %>%
      arrange(mean_val) %>%
      pull(id)
    df$id <- factor(df$id, levels = ordered_ids)
    plt <- ggplot(data = df, aes(x = factor(id), y = .data[[outcome]], color = SITE, group = SITE, shape = SITE)) +
      geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
      scale_shape_manual(values = c(16, 1, 16, 1, 16, 1, 16, 1)) +
      theme_classic(base_family = "garamond") +
      guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) + 
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.0, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.5, "cm"), legend.key.spacing.y = unit(0.0, "cm"), legend.key.spacing.x = unit(0, "cm"),
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
      ylab(y_axis_title) + scale_color_manual(values = rep(colors, each = 2))
    if (grepl("adj", name)){
      plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
    }
    plots[[paste0(outcome, suffix)]] <- plt
  }
  y_range <- range(c(plots[[outcome]]$data[[outcome]], plots[[paste0(outcome, "_adj")]]$data[[outcome]]), na.rm = TRUE)
  plots[[outcome]] <- plots[[outcome]] + coord_cartesian(ylim = y_range)
  if(outcome %in% c("TGMV", "TCV", "HCV")){
    plots[[outcome]] <- plots[[outcome]] + labs(x = "individuals (sorted by mean across batches)") + theme(axis.title.x = element_text(size = 5, margin = margin(t = 6)))
  }
  plots[[paste0(outcome, "_adj")]] <- plots[[paste0(outcome, "_adj")]] + coord_cartesian(ylim = y_range)
  comparison <- wrap_plots(list(plots[[outcome]], plots[[paste0(outcome, "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
  comparisons[[outcome]] <- comparison
}
for (name in names(wmhv_dfs)){
  df <- wmhv_dfs[[name]]
  suffix <- ""
  y_axis_title <- "WMHV"
  if (grepl("adj", name)){
    suffix <- "_adj"
    y_axis_title <- paste0("WMHV harmonized")
  }
  ordered_ids <- df %>%
    group_by(id) %>%
    summarise(mean_val = mean(.data[["WMHV"]], na.rm = TRUE)) %>%
    arrange(mean_val) %>%
    pull(id)
  df$id <- factor(df$id, levels = ordered_ids)
  plt <- ggplot(data = df, aes(x = factor(id), y = .data[["WMHV"]], color = SITE, group = SITE, shape = SITE)) +
    geom_point(size = 0.75, alpha = 0.85, stroke = 0.4) +
    scale_shape_manual(values = c(16, 1)) +
    theme_classic(base_family = "garamond") +
    guides(color = guide_legend(ncol = 4, override.aes = list(size = 1.6)), shape = guide_legend(ncol = 4, override.aes = list(size = 1.6))) +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(size = 4), 
            axis.title.y = element_text(size = 5), legend.text = element_text(size = 5, margin = margin(r = 8)), legend.title = element_blank(), 
            legend.spacing.y = unit(0.0, "cm"), legend.spacing.x = unit(0.1, "cm"), legend.margin = margin(0, 0, 0, 0), legend.box.spacing = unit(0, "cm"),
      legend.key.width = unit(0.5, "cm"), legend.key.height = unit(0.5, "cm"), legend.key.spacing.y = unit(0.0, "cm"), legend.key.spacing.x = unit(0, "cm"),
      axis.line.x = element_line(linewidth = 0.3), axis.line.y = element_line(linewidth = 0.3)) +
    ylab(y_axis_title) + scale_color_manual(values = rep(colors[3], 2))
  if (grepl("adj", name)){
    plt <- plt + theme(axis.text.y  = element_blank(), axis.ticks.y = element_blank())
  }
  plots[[paste0("WMHV", suffix)]] <- plt
}
y_range <- range(c(plots[["WMHV"]]$data[["WMHV"]], plots[[paste0("WMHV", "_adj")]]$data[["WMHV"]]), na.rm = TRUE)
plots[["WMHV"]] <- plots[["WMHV"]] + coord_cartesian(ylim = y_range) + labs(x = "individuals (sorted by mean across batches)") + theme(axis.title.x = element_text(size = 5, margin = margin(t = 6)))
plots[[paste0("WMHV", "_adj")]] <- plots[[paste0("WMHV", "_adj")]] + coord_cartesian(ylim = y_range)
comparison <- wrap_plots(list(plots[["WMHV"]], plots[[paste0("WMHV", "_adj")]]), nrow = 1) + plot_layout(guides = "collect") & theme(legend.position = "right")
comparisons[["WMHV"]] <- comparison

legend_only_plot <- ggplot() + 
  theme_void() +
  annotation_custom(cowplot::get_legend(comparisons[[1]]))
p_combined <- wrap_plots(comparisons[[1]], comparisons[[2]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update2
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline TWMV = Total White Matter Volume; TCV = Total Cortical Volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[3]], comparisons[[4]],
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update3
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[5]], comparisons[[6]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

```{r}
#| label: fig-update4
#| fig-cap: "Distribution of volumetric estimates by batch before and after harmonization in cross-sectional data \\newline AV = bilateral amygdala volume; WMHV = white matter hyperintensity volume"
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "100%"
p_combined <- wrap_plots(comparisons[[7]], comparisons[[8]], 
                         nrow = 2, ncol = 1, guides = "collect") + plot_layout(widths = 1) &
  theme(legend.position = "none", , plot.margin = margin(4,4,4,4))
cowplot::plot_grid(p_combined, legend_only_plot, ncol = 1, rel_heights = c(1, 0.1))
```

\newpage
:::

\FloatBarrier


The agreement between WMHV using the same segmentation method on different MRI machines in the cross-sectional data showed an ICC of `r round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC"], digits = 3)` (95% CI: `r  round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI"], digits = 3)`--`r  round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI"], digits = 3)`) before and of `r round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC_adj"], digits = 3)` (95% CI: `r  round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC_lower_95_CI_adj"], digits = 3)`--`r  round(update_lme_icc_res[update_lme_icc_res$ROI == "WMHV", "ICC_upper_95_CI_adj"], digits = 3)`) after harmonization.
Measurements on the SKYRA were on average `r round(update_pd_matrices$WMHV_norm_mean_pd_mat[1,2], digits = 1)`% larger before harmonization but only `r abs(round(update_pd_matrices$WMHV_adj_mean_pd_mat_adj[1,2], digits = 1))`% smaller after the application of neuroHarmonize.
A paired t-test of the batches was significant prior to harmonization (p = `r formatC(update_pwcs$WMHV_norm_pwc$p, digits = 2, format = "e")`; t(`r update_pwcs$WMHV_adj_pwc_adj$df`) = `r formatC(update_pwcs$WMHV_norm_pwc$statistic, digits = 2, format = "g")`; d = `r  round(abs((update_pwcs$WMHV_norm_pwc$statistic)/(sqrt(update_pwcs$WMHV_norm_pwc$n1))), digits = 2)`) but ceased to be so afterwards (p = `r  update_pwcs$WMHV_adj_pwc_adj$p`; t(`r update_pwcs$WMHV_adj_pwc_adj$df`) = `r formatC(update_pwcs$WMHV_adj_pwc_adj$statistic, digits = 2, format = "g")`; d = `r round(abs((update_pwcs$WMHV_adj_pwc_adj$statistic)/(sqrt(update_pwcs$WMHV_adj_pwc_adj$n1))), digits = 2)`).

# Reproduction of supplementary tables

```{r scanner_software_comp}
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp <- data.frame(matrix(nrow = 7, ncol = 3))
colnames(scanner_software_comp) <- c("ROI", "within_scanner_diff", "within_software_diff")
scanner_software_comp$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
# start with scanner update
pd <- update_pd_matrices
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  mat <- pd[[paste0(outcome,"_mean_pd_mat")]]
  transf_mat <- abs(mat) %>%
    as.data.frame() %>%
    mutate(row_name = rownames(.)) %>%
    pivot_longer(-row_name, names_to = "col_name", values_to = "value") %>%
    filter(row_name != col_name) %>%
    mutate(pair_id = map2_chr(row_name, col_name, ~ paste(sort(c(.x, .y)), collapse = "_"))) %>%
    distinct(pair_id, .keep_all = TRUE) %>%
    separate(row_name, into = c("metric1","scanner1","suffix1"), sep = "_") %>%
    separate(col_name, into = c("metric2","scanner2","suffix2"), sep = "_") 
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_scanner_diff"] <- transf_mat %>%
    filter(scanner1 == scanner2) %>%
    filter(suffix1 != suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
  scanner_software_comp[scanner_software_comp$ROI == outcome, "within_software_diff"] <- transf_mat %>%
    filter(scanner1 != scanner2) %>%
    filter(suffix1 == suffix2) %>%
    summarise(mean_value = mean(value, na.rm = TRUE))
}
scanner_software_comp[,2:3] <- round(scanner_software_comp[,2:3], digits = 2)
```

```{r}
#| label: tbl-between-within
#| tbl-cap: "Average percent differences between softwares and between scanners in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
scanner_software_comp |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "between software", "between scanners"),
    booktabs  = TRUE,
    linesep   = "",
    format    = "latex",
    align     = c("l", "c", "c")
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; between software = average percent difference comparing two segmentations with different software of the same image;",
      "between scanners = average percent difference comparing two segmentations with identical software of images from different scanners"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

::: landscape
```{r}
#| label: tbl-iced-icc
#| tbl-cap: "ICED ICCs before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- lapply(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], as.numeric)
iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] <- round(iced[, c(paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))], digits = 2)
iced[, c("ROI", paste0(rep(c("ICC_SKYRA_cat", "ICC_VERIO_cat", "ICC_SKYRA_recon6", "ICC_VERIO_recon6", 
                              "ICC_SKYRA_recon7", "ICC_VERIO_recon7", "ICC_SKYRA_samseg", "ICC_VERIO_samseg"), 
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm.",
                  "ICC", "harm.", "ICC", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "              = 1,
    "SKYRA cat"      = 2,
    "VERIO cat"      = 2,
    "SKYRA recon6"   = 2,
    "VERIO recon6"   = 2,
    "SKYRA recon7"   = 2,
    "VERIO recon7"   = 2,
    "SKYRA samseg"   = 2,
    "VERIO samseg"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; harm. = harmonized ICC; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-variances
#| tbl-cap: "ICED variances before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truevar", "SKYRAvar", "VERIOvar", "catvar", "recon6var", "recon7var", "samsegvar", "obsvar"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm.",
                  "var", "harm.", "var", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 16))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "SKYRA"    = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2,
    "samseg"   = 2,
    "residual"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9, 11, 13, 15), border_right = TRUE) |>
  footnote(
    general = paste0(
      "ROI = region of interest; var = latent variable variance; harm. = harmonized latent variable variance; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-iced-means
#| tbl-cap: "ICED means before and after harmonization in cross-sectional data"
#| echo: false
#| warning: false
#| message: false
#| error: false
iced[, c("ROI", paste0(rep(c("truemean", "VERIOmean", "catmean", "recon6mean", "recon7mean"),
                            each = 2), c("", "_adj")))] |>
  kbl(
    row.names = FALSE,
    col.names = c("ROI",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm.", "mean", "harm.",
                  "mean", "harm."),
    booktabs  = TRUE,
    linesep   = "",
    align     = c("l", rep("c", 10))
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  add_header_above(c(
    " "        = 1,
    "true"  = 2,
    "VERIO"    = 2,
    "cat"      = 2,
    "recon6"   = 2,
    "recon7"   = 2
  )) |>
  column_spec(1, border_right = TRUE) |>
  column_spec(c(3, 5, 7, 9), border_right = TRUE) |>
  footnote(
    general = paste0(
      "true = true score with SKYRA and samseg as reference scanner and segmentation, respectively; ",
      "ROI = region of interest; mean = latent variable mean; harm. = harmonized latent variable mean; ",
      "TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```
:::

```{r update_prep_no_correction}
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_mat <- update_pd_matrices
pd_sub2 <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(pd_sub2) <- c("ROI", "pre", "post", "mean_pd_pre", "mean_pd_post")
pd_sub2$ROI <- c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction")]]
  post <- pd_mat[[paste0(outcome, "_mean_pd_mat_no_correction_adj")]]
  n_comparisons <- sum(1:(ncol(pre)-1))
  n_exceeding_pre <- length(which(pre >2)) # enumerate the comparisons with a PD > 2%
  n_exceeding_post <- length(which(post >2)) # enumerate the comparisons with a PD > 2%
  pd_sub2[pd_sub2$ROI == outcome, "pre"] <- paste0(n_exceeding_pre, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "post"] <- paste0(n_exceeding_post, "/", n_comparisons)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_pre"] <- mean(abs(pre), na.rm = T)
  pd_sub2[pd_sub2$ROI == outcome, "mean_pd_post"] <- mean(abs(post), na.rm = T)
}
pd_sub2$mean_pd_pre <- round(pd_sub2$mean_pd_pre, digits = 2)
pd_sub2$mean_pd_post <- round(pd_sub2$mean_pd_post, digits = 3)
anovas <- update_anova_res
pwcs <- update_pwcs
for (outcome in c("TIV", "TGMV", "TWMV", "TCV", "LVV",  "HCV", "AV")){
  pre <- pwcs[[paste0(outcome, "_pwc_no_correction")]]
  post <- pwcs[[paste0(outcome, "_pwc_no_correction_adj")]]
  n_pwcs <- nrow(pre)
  n_sig_pre <- length(which(pre$p.adj.signif != "ns"))
  n_sig_post <- length(which(post$p.adj.signif != "ns"))
  anovas[anovas$ROI == outcome, "n_sig_pre"] <- paste0(n_sig_pre, "/", n_pwcs)
  anovas[anovas$ROI == outcome, "n_sig_post"] <- paste0(n_sig_post, "/", n_pwcs)
}
anovas$p_no_correction <- ifelse(anovas$p_no_correction < 0.001, formatC(anovas$p_no_correction, format = "e", digits = 2), anovas$p_no_correction)
anovas$p_no_correction_adj <- ifelse(anovas$p_no_correction_adj < 0.001, formatC(anovas$p_no_correction_adj, format = "e", digits = 2), anovas$p_no_correction_adj)
```

```{r}
#| label: tbl-update-icc-no-correction
#| tbl-cap: "ICCs before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(knitr)
library(kableExtra)
df2 <- update_lme_icc_res_no_wmhv %>%
  mutate(
    ICC_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction, ICC_lower_95_CI_no_correction, ICC_upper_95_CI_no_correction),
    ICC_adj_CI = sprintf("%.2f (%.2f, %.2f)", ICC_no_correction_adj, ICC_lower_95_CI_no_correction_adj, ICC_upper_95_CI_no_correction_adj))

df2[, c("ROI", "ICC_CI", "ICC_adj_CI")] |>
  kbl(
    col.names = c(
      "ROI",
      "ICC (95% CI)",
      "ICC (95% CI) harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = "lccc"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2:4, width = "4cm") |>
  footnote(
    general = paste0(
      "ROI = region of interest; ICC = intraclass correlation coefficient; ",
      "CI = confidence interval; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; ",
      "TGMV = total grey matter volume; TWMV = total white matter volume; ",
      "TCV = total cortical volume; LVV = bilateral lateral ventricle volume; ",
      "HCV = bilateral hippocampal volume; AV = bilateral amygdala volume"
    ),
    general_title = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-pd-no-correction
#| tbl-cap: "Percent differences before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
pd_sub2 |>
  kbl(
    col.names = c(
      "ROI",
      "PD > 2%",
      "PD > 2% harm.",
      "mean PD",
      "mean PD harm."
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r") 
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
    general = paste0(
      "ROI = region of interest; PD = percentage difference; harm. = harmonized with neuroHarmonize; ",
      "TIV = total intracranial volume; TGMV = total grey matter volume; ",
      "TWMV = total white matter volume; TCV = total cortical volume; ",
      "LVV = bilateral lateral ventricle volume; HCV = bilateral hippocampal volume; ",
      "AV = bilateral amygdala volume"
    ),
    general_title  = "",
    threeparttable = TRUE
  )
```

```{r}
#| label: tbl-update-signif-no-correction
#| tbl-cap: "Statistical significance of differences before and after harmonization without manually edited scans - cross-sectional dataset"
#| echo: false
#| warning: false
#| message: false
#| error: false
anovas[, c("ROI", "p_no_correction", "ges_no_correction", "n_sig_pre", "p_no_correction_adj", "ges_no_correction_adj", "n_sig_post")] |>
  kbl(
    escape = FALSE,
    col.names = c(
      "ROI",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs",
      "p-value",
      "$\\eta^2$",
      "sig. PWCs"
    ),
    booktabs = TRUE,
    linesep  = "",
    align    = c("l", "r", "r", "r", "r", "r", "r")
  ) |>
    add_header_above(c(
    " " = 1,
    "unadjusted" = 3,
    "harmonized" = 3
  )) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size     = 10
  ) |>
  footnote(
  general = paste0(
    "ROI = region of interest; p-value = p-value from ANOVA; ",
    "$\\\\eta^2$", " = generalized ", "$\\\\eta^2$", " from ANOVA;",
    "harm. = harmonized with neuroHarmonize; ",
    "sig. PWCs = share of significant pairwise comparisons; ",
    "TIV = total intracranial volume; TGMV = total grey matter volume; ",
    "TWMV = total white matter volume; TCV = total cortical volume; ",
    "LVV = bilateral lateral ventricle volume; ",
    "HCV = bilateral hippocampal volume; ",
    "AV = bilateral amygdala volume"
  ),
  general_title  = "",
  threeparttable = TRUE,
  escape = FALSE
  )
```

# Reproduction of supplementary figures

```{r update_raincloud_prep}
#| echo: false
#| warning: false
#| message: false
#| error: false
library(ggdist)
df <- read.csv("https://osf.io/deru3/?action=download", sep = ",", header=TRUE)
df_wmhv <- read.csv("https://osf.io/mxde5/?action=download", sep = ",", header=TRUE)
df$SITE <- gsub("_", " ", df$SITE)
df$SITE <- factor(df$SITE, levels = c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat"))  
df_wmhv$asinh <- asinh(df_wmhv$Lesions)
# retransform LVV
df$LVV_asinh <- df$LVV
df$LVV <- sinh(df$LVV)
colors = met.brewer(name="Java", n=5, type="discrete")
colors <- colors[c(1,2,4,5)]
colors <- rep(colors, each = 2)
make_raincloud <- function(data, col, group_val, fill_color, dotsize = 1.25) {
  data_filtered <- data %>% filter(SITE == group_val)
  
  ggplot(data_filtered, aes(x = 1, y = .data[[col]])) +
    stat_halfeye(adjust = 0.5, width = 0.6, justification = -0.2,
                 .width = 0, point_colour = NA, fill = fill_color) +
    geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.5) +
    stat_dots(side = "left", justification = 1.1, dotsize = dotsize) +
    coord_flip() +
    labs(title = as.character(group_val), x = NULL, y = col) +
    theme_classic() +
    theme(axis.text.y = element_blank(), axis.title.x = element_blank(), axis.ticks.y = element_blank())
}
# Create one plot per column
cols <- c("TIV", "TGMV", "TWMV", "TCV", "LVV", "LVV_asinh", "HCV", "AV")
group_vals <- c("SKYRA recon7", "VERIO recon7", "SKYRA recon6", "VERIO recon6", "SKYRA samseg", "VERIO samseg", "SKYRA cat", "VERIO cat")
plots_by_col <- lapply(cols, function(col) {
  
  group_plots <- mapply(make_raincloud,
                        group_val = group_vals,
                        fill_color = colors,
                        MoreArgs = list(data = df, col = col),
                        SIMPLIFY = FALSE)
  
  wrap_plots(group_plots, ncol = 2)
})

# Name the list by column for easy access
names(plots_by_col) <- cols

# create plots for WMHV
# remove highest values to enable visualisation
df_wmhv_cut <- df_wmhv[df_wmhv$Lesions < 5000,] 
add_row_label <- function(plot_list, label) {
  plot_list[[1]] <- plot_list[[1]] + 
    labs(x = label) +
    theme(axis.title.y = element_text(angle = 90, vjust = 0.5))
  plot_list
}
unharmonized <- mapply(make_raincloud,
       group_val = c("SKYRA", "VERIO"),
       fill_color = colors[5:6],
       MoreArgs = list(data = df_wmhv_cut, col = "Lesions"),
       SIMPLIFY = FALSE)
asinh  <- mapply(make_raincloud,
                 group_val = c("SKYRA", "VERIO"),
                 fill_color = colors[5:6],
                 MoreArgs = list(data = df_wmhv_cut, col = "asinh"),
                 SIMPLIFY = FALSE)
norm  <- mapply(make_raincloud,
                 group_val = c("SKYRA", "VERIO"),
                 fill_color = colors[5:6],
                 MoreArgs = list(data = df_wmhv_cut, col = "WMHV_norm"),
                 SIMPLIFY = FALSE)
unharmonized <- add_row_label(unharmonized, "original")
asinh  <- add_row_label(asinh, "asinh")
norm   <- add_row_label(norm, expression("log"[10]))
wmhv_plots <- wrap_plots(unname(c(unharmonized, asinh, norm)), nrow = 3)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tiv-raincloud
#| fig-cap: "Raincloud plots of total intracranial volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TIV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tgmv-raincloud
#| fig-cap: "Raincloud plots of total grey matter volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TGMV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-twmv-raincloud
#| fig-cap: "Raincloud plots of total white matter volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TWMV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-tcv-raincloud
#| fig-cap: "Raincloud plots of total cortical volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$TCV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-lvv-raincloud
#| fig-cap: "Raincloud plots of bilateral lateral ventricle volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$LVV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-lvv-asinh-raincloud
#| fig-cap: "Raincloud plots of bilateral lateral ventricle volume by batch after inverse hyperbolic sine transformation in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$LVV_asinh)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-hcv-raincloud
#| fig-cap: "Raincloud plots of bilateral hippocampal volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$HCV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-av-raincloud
#| fig-cap: "Raincloud plots of bilateral amygdala volume by batch in cross-sectional data."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(plots_by_col$AV)
```

```{r fig.width=6, fig.height=12}
#| label: fig-update-wmhv-raincloud
#| fig-cap: "Raincloud plots of white matter hyperintensity volume by batch in cross-sectional data. 1st row: original data. 2nd row: after inverse hyperbolic sine transformation. 3rd row: after decadic logarithm transformation. The maximum value is not depicted."
#| echo: false
#| warning: false
#| message: false
#| error: false
#| out-width: "100%"
#| out-height: "90%"
#| dpi: 900
print(wmhv_plots)
```

